### 2017-03-03 manuscript/manuscript.tex
edits to the manuscript, mostly to make sure I know how to make edits

\documentclass[twocolumn,aps,prd,superscriptaddress]{revtex4-1}  
\usepackage{amsmath,graphicx}

\begin{document}

\title{Here is our title}

\author{Justin A.\ Ellis}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}

\author{Kristina P.\ Islo}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}

\author{Stephen R.\ Taylor}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}

\author{Sarah J.\ Vigeland}
\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}

\date{\today}

\begin{abstract}
Here is our abstract.
\end{abstract}

\maketitle

\section{Introduction}

\section{Posterior Averaging}

\section{Compare Significance with Sky Scrambles/Phase Shifts}

\section{Conclusion}

\acknowledgments
We thank

JAE was supported by NASA through Einstein Fellowship Grant PF4-150120. 
SRT was supported by appointment to the NASA Postdoctoral Program 
at the Jet Propulsion Laboratory, which is administered by Oak Ridge Associated Universities 
and the Universities Space Research Association through a contract with NASA. 
KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.

\end{document}

### 2017-06-28 manuscript/manuscript.tex
changes to manuscript, plots added, etc

@@
 
+\DeclareMathOperator{\Tr}{Tr}
+
+\usepackage[usenames]{color}
+\newcommand{\sv}[1]{\textcolor{blue}{\it{\textbf{sv: #1}}} }
+\newcommand{\ki}[1]{\textcolor{green}{\it{\textbf{ki: #1}}} }
+
+\newcommand{\Agw}{\ensuremath{A_\mathrm{gw}}}
+
 \begin{document}
 
-\title{Here is our title}
+\title{Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic}
 
-\author{Justin A.\ Ellis}
-\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}
+\author{Sarah J.\ Vigeland}
+\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
 
@@
 
-\author{Sarah J.\ Vigeland}
-\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
+\author{Justin A.\ Ellis}
+\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}
 
@@
 
-\section{Posterior Averaging}
+Pulsar timing arrays (PTAs) \sv{blah blah blah}
+
+The optimal statistic is a frequentist estimator for the stochastic background amplitude. 
+It was derived by \citet{abc+2009} \sv{blah blah blah... try not to plagiarize myself}
+
+The optimal statistic is an important complement to Bayesian detection techniques. 
+It provides an independent way to assess the significance of the stochastic background, 
+and it is significantly faster to compute. It is also easy to test different spatial correlations 
+by altering the overlap reduction function $\Gamma$ in the cross-correlations. 
+In contrast, doing a full Bayesian analysis with spatial correlations is a very computationally 
+intensive procedure.
+
+However, previous work has found that the optimal statistic gives biased results \sv{expand...} 
+In order to compute the optimal statistic, the timing model parameters are fixed 
+at the maximum-likelihood values, but this leads to biases in the 
+optimal statistic value and uncertainty due to covariances between the 
+individual red noise parameters and the GW amplitude.
+
+In this paper, we present a technique for improving the accuracy of the optimal statistic 
+by marginalizing the optimal statistic over the individual red noise parameters 
+using the posterior distributions from a full Bayesian analysis. 
+We demonstrate that this hybrid approach produces a value and uncertainty 
+of the optimal statistic that is in better agreement to the $\Agw$ 
+produced from a full Bayesian analysis.
+
+This paper is organized as follows. In Sec.~\ref{sec:marg_os} 
+we lay out the procedure for computing the noise-marginalized optimal statistic 
+and compare to the fixed-noise optimal statistic. 
+In Sec.~\ref{sec:spatial} determine how well 
+the noise-marginalized optimal statistic can differentiate between different spatial correlations 
+using simulated PTA data. 
+In Sec.~\ref{sec:skyscrambles} we determine how well the noise-marginalized optimal statistic 
+can assess the significance of a GW detection by performing sky scrambles \citep{cs2016}. 
+and compare to the Bayesian analysis done by \citet{tlb+2017}. 
+We conclude in Sec.~\ref{sec:conclusion}.
+
+\section{Noise-marginalized optimal statistic}
+\label{sec:marg_os}
+
+As introduced by \citet{abc+2009}, 
+the optimal statistic is derived by analytically maximizing 
+the PTA likelihood function in the weak-signal regime. 
+The likelihood function is given by
+\begin{equation}
+	p({\bf{r}}|\phi) = \frac{\exp[-\frac{1}{2} {\bf{r}}^T \Sigma^{-1}(\phi) {\bf{r}} ]}{\sqrt{\mathrm{det}[2\pi \Sigma(\phi)]}} \,,
+\end{equation}
+where $\bf{r}$ is an array of the timing residuals 
+from the $M$ pulsars in the PTA,
+\begin{equation}
+	\bf{r} = \left[ \begin{array}{c} {\bf{r}}_1 \\ {\bf{r}}_2 \\ \vdots \\ {\bf{r}}_M \end{array} \right] \,,
+\end{equation}
+$\phi$ is the set of pulsar noise parameters, 
+and $\Sigma$ is the covariance matrix of the residuals. 
+We can write $\Sigma$ in terms of 
+autocovariance and cross-covariance matrices ${\bf{P}}_I$ and ${\bf{S}}_{IJ}$, respectively, as
+\begin{equation}
+	\Sigma = \left[ \begin{array}{cccc} {\bf{P}}_1 & {\bf{S}}_{12} & \hdots & {\bf{S}}_{1M}  \\
+							{\bf{S}}_{21} & {\bf{P}}_2 & \hdots & {\bf{S}}_{2M} \\
+							\vdots & \vdots & \ddots & \vdots \\
+							{\bf{S}}_{M1} & {\bf{S}}_{M2} & \hdots & {\bf{P}}_M \end{array} \right] \,,
+\end{equation}
+where
+\begin{eqnarray}
+	{\bf{P}} &=& \left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle \,, \\
+	{\bf{S}}_{IJ} &=& \left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{I \neq J} \,.
+\end{eqnarray}
+\sv{something about the overlap reduction function...}
+
+Analytically maximizing this likelihood gives the optimal statistic,
+\begin{equation}
+	\hat{A}^2 = \frac{\sum_{IJ} {\bf{r}}_I^T {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} {\bf{r}}_J}{\sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) } \,,
+\end{equation}
+where $\tilde{{\bf{S}}}_{IJ}$ is the amplitude-independent cross-correlation matrix,
+\begin{equation}
+	\Agw^2 \tilde{{\bf{S}}}_{IJ} = {\bf{S}}_{IJ} \,.
+\end{equation}
+The normalization of $\hat{A}^2$ has been chosen so that 
+$\langle \hat{A}^2 \rangle = \Agw^2$. 
+For a weak GW signal, the standard deviation is 
+\begin{equation}
+	\sigma_0 = \left[ \sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{-1/2} \,,
+\end{equation}
+and the signal-to-noise ratio (SNR) is
+\begin{equation}
+	\hat{\rho} = \frac{\hat{A}^2}{\sigma_0} = \frac{\sum_{IJ} {\bf{r}}_I^T {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} {\bf{r}}_J}{\left[ \sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{1/2}} \,.
+\end{equation}
+The expectation value of the SNR, averaged over all realizations of the stochastic background, is
+\begin{equation}
+	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{IJ} \Tr\left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{1/2} \,.
+\end{equation}
+
+When constructing the optimal statistic, we typically fix the red noise parameters $\phi$ to the values that 
+maximize the likelihood, which we determine from a Bayesian common red noise analysis 
+\sv{check this part}. However, 
+doing this biases the optimal statistic and SNR \sv{blah blah blah} 
+In this paper we introduce a procedure to correct this bias by 
+marginalizing over the red noise parameters using 
+posterior distributions from a Bayesian analysis.
+
+In order to compare the noise-marginalized optimal statistic to the 
+fixed-noise optimal statistic, we simulated a stochastic background 
+in a PTA with \sv{fill in simulation details...} 
+The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
+To compute the fixed-noise optimal statistic, the red noise parameters are 
+fixed to the maximum likelihood values found using \sv{Kristina fill in here...}
+
+Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
+optimal statistic for a single realization of the stochastic background, 
+and Fig.~\ref{fig:os_datasetstats} compares the fixed-noise values of $\hat{A}^2$ 
+to the mean $\hat{A}^2$ for each realization of the stochastic background. 
+The fixed-noise analysis tends to underestimate the strength of the stochastic background, 
+whereas the noise-marginalized analysis is better able to recover the injected value. 
+The mean value of the optimal statistic from the fixed-noise analysis, 
+averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
+while the mean value from the noise-marginalized analysis is $\hat{A}^2 = 2.2 \times 10^{-29}$.
+
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
+	\caption{Optimal statistic and SNR for a simulated dataset 
+			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using a fixed-noise analysis (blue, dashed line) to the 
+			mean optimal statistic and mean SNR found by marginalizing over 
+			the individual pulsars' red noise parameters (orange, solid line). 
+			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			For this particular realization of the stochastic background, the fixed-noise analysis 
+			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
+			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$.}
+	\label{fig:os_dataset_sample}
+\end{figure}
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15.pdf}
+	\caption{Optimal statistic and SNR for \sv{insert number} simulated datasets 
+			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using a fixed-noise analysis (blue) to the 
+			mean optimal statistic and mean SNR found by marginalizing over 
+			the individual pulsars' red noise parameters. The dashed vertical line 
+			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			The value of $\hat{A}^2$ from the fixed-noise analysis, 
+			averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
+			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.2 \times 10^{-29}$.}
+	\label{fig:os_datasetstats}
+\end{figure}
+
+\sv{say something about uncertainty, spread in marg os, etc}
+
+\section{Compare Significance with Monopole and Dipole}
+\label{sec:spatial}
+
+The optimal statistic
+
+\citep{thk+2016}
 
 \section{Compare Significance with Sky Scrambles/Phase Shifts}
+\label{sec:skyscrambles}
+
+Another way to test the spatial correlations is through sky scrambles \sv{blah blah blah...} 
+\citet{cs2016} \citet{tlb+2017}
+
+The optimal statistic provides a much faster way to perform sky scrambles compared to 
+a Bayesian analysis since it only requires altering the overlap reduction function. 
+We performed sky scrambles by \sv{explain here...} 
+The scrambled sky positions were generated \sv{Steve, explain here...}
+
+Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 
+for a particular data set to the 
+distribution of noise-marginalized optimal statistics obtained using sky scrambles.
+\sv{sky scrambles for all of our simulated data sets}
+
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{plots/skyscrambles_dataset50.pdf}
+	\caption{Comparison between the optimal statistic with and without sky scrambles 
+			for a simulated dataset containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			The true analysis found an $\mathrm{SNR} = 1.1$, while only 24 of the 210 sky scramble analyses 
+			found $\mathrm{SNR} \geq 1.1$ ($p=0.11$).}
+	\label{fig:skyscrambles_dataset_sample}
+\end{figure}
+
+\sv{comparison to Steve's}
 
 \section{Conclusion}
+\label{sec:conclusion}
 
@@
 
+\bibliographystyle{apsrev}
+\bibliography{master}
+
 \end{document}

### 2017-06-30 manuscript/manuscript.tex
Adding rudimentary text

@@
 \newcommand{\sv}[1]{\textcolor{blue}{\it{\textbf{sv: #1}}} }
-\newcommand{\ki}[1]{\textcolor{green}{\it{\textbf{ki: #1}}} }
+\newcommand{\ki}[1]{\textcolor{cyan}{\it{\textbf{ki: #1}}} }
 
@@
 \begin{abstract}
-Here is our abstract.
+Supermassive black hole binaries (SMBHBs) are predicted to form a high-redshift cosmological population of low-frequency gravitational wave sources which added together produce a stochastic gravitational wave background (GWB) potentially measurable via pulsar timing.
+Comparing the actual and predicted times-of-arrival (TOAs) of radio pulses beamed by millisecond pulsars allows us to search for and place limits on the stochastic background.
+While most estimating of GWB significance is done via Bayesian analysis, here we use a frequentist estimator called the noise-marginalized optimal statistic to assess the signifcance of a GWB present in simulated PTA data.
+We find \ki{complete}
+
 \end{abstract}
@@
 
-Pulsar timing arrays (PTAs) \sv{blah blah blah}
+Pulsar timing arrays (PTAs) are a conglomerate of pulsars with millisecond periods.
+Monitioring the ultra-regular emission from these pulsars with radio telescopes enables us to probe the dynamics of the spacetime through which the radio pulses travel.
+This makes PTAs ideal complements to ground- and space-based gravitational wave (GW) detectors.
+A PTAs sensitivity is constrained by single pulsar observation cadence and total experiment length.
+For NANOGrav, this means GW frequencies in the range $10^{-9} - 10^{-7}$ Hz.
+\ki{citations needed} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime.
+
+Most measuring of the stochastic background is done through Bayesian data analysis.
+\ki{baye baye baye stuff} There are many advantages to running full Bayesian analyses on data sets, but they are slow.
+The bulk of the work consists of inverting non-diagonal matrices to find the strength of Hellings-Downs cross-correlations between pulsar pairs -- the "smoking gun" of a stochastic GWB.
+
+Determining the significance of various cross-correlations between pulsar pairs is computationally expensive, requiring multi-core computing clusters many days to complete.
+As PTA data-analysis expands to include more parameters when measuring our predicted signal, e.g.
+increase in residuals with continued and new pulsar observations, uncertainty in solar system ephemerides, etc., computation time will only increase.
+Thus methods which hasten the production of results without compromising accuracy are not only preferable, but will be necessary.
+Here we explore the noise-marginalized optimal statistic as an estimator for GWB signals.
 

### 2017-06-30 manuscript/manuscript.tex
expanded Sec. 2

--- 2017-06-28 manuscript/manuscript.tex
@@
 
-Pulsar timing arrays (PTAs) \sv{blah blah blah}
+Pulsar timing arrays (PTAs) can be used to detect low-frequency gravitational waves (GWs) \sv{blah blah blah}
 
@@
 using the posterior distributions from a full Bayesian analysis. 
-We demonstrate that this hybrid approach produces a value and uncertainty 
-of the optimal statistic that is in better agreement to the $\Agw$ 
-produced from a full Bayesian analysis.
+This hybrid approach produces a more accurate 
+estimate of \Agw\ and its uncertainty.
 
@@
 In Sec.~\ref{sec:spatial} determine how well 
-the noise-marginalized optimal statistic can differentiate between different spatial correlations 
+the noise-marginalized optimal statistic can 
+differentiate between different spatial correlations 
 using simulated PTA data. 
@@
 In order to compare the noise-marginalized optimal statistic to the 
-fixed-noise optimal statistic, we simulated a stochastic background 
-in a PTA with \sv{fill in simulation details...} 
+fixed-noise optimal statistic, we injected a GW stochastic background with 
+$\Agw = 5 \times 10^{-15}$ into simulated PTA data. 
+We constructed simulated data sets using the same procedure as in \citet{tlb+2017}, 
+which consisted of 10 MSPs with \sv{Steve fill in here...}
+
 The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
-To compute the fixed-noise optimal statistic, the red noise parameters are 
-fixed to the maximum likelihood values found using \sv{Kristina fill in here...}
+We incorporate the individual pulsar noise parameters as follows. 
+We perform individual Bayesian noise analyses on all of the pulsars, and 
+fix the white noise to the maximum likelihood value. 
+To compute the fixed-noise optimal statistic, 
+the red noise parameters for each pulsar are set to the mode of 
+the joint posterior distribution for $A_\mathrm{red}$,$\gamma_\mathrm{red}$, 
+found by performing a Bayesian common red noise analysis. 
+For the noise-marginalized optimal statistic, we draw from 
+the common red noise posteriors in order to average the optimal statistic over 
+the individual pulsars' red noise.
 
@@
 \end{figure}
+
+\sv{compare to Bayesian analysis for a particular data set}
+
 \begin{figure}[ht]
@@
 
-The optimal statistic
+The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
+because using a different spatial correlation only requires changing the overlap reduction function 
+in $\tilde{{\bf{S}}}_{IJ}$. 
+\citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
+multiple spatial correlations at once in order to mitigate common noise sources such as 
+clock error and ephemeris error. 
+Here we compute the optimal statistic with monopole and dipole spatial correlations 
+for the same simulated data sets as in the previous section in order to determine 
+how well we can distinguish a GW background from a monopole or dipole signal. 
+For a monopole signal, the overlap reduction function becomes simply
+\[ \Gamma(\theta_{IJ}) = 1 \,\]
+while for a dipole signal, the overlap reduction function becomes
+\[ \Gamma(\theta_{IJ}) = \cos\theta_{IJ} \,. \]
 
-\citep{thk+2016}
+Figure~\ref{fig:} shows the mean value of $\hat{A}^2$ and the mean SNR 
+computed assuming monopole, dipole, and Hellings-Downs spatial correlations 
+for \sv{insert number} simulated data sets.
 
@@
 
-Another way to test the spatial correlations is through sky scrambles \sv{blah blah blah...} 
-\citet{cs2016} \citet{tlb+2017}
+Another way to test the spatial correlations is through ``sky scrambles,'' 
+where the overlap reduction function is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
+\citet{tlb+2017} showed how sky scrambles affect the Bayes' factor for simulated data sets. 
+Here we perform a similar analysis using frequentist methods.
 
-The optimal statistic provides a much faster way to perform sky scrambles compared to 
-a Bayesian analysis since it only requires altering the overlap reduction function. 
-We performed sky scrambles by \sv{explain here...} 
 The scrambled sky positions were generated \sv{Steve, explain here...}

### 2017-06-30 manuscript/manuscript.tex

@@
 
-Pulsar timing arrays (PTAs) can be used to detect low-frequency gravitational waves (GWs) \sv{blah blah blah}
+Pulsar timing arrays (PTAs) are a conglomerate of pulsars with millisecond periods.
+Monitioring the ultra-regular emission from these pulsars with radio telescopes enables us to probe the dynamics of the spacetime through which the radio pulses travel.
+This makes PTAs ideal complements to ground- and space-based gravitational wave (GW) detectors.
+A PTAs sensitivity is constrained by single pulsar observation cadence and total experiment length.
+For NANOGrav, this means GW frequencies in the range $10^{-9} - 10^{-7}$ Hz.
+\ki{citations needed} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime.
+
+Most measuring of the stochastic background is done through Bayesian data analysis.
+\ki{baye baye baye stuff} There are many advantages to running full Bayesian analyses on data sets, but they are slow.
+The bulk of the work consists of inverting non-diagonal matrices to find the strength of Hellings-Downs cross-correlations between pulsar pairs -- the "smoking gun" of a stochastic GWB.
+
+Determining the significance of various cross-correlations between pulsar pairs is computationally expensive, requiring multi-core computing clusters many days to complete.
+As PTA data-analysis expands to include more parameters when measuring our predicted signal, e.g.
+increase in residuals with continued and new pulsar observations, uncertainty in solar system ephemerides, etc., computation time will only increase.
+Thus methods which hasten the production of results without compromising accuracy are not only preferable, but will be necessary.
+Here we explore the noise-marginalized optimal statistic as an estimator for GWB signals.
 

### 2017-07-31 manuscript/manuscript.tex
added plots

@@
 Comparing the actual and predicted times-of-arrival (TOAs) of radio pulses beamed by millisecond pulsars allows us to search for and place limits on the stochastic background.
-While most estimating of GWB significance is done via Bayesian analysis, here we use a frequentist estimator called the noise-marginalized optimal statistic to assess the signifcance of a GWB present in simulated PTA data.
+While most estimating of GWB significance is done via Bayesian analysis, here we use a frequentist estimator called the noise-marginalized optimal statistic to assess the significance of a GWB present in simulated PTA data.
 We find \ki{complete}
@@
 Pulsar timing arrays (PTAs) are a conglomerate of pulsars with millisecond periods. 
-Monitioring the ultra-regular emission from these pulsars with radio telescopes enables us to probe the dynamics of the spacetime through which the radio pulses travel.
-This makes PTAs ideal complements to ground- and space-based gravitational wave (GW) detectors.
-A PTAs sensitivity is constrained by single pulsar observation cadence and total experiment length.
-For NANOGrav, this means GW frequencies in the range $10^{-9} - 10^{-7}$ Hz.
-\ki{citations needed} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime.
+Monitoring the ultra-regular periodic emission from these pulsars with radio telescopes 
+enables us to probe the dynamics of the spacetime through which the pulses travel. 
+PTAs are sensitive to gravitational waves (GWs) with frequencies of 
+$10^{-9} - 10^{-7}\;\mathrm{Hz}$ \ki{citations needed}, 
+which makes PTAs ideal complements to ground- and space-based GW detectors \sv{add citations}. 
+\sv{something...} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime.
 
@@
 The mean value of the optimal statistic from the fixed-noise analysis, 
-averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
-while the mean value from the noise-marginalized analysis is $\hat{A}^2 = 2.2 \times 10^{-29}$.
+averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
+while the mean value from the noise-marginalized analysis is $\hat{A}^2 = 2.5 \times 10^{-29}$.
 
@@
 	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15.pdf}
-	\caption{Optimal statistic and SNR for \sv{insert number} simulated datasets 
+	\caption{Optimal statistic and SNR for 425 simulated datasets 
 			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
@@
 			The value of $\hat{A}^2$ from the fixed-noise analysis, 
-			averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
-			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.2 \times 10^{-29}$.}
+			averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
+			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$.}
 	\label{fig:os_datasetstats}
@@
 
-Figure~\ref{fig:} shows the mean value of $\hat{A}^2$ and the mean SNR 
+Figure~\ref{fig:os_ORF} shows the mean value of $\hat{A}^2$ and the mean SNR 
 computed assuming monopole, dipole, and Hellings-Downs spatial correlations 
-for \sv{insert number} simulated data sets.
+for 425 simulated data sets.
+
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
+			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using monopole (blue), dipole (orange), and quadrupole (green) 
+			spatial correlations. 
+			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
+	\label{fig:os_ORF}
+\end{figure}
 

### 2017-07-31 manuscript/manuscript.tex
more edits to Sec. III

@@
 which makes PTAs ideal complements to ground- and space-based GW detectors \sv{add citations}. 
-\sv{something...} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime.
+\sv{something...} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime \sv{add citations}.
 
@@
 and it is significantly faster to compute. It is also easy to test different spatial correlations 
-by altering the overlap reduction function $\Gamma$ in the cross-correlations. 
+by altering the overlap reduction function (ORF) $\Gamma$ in the cross-correlations. 
 In contrast, doing a full Bayesian analysis with spatial correlations is a very computationally 
@@
 \end{eqnarray}
-\sv{something about the overlap reduction function...}
+\sv{something about the ORF...}
 
@@
 while the mean value from the noise-marginalized analysis is $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-
+The noise-marginalized analysis also gives a higher average mean SNR of 1.7, 
+compared to an average SNR of 1.0 using the fixed-noise analysis.
 \begin{figure}[ht]
@@
 			averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
-			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$.}
+			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			Furthermore, the fixed-noise analysis gives an average SNR of 1.0, while the 
+			noise-marginalized analysis gives an average mean SNR of 1.7.}
 	\label{fig:os_datasetstats}
@@
 The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
-because using a different spatial correlation only requires changing the overlap reduction function 
+because using a different spatial correlation only requires changing the ORF 
 in $\tilde{{\bf{S}}}_{IJ}$. 
@@
 how well we can distinguish a GW background from a monopole or dipole signal. 
-For a monopole signal, the overlap reduction function becomes simply
+For a monopole signal, the ORF becomes simply
 \[ \Gamma(\theta_{IJ}) = 1 \,\]
-while for a dipole signal, the overlap reduction function becomes
+while for a dipole signal, the ORF becomes
 \[ \Gamma(\theta_{IJ}) = \cos\theta_{IJ} \,. \]
 
-Figure~\ref{fig:os_ORF} shows the mean value of $\hat{A}^2$ and the mean SNR 
-computed assuming monopole, dipole, and Hellings-Downs spatial correlations 
-for 425 simulated data sets.
-
+Figure~\ref{fig:os_ORF} shows the noise-marginalized mean value of $\hat{A}^2$ and the mean SNR 
+computed assuming monopole, dipole, and quadrupole spatial correlations 
+for 425 simulated data sets. Using a monopole or dipole ORF 
+gives a lower value for the mean optimal statistic and mean SNR compared to the 
+Hellings-Downs ORF. 
+However, there is significant overlap between the distributions for the mean SNR using 
+monopole, dipole, and quadrupole ORFs, indicating that it is not easy to differentiate 
+between those three spatial correlations for an injected GW background of this amplitude. 
+We find a noise-marginalized mean SNR above 1.0 in 68\% of our simulated data sets 
+using the quadrupole ORF, and in 45\% and 24\% of our simulated data sets 
+using the monopole and dipole ORF, respectively.
 \begin{figure}[ht]
@@
 Another way to test the spatial correlations is through ``sky scrambles,'' 
-where the overlap reduction function is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
+where the ORF is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
 \citet{tlb+2017} showed how sky scrambles affect the Bayes' factor for simulated data sets. 
@@
 \acknowledgments
-We thank
-
+We thank Joe Romano and Xavier Siemens for useful discussions. 
 JAE was supported by NASA through Einstein Fellowship Grant PF4-150120. 

### 2017-07-31 codes/optstat.py
changing 2dmax option

@@ -140,19 +140,19 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
         if writenoise == '2dmax':
 
-            # create chain object and find 1d maxLL parameter values
-            cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
-            ml = cp.get_ml_values(mtype='marg')
+            # # create chain object and find 1d maxLL parameter values
+            # cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
+            # ml = cp.get_ml_values(mtype='marg')
             noisefile = noisedir + '/{0}_noise.txt'.format(psr)
             
             # load chain and find 2d maxLL values for RN parameters
-            pars = np.loadtxt('chains/{0}/noise/{1}/pars.txt'.format(dataset, psr), dtype='S42')
-            chain = np.loadtxt('chains/{0}/noise/{1}/chain_1.txt'.format(dataset, psr))
+            pars = np.loadtxt('../data/chains/{0}/fix_spec_nf_30/pars.txt'.format(dataset), dtype='S42')
+            chain = np.loadtxt('../data/chains/{0}/fix_spec_nf_30/chain_1.txt'.format(dataset))
             burn = int(0.25*chain.shape[0])
     
             # 2-d histogram 
-            RN_amplitude_chain = chain[:,np.argwhere(pars == 'RN-Amplitude')[0][0]]
-            RN_spectral_index_chain = chain[:,np.argwhere(pars == 'RN-spectral-index')[0][0]]
+            RN_amplitude_chain = chain[:,np.argwhere(pars == 'RN-Amplitude_{0}'.format(psr))[0][0]]
+            RN_spectral_index_chain = chain[:,np.argwhere(pars == 'RN-spectral-index_{0}'.format(psr))[0][0]]
             RN_Amplitude, RN_spectral_index = getMax2d(RN_amplitude_chain, RN_spectral_index_chain)
 
             with open(noisefile, 'w') as f:

### 2017-08-01 manuscript/manuscript.tex
added comparison to Bayesian analysis for a single dataset

@@
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
-optimal statistic for a single realization of the stochastic background, 
-and Fig.~\ref{fig:os_datasetstats} compares the fixed-noise values of $\hat{A}^2$ 
-to the mean $\hat{A}^2$ for each realization of the stochastic background. 
+optimal statistic for a single realization of the stochastic background. 
+We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
+for a common red process. For this particular realization of the stochastic background, 
+the values of $\hat{A}^2$ from the noise-marginalized and Bayesian analyses are in good agreement 
+with each other and the injected GW stochastic background, 
+but the value from the fixed-noise analysis is significantly lower. 
+The noise-marginalized analysis also gives a larger mean SNR compared to the fixed-noise analysis.
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
+	\caption{Optimal statistic and SNR for a simulated dataset 
+			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
+			mean optimal statistic and mean SNR found by marginalizing over 
+			the individual pulsars' red noise parameters (solid orange line). 
+			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
+			for a common red process (solid green line). 
+			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			For this particular realization of the stochastic background, the fixed-noise analysis 
+			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
+			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
+			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
+	\label{fig:os_dataset_sample}
+\end{figure}
+
+Figure~\ref{fig:os_datasetstats} compares the fixed-noise values of $\hat{A}^2$ 
+to the mean $\hat{A}^2$ for 425 different realizations of the stochastic background. 
 The fixed-noise analysis tends to underestimate the strength of the stochastic background, 
@@
 compared to an average SNR of 1.0 using the fixed-noise analysis.
-\begin{figure}[ht]
-	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
-	\caption{Optimal statistic and SNR for a simulated dataset 
-			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using a fixed-noise analysis (blue, dashed line) to the 
-			mean optimal statistic and mean SNR found by marginalizing over 
-			the individual pulsars' red noise parameters (orange, solid line). 
-			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			For this particular realization of the stochastic background, the fixed-noise analysis 
-			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
-			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$.}
-	\label{fig:os_dataset_sample}
-\end{figure}
-
-\sv{compare to Bayesian analysis for a particular data set}
-
 \begin{figure}[ht]
@@
 \end{figure}
-
-\sv{say something about uncertainty, spread in marg os, etc}
 
@@
 
-Another way to test the spatial correlations is through ``sky scrambles,'' 
+Another technique for testing the spatial correlations is through ``sky scrambles,'' 
 where the ORF is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 

### 2017-08-01 codes/optstat.py
clarifying noise-writing options

@@ -138,11 +138,11 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
                 for p,val in zip(pars,maxpost_sample):
                     f.write('%s %g\n'%(p, val))
 
-        if writenoise == '2dmax':
+        if writenoise == '2dmax_common':
 
-            # # create chain object and find 1d maxLL parameter values
-            # cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
-            # ml = cp.get_ml_values(mtype='marg')
+            # create chain object and find 1d maxLL parameter values
+            cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
+            ml = cp.get_ml_values(mtype='marg')
             noisefile = noisedir + '/{0}_noise.txt'.format(psr)
             
             # load chain and find 2d maxLL values for RN parameters
@@ -169,7 +169,37 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
                     else:
                         f.write('%s %g\n'%(key, val))
-            
+
+        if writenoise == '2dmax_indiv':
+            # create chain object and find 1d maxLL parameter values
+            cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
+            ml = cp.get_ml_values(mtype='marg')
+            noisefile = noisedir + '/{0}_noise.txt'.format(psr) 
+
+            # load chain and find 2d maxLL values for RN parameters
+            pars = np.loadtxt('chains/{0}/noise/{1}/pars.txt'.format(dataset, psr), dtype='S42')
+            chain = np.loadtxt('chains/{0}/noise/{1}/chain_1.txt'.format(dataset, psr))
+            burn = int(0.25*chain.shape[0])
+
+            RN_amplitude_chain = chain[:,np.argwhere(pars == 'RN-Amplitude')[0][0]]
+            RN_spectral_index_chain = chain[:,np.argwhere(pars == 'RN-spectral-index')[0][0]]
+
+
+             with open(noisefile, 'w') as f:
+
+                for key, val in ml.items():
+
+                    if key == 'RN-Amplitude':
+                        val = RN_Amplitude
+                        f.write('%s %g\n'%(key, val))
+
+                    elif key == 'RN-spectral-index':
+                        val = RN_spectral_index
+                        f.write('%s %g\n'%(key, val))
+
+                    else:
[…]

### 2017-08-01 codes/optstat.py
fixing noise chain location

@@ -104,8 +104,8 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
     for psr in psrs:
         print '   Running on ' + psr
         os.system('PAL2_run.py --h5File {0} --pulsar {1} \
-                  --outDir chains/{2}/noise/{1}/ --niter {3} \
-                  --mark9 --incRed --nf 30'.format(h5filename, psr, dataset, niter))
+                  --outDir {1}/chains/{3}/noise/{2}/ --niter {4} \
+                  --mark9 --incRed --nf 30'.format(h5filename, chaindir, psr, dataset, niter))
 
     # make noise files for each pulsar
     noisedir = '../data/noisefiles/' + dataset
@@ -116,7 +116,7 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
         
         if writenoise == '1dmax':
             # create chain object and find 1d maxLL values
-            cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
+            cp = pp.ChainPP('../data/chains/{0}/noise/{1}/'.format(dataset, psr))
             ml = cp.get_ml_values(mtype='marg')
             noisefile = noisedir + '/{0}_noise.txt'.format(psr)
             with open(noisefile, 'w') as f:

### 2017-08-02 codes/optstat.py
now handles exception for J1909 issues

@@ -142,7 +142,19 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
             # create chain object and find 1d maxLL parameter values
             cp = pp.ChainPP('chains/{0}/noise/{1}/'.format(dataset, psr))
-            ml = cp.get_ml_values(mtype='marg')
+            try:
+                ml = cp.get_ml_values(mtype='marg')
+            except:
+                # this sometimes arises for J1909
+
+                # 1d maximization for efac
+                ind = np.where(cp.pars == 'efac')
+                ml = cp.get_ml_values(mtype='marg', ind=ind[0]) 
+
+                # initialise the two RN parameters in ml object
+                ml.update({'RN-Amplitude': None})
+                ml.update({'RN-spectral-index': None})
+
             noisefile = noisedir + '/{0}_noise.txt'.format(psr)
             
             # load chain and find 2d maxLL values for RN parameters
@@ -150,7 +162,6 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
             chain = np.loadtxt('../data/chains/{0}/fix_spec_nf_30/chain_1.txt'.format(dataset))
             burn = int(0.25*chain.shape[0])
     
-            # 2-d histogram 
             RN_amplitude_chain = chain[:,np.argwhere(pars == 'RN-Amplitude_{0}'.format(psr))[0][0]]
             RN_spectral_index_chain = chain[:,np.argwhere(pars == 'RN-spectral-index_{0}'.format(psr))[0][0]]
             RN_Amplitude, RN_spectral_index = getMax2d(RN_amplitude_chain, RN_spectral_index_chain)
@@ -161,11 +172,11 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
                     if key == 'RN-Amplitude':
                         val = RN_Amplitude
-                        f.write('%s %g\n'%(key, val))
+                        f.write('%s %g\n'%(key, RN_Amplitude))
 
                     elif key == 'RN-spectral-index':
                         val = RN_spectral_index
-                        f.write('%s %g\n'%(key, val))
+                        f.write('%s %g\n'%(key, RN_spectral_index))
 
                     else:
                         f.write('%s %g\n'%(key, val))
@@ -173,7 +184,19 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
         if writenoise == '2dmax_indiv':
             # create chain object and find 1d maxLL parameter values
[…]

### 2017-08-02 manuscript/manuscript.tex
edits to introduction

@@
 
-Pulsar timing arrays (PTAs) are a conglomerate of pulsars with millisecond periods. 
-Monitoring the ultra-regular periodic emission from these pulsars with radio telescopes 
-enables us to probe the dynamics of the spacetime through which the pulses travel. 
-PTAs are sensitive to gravitational waves (GWs) with frequencies of 
-$10^{-9} - 10^{-7}\;\mathrm{Hz}$ \ki{citations needed}, 
-which makes PTAs ideal complements to ground- and space-based GW detectors \sv{add citations}. 
-\sv{something...} predict that the incoherent superposition of inspiraling SMBHBs form the dominant signal in this low-frequency regime \sv{add citations}.
+Low-frequency gravitational waves (GWs) with frequencies of 
+$10^{-9} - 10^{-7} \; \mathrm{Hz}$ can be observed with pulsar timing arrays (PTAs) 
+composed of millisecond pulsars (MSPs) \sv{citation}. 
+By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
+we can probe the dynamics of the spacetime through which the pulses travel. 
+The dominant astrophysical source in this frequency range is an isotropic stochastic background 
+made up of the incoherent superposition of GWs from inspiraling 
+supermassive black hole binaries (SMBHBs) 
+\citep{1995ApJ...446..543R, 2003ApJ...583..616J, 2003ApJ...590..691W}.
 

### 2017-08-03 manuscript/manuscript.tex
edits to sky scramble section

@@
 \begin{figure}[ht]
-	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
 	\caption{Optimal statistic and SNR for a simulated dataset 
@@
 \begin{figure}[ht]
-	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
 	\caption{Optimal statistic and SNR for 425 simulated datasets 
@@
 \begin{figure}[ht]
-	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
 	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
@@
 Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 
-for a particular data set to the 
-distribution of noise-marginalized optimal statistics obtained using sky scrambles.
-\sv{sky scrambles for all of our simulated data sets}
-
+for a particular data set to the distribution of noise-marginalized optimal statistics obtained using sky scrambles. 
+For this data set, the noise-marginalized mean optimal statistic is 
+$1.9\times10^{-29}$ and the mean SNR is 1.1. 
+The distributions of the noise-marginalized optimal statistic and SNR obtained using the scrambled ORFs 
+are both centered around zero, 
+and for 24 out of the 210 sky scrambles, the SNR is above the true value of $1.1$ ($p=0.11$). 
+Figure~\ref{fig:skyscrambles} shows the $p$-values for 210 sky scrambles 
+using 425 realizations of the stochastic background. 
+For 62\% of simulations the $p$-value is $<0.1$, and for 29\% of simulations the $p$-value is $<0.01$.
 \begin{figure}[ht]
-	\includegraphics[width=\columnwidth]{plots/skyscrambles_dataset50.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/skyscrambles_dataset50.pdf}
 	\caption{Comparison between the optimal statistic with and without sky scrambles 
@@
 \end{figure}
+\begin{figure}[ht]
+	\includegraphics[width=0.9\columnwidth]{plots/os_skyscrambles.pdf}
+	\caption{Distribution of $p$-values for 425 different realizations of the stochastic background. 
+			For each realization, the $p$-value is the fraction of sky scrambled ORFs that gives a 
+			noise-marginalized mean SNR greater than the mean SNR using the correct ORF.}
+	\label{fig:skyscrambles}
+\end{figure}
 
+For five realizations of the stochastic background, we also computed the Bayes' factors for the 
+true and sky scrambled ORFs in order to directly compare 
+the frequentist and Bayesian detection techniques.
 \sv{comparison to Steve's}

### 2017-08-04 manuscript/manuscript.tex
edits to text, especially Introduction

@@
 composed of millisecond pulsars (MSPs) \sv{citation}. 
-By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
-we can probe the dynamics of the spacetime through which the pulses travel. 
-The dominant astrophysical source in this frequency range is an isotropic stochastic background 
+The dominant astrophysical source in this frequency range is an isotropic stochastic 
+gravitational wave background (GWB) 
 made up of the incoherent superposition of GWs from inspiraling 
@@
 \citep{1995ApJ...446..543R, 2003ApJ...583..616J, 2003ApJ...590..691W}. 
+By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
+we can probe the dynamics of the spacetime through which the pulses travel. 
+This is done by computing the pulsar timing residuals, 
+which measure the difference between the expected and observed pulse times of arrival.
 
-Most measuring of the stochastic background is done through Bayesian data analysis.
-\ki{baye baye baye stuff} There are many advantages to running full Bayesian analyses on data sets, but they are slow.
-The bulk of the work consists of inverting non-diagonal matrices to find the strength of Hellings-Downs cross-correlations between pulsar pairs -- the "smoking gun" of a stochastic GWB.
+The primary data analysis technique used by PTAs is Bayesian inference 
+to compare the probabilities of various models for the pulsar timing residuals, 
+including one where the residuals contain evidence of an isotropic GW stochastic background. 
+Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated 
+models; however, they are computationally intensive, 
+particularly when searching for evidence of the spatial correlations characteristic of a 
+stochastic GWB -- ``the smoking gun'' of a stochastic GWB.
 
-Determining the significance of various cross-correlations between pulsar pairs is computationally expensive, requiring multi-core computing clusters many days to complete.
-As PTA data-analysis expands to include more parameters when measuring our predicted signal, e.g.
-increase in residuals with continued and new pulsar observations, uncertainty in solar system ephemerides, etc., computation time will only increase.
-Thus methods which hasten the production of results without compromising accuracy are not only preferable, but will be necessary.
-Here we explore the noise-marginalized optimal statistic as an estimator for GWB signals.
+The significance of a stochastic GWB can also be assessed using the 
+optimal statistic, a frequentist estimator for the GWB amplitude. 
+It is an important complement to Bayesian detection techniques 
+because it provides an independent detection procedure, 
+and it is significantly faster to compute. 
+In particular, different spatial correlations can be tested with the optimal statistic 
+in less than a second, 
+whereas a Bayesian analysis with spatial correlations requires 
+several weeks to run on a supercomputing cluster.
 
-The optimal statistic is a frequentist estimator for the stochastic background amplitude. 
-It was derived by \citet{abc+2009} \sv{blah blah blah... try not to plagiarize myself}
-
-The optimal statistic is an important complement to Bayesian detection techniques. 
-It provides an independent way to assess the significance of the stochastic background, 
-and it is significantly faster to compute. It is also easy to test different spatial correlations 
-by altering the overlap reduction function (ORF) $\Gamma$ in the cross-correlations. 
-In contrast, doing a full Bayesian analysis with spatial correlations is a very computationally 
-intensive procedure.
-
-However, previous work has found that the optimal statistic gives biased results \sv{expand...} 
-In order to compute the optimal statistic, the timing model parameters are fixed 
-at the maximum-likelihood values, but this leads to biases in the 
-optimal statistic value and uncertainty due to covariances between the 
-individual red noise parameters and the GW amplitude.
-
+However, previous work has found that the optimal statistic gives biased results 
+for PTAs with significant red noise in the individual pulsars due to the 
+strong covariance between the individual red noise parameters 
+and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
 In this paper, we present a technique for improving the accuracy of the optimal statistic 
@@
 estimate of \Agw\ and its uncertainty. 
+This method is more computationally intensive than previous methods 
+for computing the optimal statistic, but still requires only minutes. 
+Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
+can be used to compute the optimal statistic with any choice of the ORF, 
+allowing additional spatial correlations to be tested very quickly.
 
@@
 can assess the significance of a GW detection by performing sky scrambles \citep{cs2016}. 
-and compare to the Bayesian analysis done by \citet{tlb+2017}. 
+We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
 We conclude in Sec.~\ref{sec:conclusion}.
@@
 the PTA likelihood function in the weak-signal regime. 
-The likelihood function is given by
+We begin by finding the timing residuals for a given pulsar ${\bf{r}}_I$, which can be written as 
+\begin{equation}
+	{\bf{r}}_I = M {\bf{\epsilon}}_I + F {\bf{a}}_I + U {\bf{j}}_I + {\bf{n}}_I \,,\\
+\end{equation}
+where $M$ is the design matrix, 
+$\epsilon_I$ is a vector of the timing model parameter uncertainties, 
+${\bf{a}}_I$ are the red noise parameters describing both individual pulsars' red noise 
+and common red noise in all pulsars, 
+${\bf{j}}_I$ are white noise parameters describing frequency-correlated white noise, 
+and ${\bf{n}}_I$ are white noise parameters describing uncorrelated white noise. 
+The likelihood function is constructed from the residuals as given by
 \begin{equation}
@@
 \begin{eqnarray}
-	{\bf{P}} &=& \left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle \,, \\
+	{\bf{P}}_I &=& \left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle \,, \\
 	{\bf{S}}_{IJ} &=& \left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{I \neq J} \,.
 \end{eqnarray}
-\sv{something about the ORF...}
+In the frequency domain, these can be written as
+\begin{eqnarray}
+	\left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle_{ij} &=& {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_I(f) \right] {\bf{R}}_I^T \,, \\
+	\left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{ij} &=& \chi_{IJ} {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_J^T \,,
+\end{eqnarray}
+where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
+two observations, 
+$\mathbf{R}$ is the timing model fitting matrix, 
+$\chi_{IJ}$ is the overlap reduction function (ORF) that describes the spatial correlation between two pulsars' residuals, 
+$P_\mathrm{gw}(f)$ is the GW power spectrum, 
+and $P_I(f)$ is the power spectrum for the $I$th pulsar.
 
@@
 
-When constructing the optimal statistic, we typically fix the red noise parameters $\phi$ to the values that 
-maximize the likelihood, which we determine from a Bayesian common red noise analysis 
-\sv{check this part}. However, 
-doing this biases the optimal statistic and SNR \sv{blah blah blah} 
+When constructing the residuals, 
+we typically fix the red noise parameters to the values that 
+maximize the likelihood. 
+However, this leads to a bias in the optimal statistic and SNR because 
+the individual red noise parameters and the common red noise parameters 
+are highly covariant. 
+In this section, we compare three different techniques for computing the 
+optimal statistic for PTAs with significant red noise. 
+First, we fix the individual pulsars' red noise parameters to the 
+maximum-likelihood values from individual pulsar analyses. 
+Second, we fix the individual pulsars' red noise parameters to the 
+values maximum-likelihood values from a common analysis 
+involving all of the pulsars in our PTA. 
+For the third method, we
+
 In this paper we introduce a procedure to correct this bias by 
@@
 which consisted of 10 MSPs with \sv{Steve fill in here...} 
-
 The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
@@
 fix the white noise to the maximum likelihood value. 
-To compute the fixed-noise optimal statistic, 
-the red noise parameters for each pulsar are set to the mode of 
-the joint posterior distribution for $A_\mathrm{red}$,$\gamma_\mathrm{red}$, 
-found by performing a Bayesian common red noise analysis. 
-For the noise-marginalized optimal statistic, we draw from 
+For the noise-marginalized optimal statistic, we draw the red noise parameters from 
 the common red noise posteriors in order to average the optimal statistic over 
 the individual pulsars' red noise. 
+For comparison, we also compute the fixed-noise optimal statistic in two ways -- 
+first, by fixing the red noise parameters at the maximum-likelihood values from individual pulsar analyses; 
+second, by fixing the red noise parameters at the maximum-likelihood values from a common analysis.
 
@@
 
-Figure~\ref{fig:os_datasetstats} compares the fixed-noise values of $\hat{A}^2$ 
+In Fig.~\ref{fig:os_datasetstats} we compare the fixed-noise values of $\hat{A}^2$ 
 to the mean $\hat{A}^2$ for 425 different realizations of the stochastic background. 

### 2017-08-04 manuscript/manuscript.tex
edits to text, especially Introduction

@@
 we can probe the dynamics of the spacetime through which the pulses travel. 
-This is done by computing the pulsar timing residuals, 
+This is done by searching for correlations in the pulsar timing residuals, 
 which measure the difference between the expected and observed pulse times of arrival.
@@
 models; however, they are computationally intensive, 
-particularly when searching for evidence of the spatial correlations characteristic of a 
-stochastic GWB -- ``the smoking gun'' of a stochastic GWB.
+particularly when searching for evidence of dipole spatial correlations -- 
+``the smoking gun'' of a stochastic GWB.
 
@@
 involving all of the pulsars in our PTA. 
-For the third method, we
+For the third method, we draw values of the individual pulsars' 
+red noise parameters from the posteriors generated by a Bayesian analysis 
+that includes both individual red noise and a common red noise process, 
+allowing us to marginalize over the individual red noise.
 
-In this paper we introduce a procedure to correct this bias by 
-marginalizing over the red noise parameters using 
-posterior distributions from a Bayesian analysis.
-
-In order to compare the noise-marginalized optimal statistic to the 
-fixed-noise optimal statistic, we injected a GW stochastic background with 
-$\Agw = 5 \times 10^{-15}$ into simulated PTA data. 
-We constructed simulated data sets using the same procedure as in \citet{tlb+2017}, 
-which consisted of 10 MSPs with \sv{Steve fill in here...} 
+We used these three techniques to compute the optimal statistic for simulated data sets. 
+We constructed simulated data sets using the same procedure as in \citet{tlb+2017}. 
+The simulated PTAs consisted of 10 MSPs with \sv{Steve fill in here...}, 
+and the injected GWB amplitude was $\Agw = 5\times10^{-15}$. 
 The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}.
-We incorporate the individual pulsar noise parameters as follows. 
-We perform individual Bayesian noise analyses on all of the pulsars, and 
-fix the white noise to the maximum likelihood value. 
-For the noise-marginalized optimal statistic, we draw the red noise parameters from 
-the common red noise posteriors in order to average the optimal statistic over 
-the individual pulsars' red noise. 
-For comparison, we also compute the fixed-noise optimal statistic in two ways -- 
-first, by fixing the red noise parameters at the maximum-likelihood values from individual pulsar analyses; 
-second, by fixing the red noise parameters at the maximum-likelihood values from a common analysis.
 

### 2017-08-04 manuscript/manuscript.tex
edits to sky scramble section

@@
 We constructed simulated data sets using the same procedure as in \citet{tlb+2017}. 
-The simulated PTAs consisted of 10 MSPs with \sv{Steve fill in here...}, 
-and the injected GWB amplitude was $\Agw = 5\times10^{-15}$. 
+The simulated PTAs consisted of 10 MSPs with 
+total observation times and noise properties chosen to be 
+similar to the \sv{Steve fill in here...} (see Table~\ref{tab:sim}). 
+We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.
+\begin{table}[htb]
+	\setlength{\tabcolsep}{5pt}
+	\caption{Pulsar parameters used in simulated PTA}
+	\begin{center}
+	\begin{tabular}{ccccc}
+		\hline\hline
+    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $\log_{10}A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
+		\hline
+		J0613$-$0200 & 16.054 & 1.58 & -13.90 & 3.18 \\
+		J0751+1807 & 17.606 & 2.60 & -14.14 & 2.58 \\
+		J1012+5307 & 16.831 & 1.47 & -13.09 & 1.65 \\
+		J1640+2224 & 16.735 & 1.99 & -13.24 & 0.03 \\
+		J1643$-$1224 & 17.300 & 1.65 & -18.56 & 4.04 \\
+		J1713+0747 & 17.657 & 0.26 & -14.90 & 4.85 \\
+		J1744$-$1134 & 17.250 & 0.65 & -13.60 & 2.00 \\
+		J1857+0943 & 17.310 & 1.51 & -16.00 & 1.35 \\
+		J1909$-$3744 & 9.379 & 0.12 & -13.99 & 2.06 \\
+		J2145$-$0750 & 17.161 & 1.19 & -13.87 & 4.02 \\
+    		\hline\hline
+	\end{tabular}
+	\end{center}
+	\label{tab:sim}
+\end{table}
+
 The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
-
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
@@
 \begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
+	\includegraphics[width=0.95\columnwidth]{plots/os_dataset50.pdf}
 	\caption{Optimal statistic and SNR for a simulated dataset 
@@
 \begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
+	\includegraphics[width=0.95\columnwidth]{plots/os_datasetstats_5e-15.pdf}
 	\caption{Optimal statistic and SNR for 425 simulated datasets 
@@
 
-\section{Compare Significance with Monopole and Dipole}
+\section{Monopole and Dipole Spatial Correlations}
 \label{sec:spatial}
@@
 using the quadrupole ORF, and in 45\% and 24\% of our simulated data sets 
-using the monopole and dipole ORF, respectively.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
+using the monopole and dipole ORFs, respectively.
+\begin{figure}[htb]
+	\includegraphics[width=0.95\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
 	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
-			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using monopole (blue), dipole (orange), and quadrupole (green) 
-			spatial correlations. 
+			containing an injected isotropic stochastic GWB with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using monopole (blue), dipole (orange), 
+			and quadrupole (green) spatial correlations. 
 			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
@@
 
-\section{Compare Significance with Sky Scrambles/Phase Shifts}
+\section{Sky Scrambles}
 \label{sec:skyscrambles}
@@
 
-The scrambled sky positions were generated \sv{Steve, explain here...}
+We used the same scrambled sky positions as were used in \citet{tlb+2017}. 
+These positions were chosen so as to minimize the overlap  
+between the scrambled ORFs and the true ORF, 
+which can be assessed using the ``match statistic'' \citep{cs2016}:
+\begin{equation}
+	\bar{M} = \frac{\sum_{I,J \neq I} \Gamma_{IJ} \Gamma'_{IJ}}{\sqrt{ \left( \sum_{I, J \neq I} \Gamma_{IJ} \Gamma_{IJ} \right) \left( \sum_{I, J \neq I} \Gamma'_{IJ} \Gamma'_{IJ} \right)}} \,,
+\end{equation}
+where $I$,$J$ index the pulsar number, and $\Gamma$ and $\Gamma'$ are two different ORFs. 
+The 210 sky scrambles used in this paper were generated using a particle swarm optimization (PSO) algorithm 
+with the requirement that the scrambled ORFs had $\bar{M} < 0.1$ when compared to the true ORF and to each other.
+\sv{Steve, check this section}
 
@@
 For 62\% of simulations the $p$-value is $<0.1$, and for 29\% of simulations the $p$-value is $<0.01$.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/skyscrambles_dataset50.pdf}
-	\caption{Comparison between the optimal statistic with and without sky scrambles 
-			for a simulated dataset containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			The true analysis found an $\mathrm{SNR} = 1.1$, while only 24 of the 210 sky scramble analyses 
-			found $\mathrm{SNR} \geq 1.1$ ($p=0.11$).}
+\begin{figure}[htb]
+	\includegraphics[width=0.95\columnwidth]{plots/skyscrambles_dataset50.pdf}
+	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
+			with and without sky scrambles for a simulated dataset 
+			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+			Using the true ORF, we found an SNR of 1.1, while only 24 of the 210 analyses using scrambled ORFs 
+			found an SNR $\geq 1.1$ ($p=0.11$).}
 	\label{fig:skyscrambles_dataset_sample}
 \end{figure}
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_skyscrambles.pdf}
+\begin{figure}[htb]
+	\includegraphics[width=0.95\columnwidth]{plots/os_skyscrambles.pdf}
 	\caption{Distribution of $p$-values for 425 different realizations of the stochastic background. 
@@
 
+The noise-marginalized optimal statistic is a computationally inexpensive 
+method for assessing the significance of an isotropic stochastic GWB in 
+PTA data.
+
 \acknowledgments
 We thank Joe Romano and Xavier Siemens for useful discussions. 
+KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.
 JAE was supported by NASA through Einstein Fellowship Grant PF4-150120. 
@@
 and the Universities Space Research Association through a contract with NASA.
-KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.
 

### 2017-08-04 manuscript/manuscript.tex
added citations

@@
 where $I$,$J$ index the pulsar number, and $\Gamma$ and $\Gamma'$ are two different ORFs. 
-The 210 sky scrambles used in this paper were generated using a particle swarm optimization (PSO) algorithm 
+The 210 sky scrambles used in this paper were generated 
+using a particle swarm optimization (PSO) algorithm \citep{ke1995,se1998} 
 with the requirement that the scrambled ORFs had $\bar{M} < 0.1$ when compared to the true ORF and to each other.

### 2017-08-04 manuscript/manuscript.tex
minor edits

@@
 to compare the probabilities of various models for the pulsar timing residuals, 
-including one where the residuals contain evidence of an isotropic GW stochastic background. 
-Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated 
-models; however, they are computationally intensive, 
+including one where the residuals contain evidence of an isotropic GW stochastic background \sv{add citation}. 
+Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
+However, running a full Bayesian analysis is computationally intensive, 
 particularly when searching for evidence of dipole spatial correlations -- 
@@
 The significance of a stochastic GWB can also be assessed using the 
-optimal statistic, a frequentist estimator for the GWB amplitude. 
+optimal statistic, a frequentist estimator for the GWB amplitude \sv{add citation}. 
 It is an important complement to Bayesian detection techniques 
@@
 total observation times and noise properties chosen to be 
-similar to the \sv{Steve fill in here...} (see Table~\ref{tab:sim}). 
+similar to the properties of the EPTA (see Table~\ref{tab:sim})  \sv{add citation}. 
 We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.

### 2017-08-04 manuscript/manuscript.tex
added more references

@@
 \begin{abstract}
-Supermassive black hole binaries (SMBHBs) are predicted to form a high-redshift cosmological population of low-frequency gravitational wave sources which added together produce a stochastic gravitational wave background (GWB) potentially measurable via pulsar timing.
-Comparing the actual and predicted times-of-arrival (TOAs) of radio pulses beamed by millisecond pulsars allows us to search for and place limits on the stochastic background.
-While most estimating of GWB significance is done via Bayesian analysis, here we use a frequentist estimator called the noise-marginalized optimal statistic to assess the significance of a GWB present in simulated PTA data.
-We find \ki{complete}
+Supermassive black hole binaries (SMBHBs) are predicted to form a 
+high-redshift cosmological population of low-frequency gravitational wave sources 
+which added together produce a stochastic gravitational wave background (GWB) 
+that can be observed by pulsar timing arrays (PTAs). 
+Comparing the actual and predicted times-of-arrival (TOAs) of radio pulses beamed by 
+millisecond pulsars allows us to search for and place limits on the stochastic background. 
+While most estimating of GWB significance is done via Bayesian analysis, 
+here we use a frequentist estimator called the noise-marginalized optimal statistic 
+to assess the significance of a GWB present in simulated PTA data. We find \ki{complete}
 
@@
 $10^{-9} - 10^{-7} \; \mathrm{Hz}$ can be observed with pulsar timing arrays (PTAs) 
-composed of millisecond pulsars (MSPs) \sv{citation}. 
+composed of millisecond pulsars (MSPs) \cite{hd1983,fb1990}. 
 The dominant astrophysical source in this frequency range is an isotropic stochastic 
@@
 which measure the difference between the expected and observed pulse times of arrival. 
+Current upper-limits on the stochastic background from PTAs are approaching 
+theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,NANOGrav9yr}
 
@@
 to compare the probabilities of various models for the pulsar timing residuals, 
-including one where the residuals contain evidence of an isotropic GW stochastic background \sv{add citation}. 
+including one where the residuals contain evidence of an isotropic GW stochastic background 
+\citep{vlm+2009,lah+2013}. 
 Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
@@
 The significance of a stochastic GWB can also be assessed using the 
-optimal statistic, a frequentist estimator for the GWB amplitude \sv{add citation}. 
+optimal statistic, a frequentist estimator for the GWB amplitude \citep{abc+2009,demorest+2013,ccs+2015}. 
 It is an important complement to Bayesian detection techniques 
@@
 
+\sv{Justin, please bring notation up to date}
 As introduced by \citet{abc+2009}, 
@@
 total observation times and noise properties chosen to be 
-similar to the properties of the EPTA (see Table~\ref{tab:sim})  \sv{add citation}. 
+similar to the properties of the EPTA \cite{caballero+2016} (see Table~\ref{tab:sim}). 
 We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.

### 2017-08-05 codes/optstat.py
code now calculates OS for both types of fixed noise

@@ -78,7 +78,7 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
     h5filename = '../data/simulated_data/' + dataset + '/sim.hdf5'
 
-    # run the common red noise analysis
+    # # run the common red noise analysis
     print 'Running the common red noise analysis for the simulated dataset...'
     chaindir = '../data/chains/' + dataset
     print 'Noise chains saved in directory ' + chaindir
@@ -88,24 +88,18 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
               --noVaryNoise --mark9 --incRed --incGWB --Tspan 0 --noCorrelations --nf 30 \
               --fixSi 4.33'.format(h5filename, chaindir, niter))
     
-    # find sample with largest posterior value to be used for initialising os 
-#    print 'Retreiving maximum likelihood sample from ' + chaindir
-#    chain = np.loadtxt(chaindir + '/fix_spec_nf_30/chain_1.txt')
-#    burn = int(0.25*chain.shape[0])
-#    index = np.argmax(chain[burn:,-3])
-#    maxpost_sample = chain[index,:]
-#    np.save(chaindir + '/init_RN_params.npy', maxpost_sample)
 
     # run the individual noise analyses for each pulsar
     print 'Running the individual white noise analyses for each pulsar...'
+
     pfile = h5.File(h5filename)
     psrs = list(pfile.keys())
 
     for psr in psrs:
-        print '   Running on ' + psr
+        print psr
         os.system('PAL2_run.py --h5File {0} --pulsar {1} \
-                  --outDir {1}/chains/{3}/noise/{2}/ --niter {4} \
-                  --mark9 --incRed --nf 30'.format(h5filename, chaindir, psr, dataset, niter))
+                  --outDir ../data/chains/{2}/noise/{1}/ --niter {3} \
+                  --mark9 --incRed --nf 30'.format(h5filename, psr, dataset, niter))
 
     # make noise files for each pulsar
     noisedir = '../data/noisefiles/' + dataset
@@ -126,7 +120,7 @@ def run_noise_analysis(dataset, writenoise, niter=1000000):
 
         if writenoise == 'maxsample':
             # load chain and parameters names
-            pars = list(np.loadtxt('chains/{0}/noise/{1}/pars.txt'.format(dataset, psr), dtype='S42'))
+            pars = list(np.loadtxt('../data/chains/{0}/noise/{1}/pars.txt'.format(dataset, psr), dtype='S42'))
[…]

### 2017-08-06 codes/optstat.py
Changing model so incEphemMixture=False

@@ -281,7 +281,7 @@ def init_os_individual(dataset, h5file, psrlist, nf, noisedir=None, noVaryNoise=
     # make model dictionary, can play with number of frequencies here
     # Note: Tmax=0 means it will use the largest time span in the array to set the frequency bin size
     fullmodel = model.makeModelDict(incRedNoise=True, incEquad=incEquad, incJitterEquad=incJitterEquad,
-                                    likfunc='mark9', incGWB=True, nfreqs=nf, Tmax=0)
+                                    likfunc='mark9', incGWB=True, incEphemMixture=False, nfreqs=nf, Tmax=0)
 
     # loop through pulsars and fix white noise parameters to values in file
     if noisedir is not None:
@@ -334,7 +334,7 @@ def init_os_common(dataset, h5file, psrlist, nf, noisedir=None, noVaryNoise=Fals
     # make model dictionary, can play with number of frequencies here
     # Note: Tmax=0 means it will use the largest time span in the array to set the frequency bin size
     fullmodel = model.makeModelDict(incRedNoise=True, incEquad=incEquad, incJitterEquad=incJitterEquad,
-                                    likfunc='mark9', incGWB=True, nfreqs=nf, Tmax=0)
+                                    likfunc='mark9', incGWB=True, incEphemMixture=False, nfreqs=nf, Tmax=0)
 
     # loop through pulsars and fix white noise parameters to values in file
     if noisedir is not None:

### 2017-08-08 codes/combine_data.py
Script that collates results from the two fixed noise methods

@@ -1,0 +1,35 @@
+# Collates optimal statistic data into one file containing 
+# both types of fixed noise and noise-marginalized results 
+
+
+import numpy as np
+import os,sys,glob,math
+
+datadir = '../data/optstat/'
+datasetname = 'dataset'
+ndatasets = 100
+
+outputfilename = datadir + 'stats_all.dat'
+
+f = open(outputfilename, 'w')
+
+for n in range(ndatasets):
+    name = '../data/optstat_individual/' + dataset + str(n)
+    if os.path.exists(name):
+        data_init_individual = np.loadtxt(name + '/init_os.dat', usecols=1)
+        data_individual = np.loadtxt(name + '/marg_os.dat')
+    name = '../data/optstat_common/' + dataset + str(n)
+    if os.path.exists(name):
+        data_init_common = np.loadtxt(name + '/init_os.dat', usecols=1)
+        data_common = np.loadtxt(name + '/marg_os.dat')
+
+        f.write('{0:<10}  {1:>10.3e}  {2:>10.3e}  {3:>6.3f}  {4:>10.3e}  {5:>10.3e}  {6:>6.3f} {7:>10.3e}  {8:>10.3e}  {9:>6.3f}  {10:>10.3e}  {11:>10.3e}  {12:>6.3f}\n'.format(datasetname+str(n),
+                                                             data_init_individual[0], data_init_individual[1], data_init_individual[2],
+                                                             np.mean(data_individual[:,0]),
+                                                             np.mean(data_individual[:,1]),
+                                                             np.mean(data_individual[:,2]), 
+                                                             data_init_common[0], data_init_common[1], data_init_common[2],
+                                                             np.mean(data_common[:,0]),
+                                                             np.mean(data_common[:,1]),
+                                                             np.mean(data_common[:,2])))
+f.close()

### 2017-08-08 manuscript/manuscript.tex
added CiteNeed command

@@
 \usepackage[usenames]{color}
+\newcommand{\CiteNeed}{\textcolor{red}{\scriptsize[citation needed]}}
 \newcommand{\sv}[1]{\textcolor{blue}{\it{\textbf{sv: #1}}} }
@@
 
-The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
+The optimal statistic was computed using the \sv{fill in...} in PAL2 \CiteNeed. 
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
@@
 \begin{figure}[ht]
-	\includegraphics[width=0.95\columnwidth]{plots/os_dataset50.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
 	\caption{Optimal statistic and SNR for a simulated dataset 
@@
 \begin{figure}[ht]
-	\includegraphics[width=0.95\columnwidth]{plots/os_datasetstats_5e-15.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
 	\caption{Optimal statistic and SNR for 425 simulated datasets 
@@
 \begin{figure}[htb]
-	\includegraphics[width=0.95\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
 	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
@@
 \begin{figure}[htb]
-	\includegraphics[width=0.95\columnwidth]{plots/skyscrambles_dataset50.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/skyscrambles_dataset50.pdf}
 	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
@@
 \begin{figure}[htb]
-	\includegraphics[width=0.95\columnwidth]{plots/os_skyscrambles.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/os_skyscrambles.pdf}
 	\caption{Distribution of $p$-values for 425 different realizations of the stochastic background. 

### 2017-08-11 codes/makesims_nano.py
Added new nanograv-like data simulator

@@ -1,0 +1,69 @@
+import numpy as np
+import os,sys,glob,math
+import json
+
+import libstempo.toasim as LT
+
+def create_dataset(dataset, Agwb):
+    """ Create simulated dataset using 18 pulsars from NANOGrav 9-year
+    stochastic analysis. Will use 11-year data span and red noise values
+    with white noise values taken from most recent time-to-detection
+    simulations.
+
+    :param dataset: Name of output dataset.
+    :param Agwb: Amplitude of injected GWB
+    """
+
+    print 'Getting pulsar parameters for simulated dataset...'
+    # get simulation data
+    with open('nano9_simdata.json', 'r') as fp:
+        pdict = json.load(fp)
+
+    # get red noise dictionary
+    with open('nano_red_dict.json', 'r') as fp:
+        red_dict = json.load(fp)
+
+    # get parfiles
+    parfiles = glob.glob('../data/nano9_stipped_parfiles/*.par')
+    parfiles = [p for p in parfiles if p.split('/')[-1].split('_')[0]
+                in pdict.keys()]
+
+    datadir = '../data/simulated_data/' + dataset
+    if not os.path.exists(datadir):
+        os.makedirs(datadir)
+    print 'Making simulated data in directory ' + datadir
+
+    psrs = []
+    for pf in parfiles:
+        pname = pf.split('/')[-1].split('_')[0]
+        psrs.append(LT.fakepulsar(pf, pdict[pname][0], pdict[pname][1]))
+
+    for psr in psrs:
+        # white noise
+        LT.add_efac(psr)
+
+        # red noise
+        if pname in red_dict:
+            LT.add_rednoise(psr, red_dict[pname][0], red_dict[pname][1],
+                            components=30)
+
+    # GWB
+    LT.createGWB(psrs, Agwb, 13./3., seed=None)
+
+    for psr in psrs:
+        psr.fit(iters=2)
+
+        ## Save par and tim files
+        psr.savepar(datadir+'/{0}_optstatsim.par'.format(psr.name))
+        psr.savetim(datadir+'/{0}_optstatsim.tim'.format(psr.name))
+
+    # make the hdf5 file for the simulated dataset
+    print 'Making the hdf5 file for the simulated dataset...'
+    datadir = '../data/simulated_data/' + dataset
+    h5filename = datadir + '/sim.hdf5'
+
+    os.system('python makeH5file.py \
+              --pardir {0} --timdir {0} \
+              --h5File {1}'.format(datadir, h5filename));
+
+    print 'Finished!'

### 2017-08-11 .ci/build_paper.sh
Trying to set up automated paper generation with travis

@@ -1,0 +1,28 @@
+#!/bin/bash -x
+
+if git diff --name-only $TRAVIS_COMMIT_RANGE | grep 'manuscript/'
+then
+  # Install tectonic using conda
+  wget https://repo.continuum.io/miniconda/Miniconda3-latest-Linux-x86_64.sh -O miniconda.sh;
+  bash miniconda.sh -b -p $HOME/miniconda
+  export PATH="$HOME/miniconda/bin:$PATH"
+  hash -r
+  conda config --set always_yes yes --set changeps1 no
+  conda update -q conda
+  conda info -a
+  conda create --yes -n manuscript
+  source activate manuscript
+  conda install -c conda-forge -c pkgw-forge tectonic
+
+  # Build the paper using tectonic
+  cd manuscript
+  tectonic manuscript.tex --print
+
+  # Force push the paper to GitHub
+  cd $TRAVIS_BUILD_DIR
+  git checkout --orphan $TRAVIS_BRANCH-pdf
+  git rm -rf .
+  git add -f manuscript/manuscript.pdf
+  git -c user.name='travis' -c user.email='travis' commit -m "building the paper"
+  git push -q -f https://$GITHUB_USER:$GITHUB_API_KEY@github.com/$TRAVIS_REPO_SLUG $TRAVIS_BRANCH-pdf
+fi

### 2017-12-20 manuscript/manuscript.tex
Trying to set up automated paper generation with travis
Trying travis
Editing wording in introduction

@@
 
+Test travis...
+
 \end{abstract}
@@
 
-Low-frequency gravitational waves (GWs) with frequencies of 
+Long-wavelength gravitational waves (GWs) with frequencies of 
 $10^{-9} - 10^{-7} \; \mathrm{Hz}$ can be observed with pulsar timing arrays (PTAs) 
@@
 However, running a full Bayesian analysis is computationally intensive, 
-particularly when searching for evidence of dipole spatial correlations -- 
+particularly when searching for evidence of dipole spatial correlations:
 ``the smoking gun'' of a stochastic GWB.
@@
 optimal statistic, a frequentist estimator for the GWB amplitude \citep{abc+2009,demorest+2013,ccs+2015}. 
-It is an important complement to Bayesian detection techniques 
-because it provides an independent detection procedure, 
-and it is significantly faster to compute. 
-In particular, different spatial correlations can be tested with the optimal statistic 
-in less than a second, 
-whereas a Bayesian analysis with spatial correlations requires 
-several weeks to run on a supercomputing cluster.
+%It is an important complement to Bayesian detection techniques 
+%because it provides an independent detection procedure, 
+%and it is significantly faster to compute. 
+%In particular, different spatial correlations can be tested with the optimal statistic 
+%in less than a second, 
+%whereas a Bayesian analysis with spatial correlations requires 
+%several weeks to run on a supercomputing cluster.
+Not only does it provide an independent detection procedure, complementing a more robust Bayesian analysis, but its calculation requires significantly less time to compute.
+In particular, the optimal statistic produces results for various spatial correlation functions, a bottleneck within the more traditional stochastic GWB analysis, within seconds; a full Bayesian analysis including correlations has to run for several weeks on a supercomputing cluster.
 
@@
 and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
-In this paper, we present a technique for improving the accuracy of the optimal statistic 
-by marginalizing the optimal statistic over the individual red noise parameters 
-using the posterior distributions from a full Bayesian analysis. 
-This hybrid approach produces a more accurate 
-estimate of \Agw\ and its uncertainty. 
-This method is more computationally intensive than previous methods 
-for computing the optimal statistic, but still requires only minutes. 
-Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
-can be used to compute the optimal statistic with any choice of the ORF, 
-allowing additional spatial correlations to be tested very quickly.
+%In this paper, we present a technique for improving the accuracy of the optimal statistic 
+%by marginalizing the optimal statistic over the individual red noise parameters 
+%using the posterior distributions from a full Bayesian analysis. 
+%This hybrid approach produces a more accurate 
+%estimate of \Agw\ and its uncertainty. 
+Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: marginalizing over the individual pulsar red noise parameters using the posterior distributions from a full PTA Bayesian analysis.
+This hybrid approach produces a more precise estimate of the stochastic GWB amplitude,
+\Agw\/, and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
+%This method is more computationally intensive than previous methods 
+%for computing the optimal statistic, but still requires only minutes. 
+%Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
+%can be used to compute the optimal statistic with any choice of the ORF, 
+%allowing additional spatial correlations to be tested very quickly.
+Furthermore, the same Bayesian analysis drawn upon by the noise marginalization can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), allowing for quick testing of various spatial correlations.
 

### 2018-03-18 codes/convert_finished_to_fits.py
added function to convert chain_1.txt files to chain.fits files

@@ -1,0 +1,49 @@
+import numpy as np
+import os.path
+import glob
+from astropy.io import fits
+from astropy.table import Table
+
+
+def chain_txt_to_fits(chaindir):
+    
+    txtfile = '{0}/chain_1.txt'.format(chaindir)
+    parfile = '{0}/pars.txt'.format(chaindir)
+    
+    chain = np.loadtxt(txtfile)
+    pars = np.loadtxt(parfile, dtype='S42')
+    
+    # Append pars with MCMC information, these are always the 
+    # last four entries in a chain sample
+    super_pars = np.array(['lnprob', 'lnlike', 'acceptance rate', 'parallel-tempering acceptance rate'])
+
+    pars = np.concatenate((pars, super_pars))
+    
+    # initiate array of of columns
+    c = [[]] * len(pars)
+
+    # fill in columns with chain data
+    for ii, par in enumerate(pars):
+        c[ii] = fits.Column(name=par, format='D', array=chain[:,ii]) # format is set to float64: 'D'
+
+    # convert collection of columns to Bin HDU table
+    tbhdu = fits.BinTableHDU.from_columns(c)
+
+    # save as fits file
+    print 'Writing fits file {0}/chain.fits'.format(chaindir)
+    tbhdu.writeto('{0}/chain.fits'.format(chaindir))
+
+    print 'Removing chain file ' + txtfile
+    os.remove(txtfile)
+    
+
+if __name__ == '__main__':
+
+    import argparse
+
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--chaindir')
+
+    args = parser.parse_args()
+
+    chain_txt_to_fits(chaindir=args.chaindir)

### 2018-03-19 codes/optstat_v3.py
added updated version of optstat code

@@ -1,0 +1,711 @@
+import numpy as np
+import os,sys,glob,math
+import h5py as h5
+import scipy.interpolate as interp
+import scipy.ndimage.filters as filter
+
+from PAL2 import PALmodels, PALutils, pputils as pp
+
+import makesims_nano
+
+
+def get_ml_vals(cp):
+
+    ind = np.arange(0, cp.ndim)
+    ind = np.atleast_1d(ind)
+
+    x = OrderedDict()
+
+    for i in ind:
+        x[cp.pars[i]] = bu.getMax(cp.chain[:,i])
+
+    return x
+
+
+def getMax2d(samples1, samples2, weights=None, smooth=True, bins=[40, 40],
+            x_range=None, y_range=None, logx=False, logy=False, logz=False):
+    
+    if x_range is None:
+        xmin = np.min(samples1)
+        xmax = np.max(samples1)
+    else:
+        xmin = x_range[0]
+        xmax = x_range[1]
+
+    if y_range is None:
+        ymin = np.min(samples2)
+        ymax = np.max(samples2)
+    else:
+        ymin = y_range[0]
+        ymax = y_range[1]
+
+    if logx:
+        bins[0] = np.logspace(np.log10(xmin), np.log10(xmax), bins[0])
+    
+    if logy:
+        bins[1] = np.logspace(np.log10(ymin), np.log10(ymax), bins[1])
+
+    hist2d,xedges,yedges = np.histogram2d(samples1, samples2, weights=weights, \
+            bins=bins,range=[[xmin,xmax],[ymin,ymax]])
+    extent = [xedges[0], xedges[-1], yedges[0], yedges[-1] ]
+
+    if logz:
+        for ii in range(hist2d.shape[0]):
+            for jj in range(hist2d.shape[1]):
+                if hist2d[ii,jj] <= 0:
+                    hist2d[ii,jj] = 1
+
+    
+    xedges = np.delete(xedges, -1) + 0.5*(xedges[1] - xedges[0])
+    yedges = np.delete(yedges, -1) + 0.5*(yedges[1] - yedges[0])
+    
+    # gaussian smoothing
+    if smooth:
+        hist2d = filter.gaussian_filter(hist2d, sigma=0.75)
+
+    # interpolation
+    f = interp.interp2d(xedges, yedges, hist2d, kind='cubic')
+    xedges = np.linspace(xedges.min(), xedges.max(), 10000)
[…]

### 2018-03-19 manuscript/manuscript.tex
updates to manuscript

@@
 \DeclareMathOperator{\Tr}{Tr}
+
+\usepackage{natbib}
+\usepackage{amsmath,amssymb}
 
@@
 can assess the significance of a GW detection by performing sky scrambles \citep{cs2016}. 
-We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
+%We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
 We conclude in Sec.~\ref{sec:conclusion}.
@@
 
-\sv{Justin, please bring notation up to date}
+\sv{bring notation up to date}
 As introduced by \citet{abc+2009}, 
@@
 Second, we fix the individual pulsars' red noise parameters to the 
-values maximum-likelihood values from a common analysis 
-involving all of the pulsars in our PTA. 
+values that jointly maximize the likelihood for a common analysis 
+of all of the pulsars in our PTA 
+that searches over the pulsars' red noise parameters and 
+a common red process. 
 For the third method, we draw values of the individual pulsars' 
@@
 We used these three techniques to compute the optimal statistic for simulated data sets. 
-We constructed simulated data sets using the same procedure as in \citet{tlb+2017}. 
-The simulated PTAs consisted of 10 MSPs with 
-total observation times and noise properties chosen to be 
-similar to the properties of the EPTA \cite{caballero+2016} (see Table~\ref{tab:sim}). 
-We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.
+We simulated PTA data sets with comprised of 18 MSPs 
+with the same sky locations, observation times, and noise properties 
+as the MSPs in the NANOGrav 9-year data set 
+with at least 3 years of observation time \citep{abb+15} 
+(see Table~\ref{tab:sim} for more details). 
+To these data sets we injected an isotropic stochastic GWB 
+\sv{insert strength of GWB... are we going to have one value or more than one?}.
 \begin{table}[htb]
 	\setlength{\tabcolsep}{5pt}
-	\caption{Pulsar parameters used in simulated PTA}
+	\caption{Pulsar parameters used in simulated PTAs}
 	\begin{center}
@@
 		\hline
+		J0030+0451 & & & \\
 		J0613$-$0200 & 16.054 & 1.58 & -13.90 & 3.18 \\
-		J0751+1807 & 17.606 & 2.60 & -14.14 & 2.58 \\
 		J1012+5307 & 16.831 & 1.47 & -13.09 & 1.65 \\
+		J1024$-$0719 & & & \\
+		J1455$-$3330 & & & \\
+		J1600$-$3053 & & & \\
+		J1614$-$2230 & & & \\
 		J1640+2224 & 16.735 & 1.99 & -13.24 & 0.03 \\
-		J1643$-$1224 & 17.300 & 1.65 & -18.56 & 4.04 \\
 		J1713+0747 & 17.657 & 0.26 & -14.90 & 4.85 \\
+		J1741+1351 & & & \\
 		J1744$-$1134 & 17.250 & 0.65 & -13.60 & 2.00 \\
-		J1857+0943 & 17.310 & 1.51 & -16.00 & 1.35 \\
+		B1855+09 & & & \\
+		J1853+1303 & & & \\
 		J1909$-$3744 & 9.379 & 0.12 & -13.99 & 2.06 \\
+		J1918$-$0642 & & & \\
+		J2010$-$1323 & & & \\
 		J2145$-$0750 & 17.161 & 1.19 & -13.87 & 4.02 \\
+		J2317+1439 & & & \\
     		\hline\hline
@@
 
-The optimal statistic was computed using the \sv{fill in...} in PAL2 \CiteNeed. 
+%We constructed simulated data sets using the same procedure as in \citet{tlb+2017}. 
+%The simulated PTAs consisted of 10 MSPs with 
+%total observation times and noise properties chosen to be 
+%similar to the properties of the EPTA \cite{caballero+2016} (see Table~\ref{tab:sim}). 
+%We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.
+%\begin{table}[htb]
+%	\setlength{\tabcolsep}{5pt}
+%	\caption{Pulsar parameters used in simulated PTA}
+%	\begin{center}
+%	\begin{tabular}{ccccc}
+%		\hline\hline
+%    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $\log_{10}A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
+%		\hline
+%		J0613$-$0200 & 16.054 & 1.58 & -13.90 & 3.18 \\
+%		J0751+1807 & 17.606 & 2.60 & -14.14 & 2.58 \\
+%		J1012+5307 & 16.831 & 1.47 & -13.09 & 1.65 \\
+%		J1640+2224 & 16.735 & 1.99 & -13.24 & 0.03 \\
+%		J1643$-$1224 & 17.300 & 1.65 & -18.56 & 4.04 \\
+%		J1713+0747 & 17.657 & 0.26 & -14.90 & 4.85 \\
+%		J1744$-$1134 & 17.250 & 0.65 & -13.60 & 2.00 \\
+%		J1857+0943 & 17.310 & 1.51 & -16.00 & 1.35 \\
+%		J1909$-$3744 & 9.379 & 0.12 & -13.99 & 2.06 \\
+%		J2145$-$0750 & 17.161 & 1.19 & -13.87 & 4.02 \\
+%    		\hline\hline
+%	\end{tabular}
+%	\end{center}
+%	\label{tab:sim}
+%\end{table}
+
+The optimal statistic was computed using the PTA data analysis package 
+PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
+\citep{evh17a}. 
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 

### 2018-03-19 manuscript/manuscript.tex
updated affiliations

@@
 \affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}
+\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, California 91125, USA}
 
 \author{Justin A.\ Ellis}
-\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}
+\affiliation{Department of Physics and Astronomy, West Virginia University, P.O. Box 6315, Morgantown, WV 26506, USA}
+\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505, USA}
 

### 2018-03-19 manuscript/manuscript.tex
commented out figures as reminder that they all need to be redone

@@
 The noise-marginalized analysis also gives a larger mean SNR compared to the fixed-noise analysis.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
-	\caption{Optimal statistic and SNR for a simulated dataset 
-			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
-			mean optimal statistic and mean SNR found by marginalizing over 
-			the individual pulsars' red noise parameters (solid orange line). 
-			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
-			for a common red process (solid green line). 
-			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			For this particular realization of the stochastic background, the fixed-noise analysis 
-			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
-			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
-			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
-	\label{fig:os_dataset_sample}
-\end{figure}
+%\begin{figure}[ht]
+%	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
+%	\caption{Optimal statistic and SNR for a simulated dataset 
+%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+%			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
+%			mean optimal statistic and mean SNR found by marginalizing over 
+%			the individual pulsars' red noise parameters (solid orange line). 
+%			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
+%			for a common red process (solid green line). 
+%			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+%			For this particular realization of the stochastic background, the fixed-noise analysis 
+%			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
+%			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
+%			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
+%	\label{fig:os_dataset_sample}
+%\end{figure}
 
@@
 compared to an average SNR of 1.0 using the fixed-noise analysis.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
-	\caption{Optimal statistic and SNR for 425 simulated datasets 
-			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using a fixed-noise analysis (blue) to the 
-			mean optimal statistic and mean SNR found by marginalizing over 
-			the individual pulsars' red noise parameters. The dashed vertical line 
-			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			The value of $\hat{A}^2$ from the fixed-noise analysis, 
-			averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
-			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			Furthermore, the fixed-noise analysis gives an average SNR of 1.0, while the 
-			noise-marginalized analysis gives an average mean SNR of 1.7.}
-	\label{fig:os_datasetstats}
-\end{figure}
+%\begin{figure}[ht]
+%	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
+%	\caption{Optimal statistic and SNR for 425 simulated datasets 
+%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+%			We compare the values found using a fixed-noise analysis (blue) to the 
+%			mean optimal statistic and mean SNR found by marginalizing over 
+%			the individual pulsars' red noise parameters. The dashed vertical line 
+%			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+%			The value of $\hat{A}^2$ from the fixed-noise analysis, 
+%			averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
+%			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+%			Furthermore, the fixed-noise analysis gives an average SNR of 1.0, while the 
+%			noise-marginalized analysis gives an average mean SNR of 1.7.}
+%	\label{fig:os_datasetstats}
+%\end{figure}
 
@@
 using the monopole and dipole ORFs, respectively.
-\begin{figure}[htb]
-	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
-	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
-			containing an injected isotropic stochastic GWB with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using monopole (blue), dipole (orange), 
-			and quadrupole (green) spatial correlations. 
-			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
-	\label{fig:os_ORF}
-\end{figure}
+%\begin{figure}[htb]
+%	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
+%	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
+%			containing an injected isotropic stochastic GWB with $\Agw = 5\times10^{-15}$. 
+%			We compare the values found using monopole (blue), dipole (orange), 
+%			and quadrupole (green) spatial correlations. 
+%			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
+%	\label{fig:os_ORF}
+%\end{figure}
 
@@
 For 62\% of simulations the $p$-value is $<0.1$, and for 29\% of simulations the $p$-value is $<0.01$.
-\begin{figure}[htb]
-	\includegraphics[width=0.9\columnwidth]{plots/skyscrambles_dataset50.pdf}
-	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
-			with and without sky scrambles for a simulated dataset 
-			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-			Using the true ORF, we found an SNR of 1.1, while only 24 of the 210 analyses using scrambled ORFs 
-			found an SNR $\geq 1.1$ ($p=0.11$).}
-	\label{fig:skyscrambles_dataset_sample}
-\end{figure}
-\begin{figure}[htb]
-	\includegraphics[width=0.9\columnwidth]{plots/os_skyscrambles.pdf}
-	\caption{Distribution of $p$-values for 425 different realizations of the stochastic background. 
-			For each realization, the $p$-value is the fraction of sky scrambled ORFs that gives a 
-			noise-marginalized mean SNR greater than the mean SNR using the correct ORF.}
-	\label{fig:skyscrambles}
-\end{figure}
+%\begin{figure}[htb]
+%	\includegraphics[width=0.9\columnwidth]{plots/skyscrambles_dataset50.pdf}
+%	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
+%			with and without sky scrambles for a simulated dataset 
+%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
+%			Using the true ORF, we found an SNR of 1.1, while only 24 of the 210 analyses using scrambled ORFs 
+%			found an SNR $\geq 1.1$ ($p=0.11$).}
+%	\label{fig:skyscrambles_dataset_sample}
+%\end{figure}
+%\begin{figure}[htb]
+%	\includegraphics[width=0.9\columnwidth]{plots/os_skyscrambles.pdf}
+%	\caption{Distribution of $p$-values for 425 different realizations of the stochastic background. 
+%			For each realization, the $p$-value is the fraction of sky scrambled ORFs that gives a 
+%			noise-marginalized mean SNR greater than the mean SNR using the correct ORF.}
+%	\label{fig:skyscrambles}
+%\end{figure}
 

### 2018-03-19 manuscript/manuscript.tex
updated abstract

@@
 Supermassive black hole binaries (SMBHBs) are predicted to form a 
-high-redshift cosmological population of low-frequency gravitational wave sources 
-which added together produce a stochastic gravitational wave background (GWB) 
-that can be observed by pulsar timing arrays (PTAs). 
-Comparing the actual and predicted times-of-arrival (TOAs) of radio pulses beamed by 
-millisecond pulsars allows us to search for and place limits on the stochastic background. 
-While most estimating of GWB significance is done via Bayesian analysis, 
-here we use a frequentist estimator called the noise-marginalized optimal statistic 
-to assess the significance of a GWB present in simulated PTA data. We find \ki{complete}
-
-Test travis...
+high-redshift cosmological population of low-frequency gravitational wave (GW) sources, 
+which added together produce a stochastic gravitational wave background (GWB). 
+Pulsar timing arrays (PTAs) can observe this signal by looking for correlations in the 
+times of arrival (TOAs) of millisecond pulsars (MSPs). 
+In this paper, we introduce an improved frequentist estimator called the noise-marginalized optimal statistic, 
+which can be used to assess the significance of a GWB present in PTA data. 
+We find \ki{complete}
 

### 2018-03-19 codes/optstat_v3.py
fixes to code

@@ -602,6 +602,9 @@ if __name__ == '__main__':
     dataset = args.datasetname
     writenoise = args.writenoise
     
+    nreal = int(args.nreal)
+    nscrambles = int(args.nscrambles)
+    
     if not args.skipdatacreate:
         makesims_nano.create_dataset(dataset, Agwb)
     
@@ -634,6 +637,7 @@ if __name__ == '__main__':
         os.makedirs(outputdir)
 
     chaindir = '../data/chains/' + dataset + '/fix_spec_nf_30/'
+    chain = np.loadtxt(chaindir + 'chain_1.txt')
     burn = int(0.25*chain.shape[0])
 
     index = np.argmax(chain[:,-3])
@@ -664,7 +668,7 @@ if __name__ == '__main__':
         sig_ss = [ np.zeros(nreal) for i in range(nscrambles) ]
 
     for ii in range(nreal):
-        preal = chain[np.random.randint(0, chain.shape[0]), :]
+        preal = chain[np.random.randint(burn, chain.shape[0]), :-4]
         preal = np.concatenate((preal, np.array([4.33])))
         xi, rho, sig, opts[ii], sigs[ii] = model.opt_stat_mark9(preal, fixWhite=True)
         if args.computeMonopole:
@@ -672,7 +676,7 @@ if __name__ == '__main__':
         if args.computeDipole:
             opts2[ii], sigs2[ii] = compute_os(xi2, rho, sig)
         if args.computeSkyScrambles:
-            for j in range(args.nscrambles):
+            for j in range(nscrambles):
                 opt_ss[j][ii], sig_ss[j][ii] = compute_os(xi_ss[j], rho, sig)
         if ii > 1 and ii % 100 == 0:
             sys.stdout.write('\r')

### 2018-03-19 codes/optstat_v2.py
removed old version of code

@@ -1,631 +1,0 @@
-import numpy as np
-import os,sys,glob,math
-import h5py as h5
-import scipy.interpolate as interp
-import scipy.ndimage.filters as filter
-
-from PAL2 import PALmodels, PALutils, pputils as pp
-
-import makesims_nano
-
-
-def get_ml_vals(cp):
-
-    ind = np.arange(0, cp.ndim)
-    ind = np.atleast_1d(ind)
-
-    x = OrderedDict()
-
-    for i in ind:
-        x[cp.pars[i]] = bu.getMax(cp.chain[:,i])
-
-    return x
-
-
-def getMax2d(samples1, samples2, weights=None, smooth=True, bins=[40, 40],
-            x_range=None, y_range=None, logx=False, logy=False, logz=False):
-    
-    if x_range is None:
-        xmin = np.min(samples1)
-        xmax = np.max(samples1)
-    else:
-        xmin = x_range[0]
-        xmax = x_range[1]
-
-    if y_range is None:
-        ymin = np.min(samples2)
-        ymax = np.max(samples2)
-    else:
-        ymin = y_range[0]
-        ymax = y_range[1]
-
-    if logx:
-        bins[0] = np.logspace(np.log10(xmin), np.log10(xmax), bins[0])
-    
-    if logy:
-        bins[1] = np.logspace(np.log10(ymin), np.log10(ymax), bins[1])
-
-    hist2d,xedges,yedges = np.histogram2d(samples1, samples2, weights=weights, \
-            bins=bins,range=[[xmin,xmax],[ymin,ymax]])
-    extent = [xedges[0], xedges[-1], yedges[0], yedges[-1] ]
-
-    if logz:
-        for ii in range(hist2d.shape[0]):
-            for jj in range(hist2d.shape[1]):
-                if hist2d[ii,jj] <= 0:
-                    hist2d[ii,jj] = 1
-
-    
-    xedges = np.delete(xedges, -1) + 0.5*(xedges[1] - xedges[0])
-    yedges = np.delete(yedges, -1) + 0.5*(yedges[1] - yedges[0])
-    
-    # gaussian smoothing
-    if smooth:
-        hist2d = filter.gaussian_filter(hist2d, sigma=0.75)
-
-    # interpolation
-    f = interp.interp2d(xedges, yedges, hist2d, kind='cubic')
-    xedges = np.linspace(xedges.min(), xedges.max(), 10000)
[…]

### 2018-03-19 manuscript/manuscript.tex
updated description of simulated data

@@
 
-We used these three techniques to compute the optimal statistic for simulated data sets. 
-We simulated PTA data sets with comprised of 18 MSPs 
-with the same sky locations, observation times, and noise properties 
-as the MSPs in the NANOGrav 9-year data set 
-with at least 3 years of observation time \citep{abb+15} 
+We used these three techniques to compute the optimal statistic 
+for simulated ``NANOGrav-like'' data sets. 
+These consisted of 18 MSPs with observation times up to 11 years 
+and sky positions matching the 18 longest-observed pulsars in the 
+NANOGrav 11-year data set \citep{abb+17}.
+with the observation times increased so that the longest-observed pulsars 
+had TOAs spanning over 11 years. 
+We include white noise for all pulsars, plus 
+red noise parametrized as a power law, $P(f) = A_\mathrm{red} f^{-\gamma}$, 
+for those pulsars that show evidence of red noise 
 (see Table~\ref{tab:sim} for more details). 
@@
 	\setlength{\tabcolsep}{5pt}
-	\caption{Pulsar parameters used in simulated PTAs}
+	\caption{Pulsar parameters used in simulated PTAs.}
 	\begin{center}
@@
 		\hline\hline
-    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $\log_{10}A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
+    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
 		\hline
-		J0030+0451 & & & \\
-		J0613$-$0200 & 16.054 & 1.58 & -13.90 & 3.18 \\
-		J1012+5307 & 16.831 & 1.47 & -13.09 & 1.65 \\
-		J1024$-$0719 & & & \\
-		J1455$-$3330 & & & \\
-		J1600$-$3053 & & & \\
-		J1614$-$2230 & & & \\
-		J1640+2224 & 16.735 & 1.99 & -13.24 & 0.03 \\
-		J1713+0747 & 17.657 & 0.26 & -14.90 & 4.85 \\
-		J1741+1351 & & & \\
-		J1744$-$1134 & 17.250 & 0.65 & -13.60 & 2.00 \\
-		B1855+09 & & & \\
-		J1853+1303 & & & \\
-		J1909$-$3744 & 9.379 & 0.12 & -13.99 & 2.06 \\
-		J1918$-$0642 & & & \\
-		J2010$-$1323 & & & \\
-		J2145$-$0750 & 17.161 & 1.19 & -13.87 & 4.02 \\
-		J2317+1439 & & & \\
+		J0030+0451 & 11.0 & & -13.93 & 3.56 \\
+		J0613$-$0200 & 11.0 &  & -13.14 & 1.22 \\
+		J1012+5307 & 11.0 &  & -12.79 & 1.51 \\
+		J1024$-$0719 & 6.0 & & $-$ & $-$ \\
+		J1455$-$3330 & 11.0 & & $-$ & $-$ \\
+		J1600$-$3053 & 8.0 & & $-$ & $-$ \\
+		J1614$-$2230 & 7.0 & & $-$ & $-$ \\
+		J1640+2224 & 11.0 & & $-$ & $-$ \\
+		J1713+0747 & 11.0 & & -14.14 & 1.58 \\
+		J1741+1351 & 6.0 & & $-$ & $-$ \\
+		J1744$-$1134 & 11.0 & & $-$ & $-$ \\
+		B1855+09 & 11.0 & & -13.75 & 3.54 \\
+		J1853+1303 & 7.0 & & $-$ & $-$ \\
+		J1909$-$3744 & 11.0 &  & -13.84 & 1.74  \\
+		J1918$-$0642 & 11.0 & & $-$ & $-$ \\
+		J2010$-$1323 & 7.0 & & $-$ & $-$ \\
+		J2145$-$0750 & 11.0 &  & -12.69 & 1.30 \\
+		J2317+1439 & 11.0 & & $-$ & $-$ \\
     		\hline\hline
@@
 In Fig.~\ref{fig:os_datasetstats} we compare the fixed-noise values of $\hat{A}^2$ 
-to the mean $\hat{A}^2$ for 425 different realizations of the stochastic background. 
+to the mean $\hat{A}^2$ for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$. 
 The fixed-noise analysis tends to underestimate the strength of the stochastic background, 

### 2018-03-19 manuscript/manuscript.tex
added to conclusions section

@@
 For a monopole signal, the ORF becomes simply
-\[ \Gamma(\theta_{IJ}) = 1 \,\]
+$\Gamma(\theta_{IJ}) = 1$, 
 while for a dipole signal, the ORF becomes
-\[ \Gamma(\theta_{IJ}) = \cos\theta_{IJ} \,. \]
+$\Gamma(\theta_{IJ}) = \cos\theta_{IJ}$.
 
@@
 
-The noise-marginalized optimal statistic is a computationally inexpensive 
-method for assessing the significance of an isotropic stochastic GWB in 
-PTA data.
+The optimal statistic is a computationally inexpensive 
+method for assessing the significance of an isotropic stochastic GWB in PTA data. 
+The optimal statistic specifically looks for the spatial correlations characteristic 
+of a GWB. These analyses are computationally expensive using a full Bayesian approach, 
+but with the optimal statistic the significance of different spatial correlations 
+can be computed very quickly. 
+This makes the optimal statistic particularly well-suited for sky scramble analyses, 
+where the ORF is scrambled many times to simulate randomizing the pulsars' sky locations.
+
+However, the optimal statistic tends to systematically underestimate the strength of a GWB 
+when computed for fixed values of the red noise because of 
+covariances between red noise in individual pulsars and the GWB. 
+In this paper, we introduce an improved method for computing the optimal statistic 
+that marginalizes over the individual pulsars' red noises. 
+This approach \sv{expand...}
 

### 2018-04-25 manuscript/manuscript.tex
lots of edits to text, plots

@@
 \begin{abstract}
-Supermassive black hole binaries (SMBHBs) are predicted to form a 
-high-redshift cosmological population of low-frequency gravitational wave (GW) sources, 
-which added together produce a stochastic gravitational wave background (GWB). 
-Pulsar timing arrays (PTAs) can observe this signal by looking for correlations in the 
-times of arrival (TOAs) of millisecond pulsars (MSPs). 
-In this paper, we introduce an improved frequentist estimator called the noise-marginalized optimal statistic, 
-which can be used to assess the significance of a GWB present in PTA data. 
-We find \ki{complete}
-
+Observations have revealed that nearly all galaxies contain supermassive black holes (SMBHs) at their centers, 
+with masses between \sv{look up}. 
+When galaxies merge, these SMBHs form SMBH binaries (SMBHBs), 
+which emit low-frequency gravitational waves (GWs). 
+The incoherent superposition of these GW sources produce a stochastic GW background (GWB) 
+that can be observed by pulsar timing arrays (PTAs). 
+The optimal statistic is a frequentist estimator for the amplitude of a GWB 
+that specifically looks for the spatial correlations induced by the GWB in PTA data. 
+In this paper, we introduce an improved method for computing the optimal statistic 
+that marginalizes over the red noise in individual pulsars. 
+We use simulations to demonstrate that this method 
+more accurately determines the strength of a GWB, 
+and we use the noise-marginalized optimal statistic to assess the 
+significance of different spatial correlations and 
+by performing sky scrambles.
 \end{abstract}
@@
 Current upper-limits on the stochastic background from PTAs are approaching 
-theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,NANOGrav9yr}
+theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,abb+17b}
 
@@
 
-However, previous work has found that the optimal statistic gives biased results 
+However, the optimal statistic gives biased results 
 for PTAs with significant red noise in the individual pulsars due to the 
@@
 %estimate of \Agw\ and its uncertainty. 
-Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: marginalizing over the individual pulsar red noise parameters using the posterior distributions from a full PTA Bayesian analysis.
-This hybrid approach produces a more precise estimate of the stochastic GWB amplitude,
-\Agw\/, and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
+Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: 
+marginalizing over the individual pulsar red noise parameters 
+using the posterior distributions from a full Bayesian analysis of all the pulsars. 
+This hybrid approach produces a more precise estimate of the stochastic GWB amplitude \Agw\ 
+and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
 %This method is more computationally intensive than previous methods 
@@
 %allowing additional spatial correlations to be tested very quickly.
-Furthermore, the same Bayesian analysis drawn upon by the noise marginalization can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), allowing for quick testing of various spatial correlations.
+Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
+can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), 
+allowing for quick testing of various spatial correlations. 
+This technique was used to perform the frequentist searches 
+for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
+in the NANOGrav 11-year dataset \citep{abb+17b}.
 
 This paper is organized as follows. In Sec.~\ref{sec:marg_os} 
-we lay out the procedure for computing the noise-marginalized optimal statistic 
-and compare to the fixed-noise optimal statistic. 
+we lay out the procedure for computing the noise-marginalized optimal statistic. 
+We use simulations to compare the noise-marginalized optimal statistic 
+to the optimal statistic computed using fixed noise. 
 In Sec.~\ref{sec:spatial} determine how well 
 the noise-marginalized optimal statistic can 
-differentiate between different spatial correlations 
-using simulated PTA data. 
+differentiate between monopole, dipole, and Hellings-Downs spatial correlations. 
 In Sec.~\ref{sec:skyscrambles} we determine how well the noise-marginalized optimal statistic 
@@
 %We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
-We conclude in Sec.~\ref{sec:conclusion}.
+We summarize our results in Sec.~\ref{sec:conclusion} 
+and discuss future applications of the noise-marginalized optimal statistic.
 
@@
 	\setlength{\tabcolsep}{5pt}
-	\caption{Pulsar parameters used in simulated PTAs.}
+	\caption{Pulsar parameters used in simulated PTAs. \sv{finish filling in this table}}
 	\begin{center}
@@
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
-optimal statistic for a single realization of the stochastic background. 
+optimal statistic for a single realization of the stochastic background with $\Agw=$. 
 We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
@@
 to the mean $\hat{A}^2$ for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$. 
-The fixed-noise analysis tends to underestimate the strength of the stochastic background, 
-whereas the noise-marginalized analysis is better able to recover the injected value. 
-The mean value of the optimal statistic from the fixed-noise analysis, 
-averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
-while the mean value from the noise-marginalized analysis is $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-The noise-marginalized analysis also gives a higher average mean SNR of 1.7, 
-compared to an average SNR of 1.0 using the fixed-noise analysis.
-%\begin{figure}[ht]
-%	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15.pdf}
-%	\caption{Optimal statistic and SNR for 425 simulated datasets 
-%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-%			We compare the values found using a fixed-noise analysis (blue) to the 
-%			mean optimal statistic and mean SNR found by marginalizing over 
-%			the individual pulsars' red noise parameters. The dashed vertical line 
-%			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-%			The value of $\hat{A}^2$ from the fixed-noise analysis, 
-%			averaged over realizations of the stochastic background, is $\hat{A}^2 = 8.9 \times 10^{-30}$, 
-%			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-%			Furthermore, the fixed-noise analysis gives an average SNR of 1.0, while the 
-%			noise-marginalized analysis gives an average mean SNR of 1.7.}
-%	\label{fig:os_datasetstats}
-%\end{figure}
+Using the noise values from individual noise analyses 
+systematically underestimates the strength of the GWB, 
+while using the noise values from a common noise analysis more accurately recovers the injected value. 
+The mean value of the optimal statistic from the fixed-noise analysis using the individual noise results, 
+averaged over realizations of the GWB, is $\hat{A}^2 = 7.9 \times10^{-30}$ with a mean SNR of 2.3. 
+The fixed-noise and noise-marginalized analyses using the common noise results both give 
+$\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
+\begin{figure}[ht]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
+	\caption{Optimal statistic and SNR for 200 simulated datasets 
+			containing an injected GWB. 
+			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			We compare the values found by fixing the red noise 
+			based on the results from individual and common noise analyses 
+			(blue and orange lines, respectively) 
+			to the mean optimal statistic and mean SNR found by marginalizing over 
+			the red noise parameters using the results of a common noise analysis (green lines). 
+			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
+			that are significantly lower than the injected value, 
+			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
+	\label{fig:os_datasetstats}
+\end{figure}
+
+The fixed-noise and noise-marginalized analyses gave the same results 
+for an injected GWB with $\Agw = 5\times10^{-15}$, 
+but there are differences between them when analyzing datasets 
+containing smaller injected values of $\Agw$. 
+Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
+for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
+The fixed-noise analysis using the individual noise values 
+gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
+The fixed-noise analysis using the common noise values 
+gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
+which the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.  
+The difference between the fixed-noise and noise-marginalized analyses using the common noise values 
+is very small compared to the large spread in $\hat{A}^2$ and the SNR over different realizations of the GWB.
+\begin{figure}[ht]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_A1e-15.pdf}
+	\caption{Optimal statistic and SNR for 100 simulated datasets 
+			containing an injected GWB. 
+			The dashed vertical line indicates the injected value, $\hat{A}^2 = 10^{-30}$. 
+			We compare the values found by fixing the red noise 
+			based on the results from individual and common noise analyses 
+			(blue and orange lines, respectively) 
+			to the mean optimal statistic and mean SNR found by marginalizing over 
+			the red noise parameters using the results of a common noise analysis (green lines). 
+			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
+			that are significantly lower than the injected value, 
+			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
+	\label{fig:os_datasetstats2}
+\end{figure}
 
@@
 because using a different spatial correlation only requires changing the ORF 
-in $\tilde{{\bf{S}}}_{IJ}$. 
+in Eq.~\eqref{eq:}. 
 \citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
@@
 
-Figure~\ref{fig:os_ORF} shows the noise-marginalized mean value of $\hat{A}^2$ and the mean SNR 
+Our ability to distinguish between different spatial correlations 
+depends on the strength of the GWB 
+and the angular separations between pulsar pairs, $\theta_{IJ}$. 
+We can determine the overlap between ORFs corresponding to different spatial correlations 
+by computing the ``match statistic'' \citep{cs2016},
+\begin{equation}
+	\bar{M} = \frac{\sum_{I,J \neq I} \Gamma_{IJ} \Gamma'_{IJ}}{\sqrt{ \left( \sum_{I, J \neq I} \Gamma_{IJ} \Gamma_{IJ} \right) \left( \sum_{I, J \neq I} \Gamma'_{IJ} \Gamma'_{IJ} \right)}} \,,
+\end{equation}
+where $I$,$J$ index the pulsar number, and $\Gamma$ and $\Gamma'$ are two different ORFs. 
+For the 18 pulsars used in these simulations, the 
+match statistic for monopole and Hellings-Downs correlations is $\bar{M} = 0.264$, 
+and the match statistic for dipole and Hellings-Downs correlations is $\bar{M} = 0.337$. 
+These match statistics describe a fundamental limit on our ability 
+to identify the spatial correlations of a common red signal as Hellings-Downs 
+rather than monopole or dipole 
+that depends only on the number of pulsars in our PTA and their sky positions.
+
+Figure~\ref{fig:os_ORF} shows the noise-marginalized 
+mean value of $\hat{A}^2$ and the mean SNR 
 computed assuming monopole, dipole, and quadrupole spatial correlations 
-for 425 simulated data sets. Using a monopole or dipole ORF 
+for 200 simulated data sets. Using a monopole or dipole ORF 
 gives a lower value for the mean optimal statistic and mean SNR compared to the 
 Hellings-Downs ORF. 
-However, there is significant overlap between the distributions for the mean SNR using 
-monopole, dipole, and quadrupole ORFs, indicating that it is not easy to differentiate 
-between those three spatial correlations for an injected GW background of this amplitude. 
-We find a noise-marginalized mean SNR above 1.0 in 68\% of our simulated data sets 
-using the quadrupole ORF, and in 45\% and 24\% of our simulated data sets 
+We find a noise-marginalized mean SNR above 1.0 in 97\% of our simulated data sets 
+using the quadrupole ORF, and in 50\% and 68\% of our simulated data sets 
 using the monopole and dipole ORFs, respectively. 
-%\begin{figure}[htb]
-%	\includegraphics[width=0.9\columnwidth]{plots/os_datasetstats_5e-15_ORF.pdf}
-%	\caption{Noise-marginalized mean optimal statistic and mean SNR for 425 simulated datasets 
-%			containing an injected isotropic stochastic GWB with $\Agw = 5\times10^{-15}$. 
-%			We compare the values found using monopole (blue), dipole (orange), 
-%			and quadrupole (green) spatial correlations. 
-%			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
-%	\label{fig:os_ORF}
-%\end{figure}
+The mean SNR using the quadrupole ORF, averaged over realizations of the GWB, is 4.1, 
+and we find an SNR greater than this using the monopole and dipole ORFs in just 
+3\% and 3.5\% of our simulations, respectively.
+\begin{figure}[htb]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
+			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
+			We compare the values found using monopole (blue), dipole (orange), 
+			and quadrupole (green) spatial correlations. 
+			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
+	\label{fig:os_ORF}
+\end{figure}
 
@@
 \acknowledgments
-We thank Joe Romano and Xavier Siemens for useful discussions. 
+We thank Joe Romano, Xavier Siemens, and Jeff Hazboun for useful discussions. 
 KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.

### 2018-04-26 manuscript/manuscript.tex
major edits, mostly to Sec. II

@@
 However, the optimal statistic gives biased results 
-for PTAs with significant red noise in the individual pulsars due to the 
-strong covariance between the individual red noise parameters 
+when the pulsars have significant red noise 
+due to the strong covariance between the individual red noise parameters 
 and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
@@
 allowing us to marginalize over the individual red noise. 
+We used the PTA data analysis package 
+PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
+\citep{evh17a}
+to perform the noise analyses and compute the optimal statistic.
 
@@
 (see Table~\ref{tab:sim} for more details).
-To these data sets we injected an isotropic stochastic GWB 
-\sv{insert strength of GWB... are we going to have one value or more than one?}.
-\begin{table}[htb]
+\begin{table}[t]
 	\setlength{\tabcolsep}{5pt}
@@
 
-The optimal statistic was computed using the PTA data analysis package 
-PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
-\citep{evh17a}. 
 Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
-optimal statistic for a single realization of the stochastic background with $\Agw=$. 
-We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
-for a common red process. For this particular realization of the stochastic background, 
-the values of $\hat{A}^2$ from the noise-marginalized and Bayesian analyses are in good agreement 
-with each other and the injected GW stochastic background, 
-but the value from the fixed-noise analysis is significantly lower. 
-The noise-marginalized analysis also gives a larger mean SNR compared to the fixed-noise analysis.
+optimal statistic for a simulation with a GWB with $\Agw=5\times10^{-15}$. 
+%We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
+%for a common red process. 
+For this particular realization of the GWB, 
+the fixed-noise analysis 
+using the individual noise results gives 
+$\hat{A}^2 = 6.6\times10^{-30}$ with $\mathrm{SNR} = 2.4$, 
+and the fixed-noise analysis using the common noise results gives 
+$\hat{A}^2 = 2.6\times10^{-29}$ with $\mathrm{SNR} = 6.0$. 
+The noise-marginalized analysis gives $\hat{A}^2 = (2.5 \pm 0.1)\times10^{-29}$ 
+with $\mathrm{SNR}=4.8\pm0.8$. 
+The value of $\hat{A}^2$ from the fixed-noise analysis using the individual noise results 
+is significantly lower than the injected level of the GWB, while 
+the values of $\hat{A}^2$ from the fixed-noise analysis using the common noise results 
+and the noise-marginalized analysis are in good agreement with each other and the injected value. 
+The fixed-noise analysis using the individual noise results also gives a significantly lower SNR 
+than the analyses using the common noise results. 
+The mean SNR from the noise-marginalized analysis is smaller 
+than the SNR found with the fixed-noise analysis using the common noise results, 
+but the two are consistent at $1.6\sigma$.
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
+	\caption{Optimal statistic for a simulated dataset containing an injected GWB. 
+			We compare the values found using a fixed-noise analysis with noise parameters taken from 
+			individual and common noise analyses (dashed blue lines and solid orange lines, respectively) 
+			to the values found by marginalizing over 
+			the individual pulsars' red noise parameters (solid green lines). 
+			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
+			For this particular realization of the GWB, the fixed-noise analysis 
+			using the individual noise results underestimates the strength of the GWB, 
+			while the fixed-noise analysis using the common noise results 
+			and the noise-marginalized analysis more accurately find $\hat{A}^2$.}
+	\label{fig:os_dataset_sample}
+\end{figure}
 %\begin{figure}[ht]
@@
 
-In Fig.~\ref{fig:os_datasetstats} we compare the fixed-noise values of $\hat{A}^2$ 
-to the mean $\hat{A}^2$ for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$. 
+In Fig.~\ref{fig:os_datasetstats} we show the optimal statistic 
+for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
+computed using the three techniques described above. 
 Using the noise values from individual noise analyses 
@@
 The fixed-noise and noise-marginalized analyses using the common noise results both give 
-$\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
+a mean value of the optimal statistic of $\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
 	\caption{Optimal statistic and SNR for 200 simulated datasets 
@@
 			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			We compare the values found by fixing the red noise 
+			We show the values found by fixing the red noise 
 			based on the results from individual and common noise analyses 
 			(blue and orange lines, respectively) 
-			to the mean optimal statistic and mean SNR found by marginalizing over 
+			as well as the mean optimal statistic and mean SNR found by marginalizing over 
 			the red noise parameters using the results of a common noise analysis (green lines). 
@@
 			that are significantly lower than the injected value, 
-			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
+			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
+			that are closer to the injected value.}
 	\label{fig:os_datasetstats}
@@
 
-The fixed-noise and noise-marginalized analyses gave the same results 
+The fixed-noise and noise-marginalized analyses give the same results 
 for an injected GWB with $\Agw = 5\times10^{-15}$, 
@@
 gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
-which the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.  
-The difference between the fixed-noise and noise-marginalized analyses using the common noise values 
-is very small compared to the large spread in $\hat{A}^2$ and the SNR over different realizations of the GWB.
-\begin{figure}[ht]
-	\includegraphics[width=0.9\columnwidth]{plots/optstat_A1e-15.pdf}
+which the noise-marginalized analysis gives a mean value of 
+$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
 	\caption{Optimal statistic and SNR for 100 simulated datasets 
-			containing an injected GWB. 
-			The dashed vertical line indicates the injected value, $\hat{A}^2 = 10^{-30}$. 
-			We compare the values found by fixing the red noise 
+			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
+			We show the values found by fixing the red noise 
 			based on the results from individual and common noise analyses 
@@
 			that are significantly lower than the injected value, 
-			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
+			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
+			that are closer to the injected value.}
 	\label{fig:os_datasetstats2}
 \end{figure}
+
+We also compare these three methods 
+based on how far the computed values of the optimal statistic are from the injected value of of the GWB 
+compared to the uncertainty in the optimal statistic,
+\begin{equation}
+	s = \frac{\hat{A}^2 - \Agw^2}{\sigma} = \frac{\hat{A}^2 - \Agw^2}{\hat{A}^2/\rho} \,.
+\end{equation}
+Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
 
@@
 3\% and 3.5\% of our simulations, respectively.
-\begin{figure}[htb]
-	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
 	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
@@
 
-The optimal statistic is a computationally inexpensive 
-method for assessing the significance of an isotropic stochastic GWB in PTA data. 
-The optimal statistic specifically looks for the spatial correlations characteristic 
-of a GWB. These analyses are computationally expensive using a full Bayesian approach, 
-but with the optimal statistic the significance of different spatial correlations 
-can be computed very quickly. 
-This makes the optimal statistic particularly well-suited for sky scramble analyses, 
-where the ORF is scrambled many times to simulate randomizing the pulsars' sky locations.
+The optimal statistic is a frequentist estimator for the strength of 
+an isotropic stochastic GWB in PTA data that searches 
+for a common signal with a particular spatial correlation relation 
+using a matched filter approach. 
+Searching for the spatial correlations characteristic of a GWB 
+using a full Bayesian approach is computationally inexpensive, 
+whereas the optimal statistic can be computed in a fraction of the time. 
+This makes the optimal statistic particularly useful for simulations 
+or for analyses where many spatial correlations are compared 
+such as sky scramble analyses.
 

### 2018-04-26 manuscript/manuscript.tex
updated notation in Sec 2 to match NANOGrav 9-yr and 11-yr GWB papers

@@
 \newcommand{\Agw}{\ensuremath{A_\mathrm{gw}}}
+\newcommand{\res}{\ensuremath{\delta {\bf t}}}
 
@@
 
-\sv{bring notation up to date}
-As introduced by \citet{abc+2009}, 
-the optimal statistic is derived by analytically maximizing 
-the PTA likelihood function in the weak-signal regime. 
-We begin by finding the timing residuals for a given pulsar ${\bf{r}}_I$, which can be written as 
+The optimal statistic is a frequentist estimator 
+for the amplitude of an isotropic stochastic GWB, 
+and can be derived by analytically maximizing 
+the PTA likelihood function in the weak-signal regime (see~\citet{abc+2009}). 
+It is constructed from the timing residuals $\res$, 
+which can be written as 
 \begin{equation}
-	{\bf{r}}_I = M {\bf{\epsilon}}_I + F {\bf{a}}_I + U {\bf{j}}_I + {\bf{n}}_I \,,\\
+	\res = M {\bf{\epsilon}} + F {\bf{a}} + U {\bf{j}} + {\bf{n}} \,.
 \end{equation}
-where $M$ is the design matrix, 
-$\epsilon_I$ is a vector of the timing model parameter uncertainties, 
-${\bf{a}}_I$ are the red noise parameters describing both individual pulsars' red noise 
-and common red noise in all pulsars, 
-${\bf{j}}_I$ are white noise parameters describing frequency-correlated white noise, 
-and ${\bf{n}}_I$ are white noise parameters describing uncorrelated white noise. 
-The likelihood function is constructed from the residuals as given by
+The design matrix $M$ and timing model parameter uncertainties $\epsilon$ 
+describe the contributions to the residuals from uncertainties in the timing model. 
+The term $F{\bf{a}}$ describes red noise, 
+including both red noise intrinsic to the pulsar 
+and a common red noise signal common to all the pulsar (i.e., a GW signal). 
+The term $U{\bf{j}}$ describes noise that is correlated for observations made at the same time 
+at different frequencies and uncorrelated over different observing epochs, 
+while ${\bf{n}}$ describes uncorrelated white noise.
+
+The optimal statistic is constructed from the autocovariance and cross-covariance matrices of the residuals 
+${\bf{P}}_a$ and ${\bf{S}}_{ab}$,
+\begin{eqnarray}
+	{\bf{P}}_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
+	{\bf{S}}_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,.
+\end{eqnarray}
+where $a,b$ index pulsars in the PTA. 
+%In the frequency domain, these can be written as
+%\begin{eqnarray}
+%	\left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle_{ij} &=& {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_I(f) \right] {\bf{R}}_I^T \,, \\
+%	\left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{ij} &=& \chi_{IJ} {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_J^T \,,
+%\end{eqnarray}
+%where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
+%two observations, 
+%$\mathbf{R}$ is the timing model fitting matrix, 
+%$\chi_{IJ}$ is the overlap reduction function (ORF) that describes the spatial correlation between two pulsars' residuals, 
+%$P_\mathrm{gw}(f)$ is the GW power spectrum, 
+%and $P_I(f)$ is the power spectrum for the $I$th pulsar. 
+The optimal statistic $\hat{A}^2$ is given by
 \begin{equation}
-	p({\bf{r}}|\phi) = \frac{\exp[-\frac{1}{2} {\bf{r}}^T \Sigma^{-1}(\phi) {\bf{r}} ]}{\sqrt{\mathrm{det}[2\pi \Sigma(\phi)]}} \,,
+	\hat{A}^2 = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) } \,,
 \end{equation}
-where $\bf{r}$ is an array of the timing residuals 
-from the $M$ pulsars in the PTA,
+where $\tilde{{\bf{S}}}_{ab}$ is the amplitude-independent cross-correlation matrix,
 \begin{equation}
-	\bf{r} = \left[ \begin{array}{c} {\bf{r}}_1 \\ {\bf{r}}_2 \\ \vdots \\ {\bf{r}}_M \end{array} \right] \,,
+	\Agw^2 \tilde{{\bf{S}}}_{ab} = {\bf{S}}_{ab} \,.
 \end{equation}
-$\phi$ is the set of pulsar noise parameters, 
-and $\Sigma$ is the covariance matrix of the residuals. 
-We can write $\Sigma$ in terms of 
-autocovariance and cross-covariance matrices ${\bf{P}}_I$ and ${\bf{S}}_{IJ}$, respectively, as
+This definition of the optimal statistic ensures that  
+$\langle \hat{A}^2 \rangle = \Agw^2$. 
+For a weak GW signal, the variance of the optimal statistic is
 \begin{equation}
-	\Sigma = \left[ \begin{array}{cccc} {\bf{P}}_1 & {\bf{S}}_{12} & \hdots & {\bf{S}}_{1M}  \\
-							{\bf{S}}_{21} & {\bf{P}}_2 & \hdots & {\bf{S}}_{2M} \\
-							\vdots & \vdots & \ddots & \vdots \\
-							{\bf{S}}_{M1} & {\bf{S}}_{M2} & \hdots & {\bf{P}}_M \end{array} \right] \,,
-\end{equation}
-where
-\begin{eqnarray}
-	{\bf{P}}_I &=& \left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle \,, \\
-	{\bf{S}}_{IJ} &=& \left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{I \neq J} \,.
-\end{eqnarray}
-In the frequency domain, these can be written as
-\begin{eqnarray}
-	\left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle_{ij} &=& {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_I(f) \right] {\bf{R}}_I^T \,, \\
-	\left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{ij} &=& \chi_{IJ} {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_J^T \,,
-\end{eqnarray}
-where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
-two observations, 
-$\mathbf{R}$ is the timing model fitting matrix, 
-$\chi_{IJ}$ is the overlap reduction function (ORF) that describes the spatial correlation between two pulsars' residuals, 
-$P_\mathrm{gw}(f)$ is the GW power spectrum, 
-and $P_I(f)$ is the power spectrum for the $I$th pulsar.
-
-Analytically maximizing this likelihood gives the optimal statistic,
-\begin{equation}
-	\hat{A}^2 = \frac{\sum_{IJ} {\bf{r}}_I^T {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} {\bf{r}}_J}{\sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) } \,,
-\end{equation}
-where $\tilde{{\bf{S}}}_{IJ}$ is the amplitude-independent cross-correlation matrix,
-\begin{equation}
-	\Agw^2 \tilde{{\bf{S}}}_{IJ} = {\bf{S}}_{IJ} \,.
-\end{equation}
-The normalization of $\hat{A}^2$ has been chosen so that 
-$\langle \hat{A}^2 \rangle = \Agw^2$. 
-For a weak GW signal, the standard deviation is 
-\begin{equation}
-	\sigma_0 = \left[ \sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{-1/2} \,,
+	\sigma_0 = \left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{-1/2} \,,
 \end{equation}
@@
 \begin{equation}
-	\hat{\rho} = \frac{\hat{A}^2}{\sigma_0} = \frac{\sum_{IJ} {\bf{r}}_I^T {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} {\bf{r}}_J}{\left[ \sum_{IJ} \Tr \left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{1/2}} \,.
+	\rho = \frac{\hat{A}^2}{\sigma_0} = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2}} \,.
 \end{equation}
@@
 \begin{equation}
-	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{IJ} \Tr\left( {\bf{P}}_I^{-1} \tilde{{\bf{S}}}_{IJ} {\bf{P}}_J^{-1} \tilde{{\bf{S}}}_{JI} \right) \right]^{1/2} \,.
+	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{ab} \Tr\left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2} \,.
 \end{equation}
 
-When constructing the residuals, 
+When constructing the residuals $\res$, 
 we typically fix the red noise parameters to the values that 

### 2018-04-27 manuscript/manuscript.tex
added text clarifying meaning of SNR in Eq. (7)

@@
 \affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505, USA}
+
+\author{Xavier Siemens}
+\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
 
@@
 $\langle \hat{A}^2 \rangle = \Agw^2$. 
-For a weak GW signal, the variance of the optimal statistic is
+Assuming $\Agw=0$, the variance of the optimal statistic is
 \begin{equation}
@@
 \end{equation}
+The SNR $\rho$ gives the significance of $\hat{A}^2 \neq 0$. 
 The expectation value of the SNR, averaged over all realizations of the stochastic background, is
@@
 \acknowledgments
-We thank Joe Romano, Xavier Siemens, and Jeff Hazboun for useful discussions. 
+We thank Joe Romano and Jeff Hazboun for useful discussions. 
 KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.

### 2018-04-29 manuscript/manuscript.tex
updated histograms in Sec. II

--- 2018-04-26 manuscript/manuscript.tex
@@
 
-The optimal statistic is constructed from the autocovariance and cross-covariance matrices of the residuals 
+The optimal statistic is constructed from the autocovariance and cross-covariance matrices 
 ${\bf{P}}_a$ and ${\bf{S}}_{ab}$,
@@
 \end{eqnarray}
-where $a,b$ index pulsars in the PTA. 
+where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
 %In the frequency domain, these can be written as
@@
 $\langle \hat{A}^2 \rangle = \Agw^2$. 
-For a weak GW signal, the variance of the optimal statistic is
+If $\Agw=0$, the variance of the optimal statistic is
 \begin{equation}
-	\sigma_0 = \left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{-1/2} \,,
+	\sigma_0 = \left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{-1/2} \,.
 \end{equation}
-and the signal-to-noise ratio (SNR) is
+For a measured value of $\hat{A}^2$, 
+the significance of $\hat{A}^2 \neq 0$ is given by the signal-to-noise ratio (SNR) 
 \begin{equation}
@@
 (see Table~\ref{tab:sim} for more details).
-\begin{table}[t]
+\begin{table}[tb]
 	\setlength{\tabcolsep}{5pt}
@@
 but the two are consistent at $1.6\sigma$.
-\begin{figure}[t]
+\begin{figure}[htb]
 	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
-	\caption{Optimal statistic for a simulated dataset containing an injected GWB. 
-			We compare the values found using a fixed-noise analysis with noise parameters taken from 
-			individual and common noise analyses (dashed blue lines and solid orange lines, respectively) 
+	\caption{Optimal statistic for a simulated PTA dataset containing an injected GWB. 
+			We compare the values found using a fixed-noise analysis 
+			with noise parameters taken from 
+			individual and common noise analyses 
+			(dashed blue lines and solid orange lines, respectively) 
 			to the values found by marginalizing over 
-			the individual pulsars' red noise parameters (solid green lines). 
-			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			For this particular realization of the GWB, the fixed-noise analysis 
-			using the individual noise results underestimates the strength of the GWB, 
+			the individual pulsars' red noise parameters (green histograms). 
+			The dotted vertical line indicates the injected value, $\Agw^2 = 2.5 \times 10^{-29}$. 
+			The fixed-noise analysis 
+			using the individual noise results underestimates $\Agw$, 
 			while the fixed-noise analysis using the common noise results 
-			and the noise-marginalized analysis more accurately find $\hat{A}^2$.}
+			and the noise-marginalized analysis more accurately recover $\Agw$.}
 	\label{fig:os_dataset_sample}
@@
 containing smaller injected values of $\Agw$. 
-Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
-for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
-The fixed-noise analysis using the individual noise values 
-gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
-The fixed-noise analysis using the common noise values 
-gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
-which the noise-marginalized analysis gives a mean value of 
-$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
-\begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
-	\caption{Optimal statistic and SNR for 100 simulated datasets 
-			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
-			We show the values found by fixing the red noise 
-			based on the results from individual and common noise analyses 
-			(blue and orange lines, respectively) 
-			to the mean optimal statistic and mean SNR found by marginalizing over 
-			the red noise parameters using the results of a common noise analysis (green lines). 
-			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
-			that are significantly lower than the injected value, 
-			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
-			that are closer to the injected value.}
-	\label{fig:os_datasetstats2}
-\end{figure}
-
+%Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
+%for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
+%The fixed-noise analysis using the individual noise values 
+%gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
+%The fixed-noise analysis using the common noise values 
+%gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
+%which the noise-marginalized analysis gives a mean value of 
+%$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
+%\begin{figure}[t]
+%	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
+%	\caption{Optimal statistic and SNR for 100 simulated datasets 
+%			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
+%			We show the values found by fixing the red noise 
+%			based on the results from individual and common noise analyses 
+%			(blue and orange lines, respectively) 
+%			to the mean optimal statistic and mean SNR found by marginalizing over 
+%			the red noise parameters using the results of a common noise analysis (green lines). 
+%			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
+%			that are significantly lower than the injected value, 
+%			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
+%			that are closer to the injected value.}
+%	\label{fig:os_datasetstats2}
+%\end{figure}
 We also compare these three methods 
@@
 Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
+\begin{figure*}[t]
+	\includegraphics[width=0.95\columnwidth]{plots/pp_plot.pdf}
+	\caption{\sv{write caption}}
+	\label{fig:os_compare}
+\end{figure*}
 

### 2018-04-29 manuscript/manuscript.tex
updated histograms in Sec. II

@@
 systematically underestimates the strength of the GWB, 
-while using the noise values from a common noise analysis more accurately recovers the injected value. 
-The mean value of the optimal statistic from the fixed-noise analysis using the individual noise results, 
-averaged over realizations of the GWB, is $\hat{A}^2 = 7.9 \times10^{-30}$ with a mean SNR of 2.3. 
-The fixed-noise and noise-marginalized analyses using the common noise results both give 
-a mean value of the optimal statistic of $\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
+while using the noise values from a common noise analysis 
+more accurately recovers the injected value. 
+The fixed-noise analysis using the individual noise results finds 
+$\hat{A}^2 = (7.9 \pm 6.8) \times10^{-30}$ and $\rho = 2.3 \pm 1.5$, 
+averaging over realizations of the GWB. 
+The fixed-noise and noise-marginalized analyses 
+using the common noise results both give 
+$\hat{A}^2 = (2.4\pm1.2)\times10^{-29}$ and $\rho = 4.1 \pm 1.7$.
 \begin{figure}[t]
@@
 The fixed-noise and noise-marginalized analyses give the same results 
-for an injected GWB with $\Agw = 5\times10^{-15}$, 
+for $\Agw = 5\times10^{-15}$, 
 but there are differences between them when analyzing datasets 

### 2018-04-29 manuscript/manuscript.tex
added P-P plot

@@
 $\hat{A}^2 = (2.4\pm1.2)\times10^{-29}$ and $\rho = 4.1 \pm 1.7$.
-\begin{figure}[t]
+\begin{figure}[htb]
 	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
@@
 We also compare these three methods
-based on how far the computed values of the optimal statistic are from the injected value of of the GWB 
-compared to the uncertainty in the optimal statistic,
-\begin{equation}
-	s = \frac{\hat{A}^2 - \Agw^2}{\sigma} = \frac{\hat{A}^2 - \Agw^2}{\hat{A}^2/\rho} \,.
-\end{equation}
+
 Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
 \begin{figure*}[t]
-	\includegraphics[width=0.95\columnwidth]{plots/pp_plot.pdf}
-	\caption{\sv{write caption}}
+	\includegraphics[width=0.95\textwidth]{plots/pp_plot.pdf}
+	\caption{P--P plot showing the cumulative fraction of simulations for which $\Agw^2$ lies within 
+			a given confidence interval of the measured $\hat{A}^2$. 
+			The probability distribution of $\hat{A}^2$ is assumed to be a Gaussian 
+			with variance $\sigma^2_{\hat{A}^2}$. 
+			The fixed-noise optimal statistic using the individual and common noise results 
+			both give biased values of $\hat{A}^2$, particularly for small values of \Agw, 
+			while the noise-marginalized optimal statistic gives more accurate values of $\hat{A}^2$ 
+			over a large range of injected values of \Agw.}
 	\label{fig:os_compare}

### 2018-04-29 manuscript/manuscript.tex
added P-P plot

@@
 
-\author{Xavier Siemens}
-\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
+%\author{Xavier Siemens}
+%\affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
 
@@
 containing smaller injected values of $\Agw$. 
-%Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
-%for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
-%The fixed-noise analysis using the individual noise values 
-%gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
-%The fixed-noise analysis using the common noise values 
-%gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
-%which the noise-marginalized analysis gives a mean value of 
-%$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
-%\begin{figure}[t]
-%	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
-%	\caption{Optimal statistic and SNR for 100 simulated datasets 
-%			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
-%			We show the values found by fixing the red noise 
-%			based on the results from individual and common noise analyses 
-%			(blue and orange lines, respectively) 
-%			to the mean optimal statistic and mean SNR found by marginalizing over 
-%			the red noise parameters using the results of a common noise analysis (green lines). 
-%			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
-%			that are significantly lower than the injected value, 
-%			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
-%			that are closer to the injected value.}
-%	\label{fig:os_datasetstats2}
-%\end{figure}
-We also compare these three methods
+In Fig.~\ref{fig:os_compare} we show a P--P plot 
+of the cumulative fraction of simulations for which the injected $\Agw^2$ lies within 
+a given confidence interval of the measured $\hat{A}^2$, 
+assuming a Gaussian distribution with mean equal to the measured $\hat{A}^2$ 
+and variance $\sigma^2_{\hat{A}^2}$, 
+determined from the distribution of $\hat{A}^2$ for our simulations. 
+If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
+the curves should lie along a straight line with slope 1 (the dotted, diagonal line).
 
-Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
+We compare the three methods for computing the optimal statistic for 
+simulations with $\Agw=5\times10^{-15}$, $\Agw=10^{-15}$, and $\Agw=5\times10^{-16}$. 
+The fixed-noise optimal statistic using the individual noise results systematically underestimates $\hat{A}^2$ 
+(Fig.~\ref{fig:os_compare}, left panel). 
+The fixed-noise optimal statistic using the common noise results 
+estimates $\hat{A}^2$ well for large values of $\Agw$, but for small values it also underestimates $\hat{A}^2$ 
+(Fig.~\ref{fig:os_compare}, middle panel). 
+The noise-marginalized optimal statistic provides the most accurate estimate of $\hat{A}^2$ 
+over the range of \Agw\ considered here (Fig.~\ref{fig:os_compare}, right panel).
 \begin{figure*}[t]
@@
 \acknowledgments
-We thank Joe Romano and Jeff Hazboun for useful discussions. 
+We thank Joe Romano, Xavier Siemens, Jolien Creighton, and Jeff Hazboun for useful discussions. 
 KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.

### 2018-04-29 manuscript/manuscript.tex
updated Sec III

@@
 where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
-%In the frequency domain, these can be written as
-%\begin{eqnarray}
-%	\left\langle {\bf{r}}_I {\bf{r}}_I^T \right\rangle_{ij} &=& {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_I(f) \right] {\bf{R}}_I^T \,, \\
-%	\left. \left\langle {\bf{r}}_I {\bf{r}}_J^T \right\rangle \right|_{ij} &=& \chi_{IJ} {\bf{R}}_I \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_J^T \,,
-%\end{eqnarray}
-%where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
-%two observations, 
-%$\mathbf{R}$ is the timing model fitting matrix, 
-%$\chi_{IJ}$ is the overlap reduction function (ORF) that describes the spatial correlation between two pulsars' residuals, 
-%$P_\mathrm{gw}(f)$ is the GW power spectrum, 
-%and $P_I(f)$ is the power spectrum for the $I$th pulsar. 
+In the frequency domain, these can be written as
+\begin{eqnarray}
+	{\bf{P}}_a^{ij} &=& {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_a(f) \right] {\bf{R}}_a^T \,, \\
+	{\bf{S}}_{ab}^{ij} &=& \Gamma_{ab} \; {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_b^T \,, \label{eq:Sab}
+\end{eqnarray}
+where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
+two observations, 
+$\mathbf{R}$ is the timing model fitting matrix, 
+$P_\mathrm{gw}(f)$ is the GW power spectrum, 
+and $P_a(f)$ is the power spectrum for the $a$th pulsar. 
+$\Gamma_{ab}$ is the overlap reduction function (ORF) 
+that describes the spatial correlation between two pulsars' residuals \citet{hd1983}, 
+\begin{equation}
+	\Gamma_{ab} =  \,,
+\end{equation}
+where $\theta_{ab}$ is the angle between the pulsars. 
 The optimal statistic $\hat{A}^2$ is given by
@@
 We include white noise for all pulsars, plus 
-red noise parametrized as a power law, $P(f) = A_\mathrm{red} f^{-\gamma}$, 
+red noise parametrized as a power law, $P_a(f) = A_\mathrm{red} f^{-\gamma}$, 
 for those pulsars that show evidence of red noise 
 (see Table~\ref{tab:sim} for more details).
+
 \begin{table}[tb]
@@
 \end{table}
-
-%We constructed simulated data sets using the same procedure as in \citet{tlb+2017}. 
-%The simulated PTAs consisted of 10 MSPs with 
-%total observation times and noise properties chosen to be 
-%similar to the properties of the EPTA \cite{caballero+2016} (see Table~\ref{tab:sim}). 
-%We also injected a stochastic GWB amplitude with $\Agw = 5\times10^{-15}$.
 %\begin{table}[htb]
@@
 but the two are consistent at $1.6\sigma$.
-\begin{figure}[htb]
-	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
+
+\begin{figure}[tb]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
 	\caption{Optimal statistic for a simulated PTA dataset containing an injected GWB. 
@@
 \end{figure}
-%\begin{figure}[ht]
-%	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
-%	\caption{Optimal statistic and SNR for a simulated dataset 
-%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
-%			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
-%			mean optimal statistic and mean SNR found by marginalizing over 
-%			the individual pulsars' red noise parameters (solid orange line). 
-%			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
-%			for a common red process (solid green line). 
-%			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-%			For this particular realization of the stochastic background, the fixed-noise analysis 
-%			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
-%			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
-%			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
-%	\label{fig:os_dataset_sample}
-%\end{figure}
 
@@
 $\hat{A}^2 = (2.4\pm1.2)\times10^{-29}$ and $\rho = 4.1 \pm 1.7$.
-\begin{figure}[htb]
-	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
+
+\begin{figure}[tb]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
 	\caption{Optimal statistic and SNR for 200 simulated datasets 
@@
 
-The fixed-noise and noise-marginalized analyses give the same results 
-for $\Agw = 5\times10^{-15}$, 
-but there are differences between them when analyzing datasets 
-containing smaller injected values of $\Agw$. 
-In Fig.~\ref{fig:os_compare} we show a P--P plot 
-of the cumulative fraction of simulations for which the injected $\Agw^2$ lies within 
-a given confidence interval of the measured $\hat{A}^2$, 
-assuming a Gaussian distribution with mean equal to the measured $\hat{A}^2$ 
-and variance $\sigma^2_{\hat{A}^2}$, 
-determined from the distribution of $\hat{A}^2$ for our simulations. 
-If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
-the curves should lie along a straight line with slope 1 (the dotted, diagonal line).
-
-We compare the three methods for computing the optimal statistic for 
-simulations with $\Agw=5\times10^{-15}$, $\Agw=10^{-15}$, and $\Agw=5\times10^{-16}$. 
-The fixed-noise optimal statistic using the individual noise results systematically underestimates $\hat{A}^2$ 
-(Fig.~\ref{fig:os_compare}, left panel). 
-The fixed-noise optimal statistic using the common noise results 
-estimates $\hat{A}^2$ well for large values of $\Agw$, but for small values it also underestimates $\hat{A}^2$ 
-(Fig.~\ref{fig:os_compare}, middle panel). 
-The noise-marginalized optimal statistic provides the most accurate estimate of $\hat{A}^2$ 
-over the range of \Agw\ considered here (Fig.~\ref{fig:os_compare}, right panel).
-\begin{figure*}[t]
+\begin{figure*}[ht]
 	\includegraphics[width=0.95\textwidth]{plots/pp_plot.pdf}
@@
 
+The fixed-noise and noise-marginalized analyses give the same results 
+for $\Agw = 5\times10^{-15}$, 
+but there are differences between them when analyzing datasets 
+containing smaller injected values of $\Agw$. 
+In Fig.~\ref{fig:os_compare} we show a P--P plot 
+of the cumulative fraction of simulations for which the injected $\Agw^2$ lies within 
+a given confidence interval of the measured $\hat{A}^2$, 
+assuming a Gaussian distribution with mean equal to the measured $\hat{A}^2$ 
+and variance $\sigma^2_{\hat{A}^2}$, 
+determined from the distribution of $\hat{A}^2$ for our simulations. 
+If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
+the curves should lie along a straight line with slope 1 (the dotted, diagonal lines in Fig.~\ref{fig:os_compare}).
+
+We compare the three methods for computing the optimal statistic for 
+simulations with $\Agw=5\times10^{-15}$, $\Agw=10^{-15}$, and $\Agw=5\times10^{-16}$. 
+The fixed-noise optimal statistic using the individual noise results systematically underestimates $\hat{A}^2$ 
+(Fig.~\ref{fig:os_compare}, left panel). 
+The fixed-noise optimal statistic using the common noise results 
+estimates $\hat{A}^2$ well for large values of $\Agw$, but for small values it also underestimates $\hat{A}^2$ 
+(Fig.~\ref{fig:os_compare}, middle panel). 
+The noise-marginalized optimal statistic provides the most accurate estimate of $\hat{A}^2$ 
+over the range of \Agw\ considered here (Fig.~\ref{fig:os_compare}, right panel).
+
 \section{Monopole and Dipole Spatial Correlations}
@@
 because using a different spatial correlation only requires changing the ORF 
-in Eq.~\eqref{eq:}. 
+in Eq.~\eqref{eq:Sab}. 
 \citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
@@
 For a monopole signal, the ORF becomes simply
-$\Gamma(\theta_{IJ}) = 1$, 
+$\Gamma(\theta_{ab}) = 1$, 
 while for a dipole signal, the ORF becomes
-$\Gamma(\theta_{IJ}) = \cos\theta_{IJ}$.
+$\Gamma(\theta_{ab}) = \cos\theta_{ab}$.
 
@@
 depends on the strength of the GWB 
-and the angular separations between pulsar pairs, $\theta_{IJ}$. 
+and the angular separations between pulsar pairs, $\theta_{ab}$. 
 We can determine the overlap between ORFs corresponding to different spatial correlations 
@@
 \begin{equation}
-	\bar{M} = \frac{\sum_{I,J \neq I} \Gamma_{IJ} \Gamma'_{IJ}}{\sqrt{ \left( \sum_{I, J \neq I} \Gamma_{IJ} \Gamma_{IJ} \right) \left( \sum_{I, J \neq I} \Gamma'_{IJ} \Gamma'_{IJ} \right)}} \,,
+	\bar{M} = \frac{\sum_{a,b \neq a} \Gamma_{ab} \Gamma'_{ab}}{\sqrt{ \left( \sum_{a, b \neq a} \Gamma_{ab} \Gamma_{ab} \right) \left( \sum_{a, b \neq a} \Gamma'_{ab} \Gamma'_{ab} \right)}} \,,
 \end{equation}
-where $I$,$J$ index the pulsar number, and $\Gamma$ and $\Gamma'$ are two different ORFs. 
+where $\Gamma$ and $\Gamma'$ are two different ORFs. 
 For the 18 pulsars used in these simulations, the 
@@
 Hellings-Downs ORF. 
+Using quadrupole spatial correlations gives $\hat{A}^2 = (2.4 \pm 1.2) \times 10^{-29}$, 
+while using monopole spatial correlations gives $\hat{A}^2 = (2.5 \pm 3.1) \times 10^{-30}$, 
+and dipole spatial correlations gives $\hat{A}^2 = (5.3 \pm 4.2) \times 10^{-30}$. 
 We find a noise-marginalized mean SNR above 1.0 in 97\% of our simulated data sets 
@@
 \acknowledgments
-We thank Joe Romano, Xavier Siemens, Jolien Creighton, and Jeff Hazboun for useful discussions. 
+We thank Joe Romano, Xavier Siemens, Jolien Creighton, and Jeffrey Hazboun for useful discussions. 
 KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.

### 2018-04-29 manuscript/manuscript.tex
added Justin's new affiliation

@@
 \affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505, USA}
+\affiliation{Infinia ML, 202 Rigsbee Avenue, Durham, NC 27701, USA}
 
@@
 \end{equation}
-The expectation value of the SNR, averaged over all realizations of the stochastic background, is
-\begin{equation}
-	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{ab} \Tr\left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2} \,.
-\end{equation}
+%The expectation value of the SNR, averaged over all realizations of the stochastic background, is
+%\begin{equation}
+%	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{ab} \Tr\left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2} \,.
+%\end{equation}
 
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[htb]
 	\includegraphics[width=0.95\textwidth]{plots/pp_plot.pdf}
@@
 
-Another technique for testing the spatial correlations is through ``sky scrambles,'' 
+The significance of spatial correlations can also be tested with ``sky scrambles,'' 
 where the ORF is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
@@
 
-We used the same scrambled sky positions as were used in \citet{tlb+2017}. 
-These positions were chosen so as to minimize the overlap  
-between the scrambled ORFs and the true ORF, 
-which can be assessed using the ``match statistic'' \citep{cs2016}:
-\begin{equation}
-	\bar{M} = \frac{\sum_{I,J \neq I} \Gamma_{IJ} \Gamma'_{IJ}}{\sqrt{ \left( \sum_{I, J \neq I} \Gamma_{IJ} \Gamma_{IJ} \right) \left( \sum_{I, J \neq I} \Gamma'_{IJ} \Gamma'_{IJ} \right)}} \,,
-\end{equation}
-where $I$,$J$ index the pulsar number, and $\Gamma$ and $\Gamma'$ are two different ORFs. 
-The 210 sky scrambles used in this paper were generated 
-using a particle swarm optimization (PSO) algorithm \citep{ke1995,se1998} 
-with the requirement that the scrambled ORFs had $\bar{M} < 0.1$ when compared to the true ORF and to each other.
-\sv{Steve, check this section}
-
+We generated \sv{XXX} scrambled ORFs 
+using a Monte Carlo algorithm. 
+We required the scrambled ORFs to have $\bar{M} < 0.2$ 
+with respect to the true ORF and each other 
+in order to minimize the overlap. 
 Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 

### 2018-04-30 manuscript/manuscript.tex
added HD ORF

@@
 due to the strong covariance between the individual red noise parameters 
-and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
+and the GWB amplitude. 
 %In this paper, we present a technique for improving the accuracy of the optimal statistic 
@@
 $\Gamma_{ab}$ is the overlap reduction function (ORF) 
-that describes the spatial correlation between two pulsars' residuals \citet{hd1983}, 
-\begin{equation}
-	\Gamma_{ab} =  \,,
-\end{equation}
+that describes the spatial correlation between two pulsars' residuals \citep{hd1983}, 
+\begin{eqnarray}
+	\Gamma_{ab} &=& \frac{1}{2} \left[ 1 - \frac{1}{2} \left( \frac{1-\cos\theta_{ab}}{2} \right) \right. \nonumber \\
+		&& \left. + 3 \left( \frac{1-\cos\theta_{ab}}{2} \right) \ln \left( \frac{1-\cos\theta_{ab}}{2} \right) \right]  \,,
+\end{eqnarray}
 where $\theta_{ab}$ is the angle between the pulsars. 
@@
 
-When constructing the residuals $\res$, 
+When constructing the residuals $\res_a$, 
 we typically fix the red noise parameters to the values that 
@@
 are highly covariant.
+
 In this section, we compare three different techniques for computing the 
@@
 with respect to the true ORF and each other 
-in order to minimize the overlap. 
-Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 
-for a particular data set to the distribution of noise-marginalized optimal statistics obtained using sky scrambles. 
-For this data set, the noise-marginalized mean optimal statistic is 
-$1.9\times10^{-29}$ and the mean SNR is 1.1. 
-The distributions of the noise-marginalized optimal statistic and SNR obtained using the scrambled ORFs 
-are both centered around zero, 
-and for 24 out of the 210 sky scrambles, the SNR is above the true value of $1.1$ ($p=0.11$). 
-Figure~\ref{fig:skyscrambles} shows the $p$-values for 210 sky scrambles 
-using 425 realizations of the stochastic background. 
-For 62\% of simulations the $p$-value is $<0.1$, and for 29\% of simulations the $p$-value is $<0.01$.
+in order to minimize the overlap between the ORFs. 
+%Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 
+%for a particular data set to the distribution of noise-marginalized optimal statistics obtained using sky scrambles. 
+%For this data set, the noise-marginalized mean optimal statistic is 
+%$1.9\times10^{-29}$ and the mean SNR is 1.1. 
+%The distributions of the noise-marginalized optimal statistic and SNR obtained using the scrambled ORFs 
+%are both centered around zero, 
+%and for 24 out of the 210 sky scrambles, the SNR is above the true value of $1.1$ ($p=0.11$). 
+%Figure~\ref{fig:skyscrambles} shows the $p$-values for 210 sky scrambles 
+%using 425 realizations of the stochastic background. 
+%For 62\% of simulations the $p$-value is $<0.1$, and for 29\% of simulations the $p$-value is $<0.01$.
 %\begin{figure}[htb]
@@
 %\end{figure}
-
-For five realizations of the stochastic background, we also computed the Bayes' factors for the 
-true and sky scrambled ORFs in order to directly compare 
-the frequentist and Bayesian detection techniques.
-\sv{comparison to Steve's}
 

### 2018-05-01 manuscript/manuscript.tex
edited figure captions

@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
-	\caption{Optimal statistic for a simulated PTA dataset containing an injected GWB. 
-			We compare the values found using a fixed-noise analysis 
-			with noise parameters taken from 
-			individual and common noise analyses 
-			(dashed blue lines and solid orange lines, respectively) 
-			to the values found by marginalizing over 
-			the individual pulsars' red noise parameters (green histograms). 
-			The dotted vertical line indicates the injected value, $\Agw^2 = 2.5 \times 10^{-29}$. 
-			The fixed-noise analysis 
-			using the individual noise results underestimates $\Agw$, 
-			while the fixed-noise analysis using the common noise results 
-			and the noise-marginalized analysis more accurately recover $\Agw$.}
+	\caption{Optimal statistic 
+			for a simulated PTA dataset containing a GWB with $\Agw = 5\times10^{-15}$. 
+			The fixed-noise analysis using the individual noise values (dashed blue lines) 
+			systematically underestimates $\hat{A}^2$, while the 
+			fixed-noise analysis using the common noise values (solid orange lines) 
+			and the noise-marginalized analysis (green histograms) 
+			more accurately recover $\Agw$.}
 	\label{fig:os_dataset_sample}
@@
 computed using the three techniques described above. 
+For the noise-marginalized analysis, we plot 
+the mean values of $\hat{A}^2$ and $\rho$. 
 Using the noise values from individual noise analyses 
@@
 	\caption{Optimal statistic and SNR for 200 simulated datasets 
-			containing an injected GWB. 
-			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
-			We show the values found by fixing the red noise 
-			based on the results from individual and common noise analyses 
-			(blue and orange lines, respectively) 
-			as well as the mean optimal statistic and mean SNR found by marginalizing over 
-			the red noise parameters using the results of a common noise analysis (green lines). 
-			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
-			that are significantly lower than the injected value, 
-			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
-			that are closer to the injected value.}
+			containing a GWB with $\Agw = 5\times10^{-15}$. 
+			The fixed-noise analysis using the individual noise values (blue) 
+			systematically underestimates $\hat{A}^2$, while both the 
+			fixed-noise analysis using the common noise values (orange) 
+			and the noise-marginalized analysis (green) 
+			accurately recover $\Agw$.}
 	\label{fig:os_datasetstats}
@@
 
-The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
+The optimal statistic is particularly well-suited 
+to compare multiple spatial correlation relations 
 because using a different spatial correlation only requires changing the ORF 
@@
 clock error and ephemeris error. 
-Here we compute the optimal statistic with monopole and dipole spatial correlations 
-for the same simulated data sets as in the previous section in order to determine 
-how well we can distinguish a GW background from a monopole or dipole signal. 
+Here we take a different approach -- rather than simultaneously fitting 
+for signals with different spatial correlations, 
+we look at how well we can distinguish between different spatial correlations 
+by computing the optimal statistic with monopole and dipole spatial correlations 
+for the same simulated data sets as in the previous section. 
 For a monopole signal, the ORF becomes simply
-$\Gamma(\theta_{ab}) = 1$, 
+$\Gamma_{ab} = 1$, 
 while for a dipole signal, the ORF becomes
-$\Gamma(\theta_{ab}) = \cos\theta_{ab}$.
+$\Gamma_{ab} = \cos\theta_{ab}$.
 
@@
 
-We generated \sv{XXX} scrambled ORFs 
+We generated 725 scrambled ORFs 
 using a Monte Carlo algorithm. 
@@
 in order to minimize the overlap between the ORFs. 
+For each simulation, 
+we computed the noise-marginalized mean optimal statistic and mean SNR 
+for each scrambled ORF, and compared to the values found using the true ORF. 
+Figure~\ref{fig:skyscrambles_dataset_sample} shows the results of a sky scramble analysis 
+for a sample dataset with $\Agw=5\times10^{-15}$. For this particular realization, 
+the noise-marginalized optimal statistic 
+using the true ORF found $\hat{A}^2=( \pm ) \times 10^{-29}$ 
+and $\rho = \pm $. 
+None of the 725 scrambled ORFs resulted in a mean SNR greater than 4.8 ($p < 0.0014$).
+\begin{figure}[htb]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_scrambled_dataset11.pdf}
+	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
+			with and without sky scrambles for a simulated dataset 
+			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
+			The true ORF gave a mean SNR of 4.8. 
+			None of the 725 scrambled ORFs gave a mean SNR this large ($p < 0.0014$).}
+	\label{fig:skyscrambles_dataset_sample}
+\end{figure}
+
+In Fig.~\ref{fig:skyscrambles_stats}, we plot the distribution of $p$-values of the 725 sky scrambles 
+for 200 realizations of the GWB.
+
 %Figure~\ref{fig:skyscrambles_dataset_sample} compares the noise-marginalized optimal statistic 

### 2018-05-01 manuscript/manuscript.tex
added sky scramble section

@@
 \newcommand{\res}{\ensuremath{\delta {\bf t}}}
+
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{4}
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\textfraction}{0.15}
+\renewcommand{\floatpagefraction}{0.7}
 
@@
 including both red noise intrinsic to the pulsar 
-and a common red noise signal common to all the pulsar (i.e., a GW signal). 
+and a common red noise signal common to all the pulsar (such as a GW signal). 
 The term $U{\bf{j}}$ describes noise that is correlated for observations made at the same time 
@@
 The optimal statistic is constructed from the autocovariance and cross-covariance matrices 
-${\bf{P}}_a$ and ${\bf{S}}_{ab}$,
-\begin{eqnarray}
-	{\bf{P}}_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
-	{\bf{S}}_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,.
-\end{eqnarray}
-where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
+${\bf{P}}_a$ and ${\bf{S}}_{ab}$. 
+%\begin{eqnarray}
+%	{\bf{P}}_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
+%	{\bf{S}}_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,.
+%\end{eqnarray}
+%where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
 In the frequency domain, these can be written as
@@
 maximize the likelihood. 
-However, this leads to a bias in the optimal statistic and SNR because 
-the individual red noise parameters and the common red noise parameters 
+However, this leads to a bias in the optimal statistic 
+because the individual red noise and common red noise parameters 
 are highly covariant. 
-
-In this section, we compare three different techniques for computing the 
-optimal statistic for PTAs with significant red noise. 
+In this section, we compare three techniques for computing the optimal statistic. 
 First, we fix the individual pulsars' red noise parameters to the 
@@
 a common red process. 
-For the third method, we draw values of the individual pulsars' 
+For the noise-marginalized method, we draw values of the individual pulsars' 
 red noise parameters from the posteriors generated by a Bayesian analysis 
-that includes both individual red noise and a common red noise process, 
-allowing us to marginalize over the individual red noise. 
-We used the PTA data analysis package 
+that includes both individual red noise and a common red noise process.
+
+We use these methods to compute the optimal statistic 
+for simulated ``NANOGrav-like'' datasets 
+consisting of 18 MSPs with observation times up to 11 years 
+and sky positions matching the 18 longest-observed pulsars in the 
+NANOGrav 11-year data set \citep{abb+17}. 
+We include white noise for all pulsars, plus 
+red noise parametrized as a power law, $P_a(f) = A_\mathrm{red} f^{-\gamma}$, 
+for those pulsars that show evidence of red noise 
+(see Table~\ref{tab:sim} for more details).
+We use the PTA data analysis package 
 PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
@@
 
-We used these three techniques to compute the optimal statistic 
-for simulated ``NANOGrav-like'' data sets. 
-These consisted of 18 MSPs with observation times up to 11 years 
-and sky positions matching the 18 longest-observed pulsars in the 
-NANOGrav 11-year data set \citep{abb+17}.
-with the observation times increased so that the longest-observed pulsars 
-had TOAs spanning over 11 years. 
-We include white noise for all pulsars, plus 
-red noise parametrized as a power law, $P_a(f) = A_\mathrm{red} f^{-\gamma}$, 
-for those pulsars that show evidence of red noise 
-(see Table~\ref{tab:sim} for more details).
-
-\begin{table}[tb]
+\begin{table}[!tb]
 	\setlength{\tabcolsep}{5pt}
@@
 \end{table}
+
 %\begin{table}[htb]
@@
 than the analyses using the common noise results. 
-The mean SNR from the noise-marginalized analysis is smaller 
-than the SNR found with the fixed-noise analysis using the common noise results, 
-but the two are consistent at $1.6\sigma$.
-
-\begin{figure}[tb]
-	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
-	\caption{Optimal statistic 
-			for a simulated PTA dataset containing a GWB with $\Agw = 5\times10^{-15}$. 
-			The fixed-noise analysis using the individual noise values (dashed blue lines) 
-			systematically underestimates $\hat{A}^2$, while the 
-			fixed-noise analysis using the common noise values (solid orange lines) 
-			and the noise-marginalized analysis (green histograms) 
-			more accurately recover $\Agw$.}
-	\label{fig:os_dataset_sample}
-\end{figure}
+%The mean SNR from the noise-marginalized analysis is smaller 
+%than the SNR found with the fixed-noise analysis using the common noise results, 
+%but the two are consistent at $1.6\sigma$.
 
@@
 
-\begin{figure}[tb]
+\begin{figure}[!h]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
+	\caption{Optimal statistic 
+			for a simulated PTA dataset containing a GWB with $\Agw = 5\times10^{-15}$. 
+			The fixed-noise analysis using the individual noise values (dashed blue lines) 
+			systematically underestimates $\hat{A}^2$, while the 
+			fixed-noise analysis using the common noise values (solid orange lines) 
+			and the noise-marginalized analysis (green histograms) 
+			more accurately recover $\Agw$.}
+	\label{fig:os_dataset_sample}
+\end{figure}
+
+\begin{figure}[!h]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[tb]
 	\includegraphics[width=0.95\textwidth]{plots/pp_plot.pdf}
@@
 \begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
 	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
@@
 Figure~\ref{fig:skyscrambles_dataset_sample} shows the results of a sky scramble analysis 
-for a sample dataset with $\Agw=5\times10^{-15}$. For this particular realization, 
-the noise-marginalized optimal statistic 
-using the true ORF found $\hat{A}^2=( \pm ) \times 10^{-29}$ 
-and $\rho = \pm $. 
-None of the 725 scrambled ORFs resulted in a mean SNR greater than 4.8 ($p < 0.0014$).
-\begin{figure}[htb]
+for a sample dataset with $\Agw=5\times10^{-15}$. 
+For this particular realization of the GWB, 
+none of the 725 scrambled ORFs resulted in a mean SNR greater than 
+the mean SNR using the true ORF ($p < 0.0014$).
+In Fig.~\ref{fig:skyscrambles_pvalues}, we plot the distribution of $p$-values of the 725 sky scrambles. 
+For a GWB with $\Agw = 5\times10^{-15}$, 96\% of the simulations have $p \leq 0.05$, 
+and 75\% of the simulations have $p \leq 0.003$. 
+For a GWB with $\Agw = 10^{-15}$, 60\% of the simulations have $p \leq 0.05$, 
+and 33\% have $p \leq 0.003$.
+\begin{figure}[tb]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_scrambled_dataset11.pdf}
@@
 			with and without sky scrambles for a simulated dataset 
-			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
-			The true ORF gave a mean SNR of 4.8. 
-			None of the 725 scrambled ORFs gave a mean SNR this large ($p < 0.0014$).}
+			containing a GWB with $\Agw = 5\times10^{-15}$. 
+			None of the 725 scrambled ORFs gave a mean SNR as large 
+			as the mean SNR using the true ORF ($p < 0.0014$).}
 	\label{fig:skyscrambles_dataset_sample}
@@
 
-In Fig.~\ref{fig:skyscrambles_stats}, we plot the distribution of $p$-values of the 725 sky scrambles 
-for 200 realizations of the GWB.
+\begin{figure}[tb]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_scrambled.pdf}
+	\caption{Distribution of $p$-values for the noise-marginalized optimal statistic mean SNR 
+			using the true ORF compared to 725 sky scrambles. 
+			We show results for simulations with $\Agw=5\times10^{-15}$ (blue) and $\Agw=10^{-15}$ (orange).}
+	\label{fig:skyscrambles_pvalues}
+\end{figure}
 
@@
 
-The optimal statistic is a frequentist estimator for the strength of 
-an isotropic stochastic GWB in PTA data that searches 
-for a common signal with a particular spatial correlation relation 
+In this paper, we introduce an improved method for computing the optimal statistic, 
+a frequentist estimator for the strength of an isotropic stochastic GWB in PTA data 
+that searches for a common signal with a particular spatial correlation relation 
 using a matched filter approach. 
+Our method uses the output from a Bayesian analyses 
+for individual and common red signals 
+to marginalize the optimal statistic over the individual pulsars' red noises. 
+As shown in Sec.~\ref{sec:marg_os}, 
+the noise-marginalized optimal statistic more accurately recovers the GWB amplitude 
+than the fixed-noise optimal statistic, which underestimates the GWB amplitude 
+when significant red noise is present in some pulsars.
+
 Searching for the spatial correlations characteristic of a GWB 
@@
 
-However, the optimal statistic tends to systematically underestimate the strength of a GWB 
-when computed for fixed values of the red noise because of 
-covariances between red noise in individual pulsars and the GWB. 
-In this paper, we introduce an improved method for computing the optimal statistic 
-that marginalizes over the individual pulsars' red noises. 
-This approach \sv{expand...}
-
 \acknowledgments
 We thank Joe Romano, Xavier Siemens, Jolien Creighton, and Jeffrey Hazboun for useful discussions. 
-KPI and SJV are supported by NSF Physics Frontier Center Grant 1430284.
+This work was supported by NSF Physics Frontier Center Grant 1430284. 
 JAE was supported by NASA through Einstein Fellowship Grant PF4-150120. 

### 2018-05-02 manuscript/manuscript.tex
edits to text

@@
 \begin{abstract}
-Observations have revealed that nearly all galaxies contain supermassive black holes (SMBHs) at their centers, 
-with masses between \sv{look up}. 
+Observations have revealed that nearly all galaxies 
+contain supermassive black holes (SMBHs) at their centers. 
 When galaxies merge, these SMBHs form SMBH binaries (SMBHBs), 
@@
 composed of millisecond pulsars (MSPs) \cite{hd1983,fb1990}. 
-The dominant astrophysical source in this frequency range is an isotropic stochastic 
+The dominant astrophysical source in this frequency range is the isotropic stochastic 
 gravitational wave background (GWB) 
@@
 Current upper-limits on the stochastic background from PTAs are approaching 
-theoretical predictions for the stochastic GWB \citep{PPTA2013,EPTA2015,abb+17b}
+theoretical predictions for the GWB \citep{PPTA2013,EPTA2015,abb+17b}
 
@@
 to compare the probabilities of various models for the pulsar timing residuals, 
-including one where the residuals contain evidence of an isotropic GW stochastic background 
+including one where the residuals contain evidence of the GWB 
 \citep{vlm+2009,lah+2013}. 
-Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
+%Bayesian inference is a powerful technique because it allows us to compare arbitrarily complicated models. 
+Bayesian inference is a powerful tool because 
+it performs a non-linear fit and 
+accounts for degeneracies between parameters. 
 However, running a full Bayesian analysis is computationally intensive, 
 particularly when searching for evidence of dipole spatial correlations:
-``the smoking gun'' of a stochastic GWB.
+``the smoking gun'' of the GWB.
 
-The significance of a stochastic GWB can also be assessed using the 
+The significance of the GWB can also be assessed using the 
 optimal statistic, a frequentist estimator for the GWB amplitude \citep{abc+2009,demorest+2013,ccs+2015}. 
@@
 using the posterior distributions from a full Bayesian analysis of all the pulsars. 
-This hybrid approach produces a more precise estimate of the stochastic GWB amplitude \Agw\ 
-and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
+This hybrid approach produces a more precise estimate of the GWB amplitude \Agw\ 
+and its uncertainty, while requiring only a few minutes more than the more traditional method of computation. 
 %This method is more computationally intensive than previous methods 
@@
 allowing for quick testing of various spatial correlations. 
-This technique was used to perform the frequentist searches 
+This technique is used to perform the frequentist searches 
 for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
@@
 
-\begin{figure}[!h]
+\begin{figure}[tb]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
@@
 
-\begin{figure}[!h]
+\begin{figure}[tb]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
@@
 3\% and 3.5\% of our simulations, respectively.
-\begin{figure}[t]
+\begin{figure}[tb]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
@@
 for each scrambled ORF, and compared to the values found using the true ORF.
+
 Figure~\ref{fig:skyscrambles_dataset_sample} shows the results of a sky scramble analysis 
@@
 
-In this paper, we introduce an improved method for computing the optimal statistic, 
-a frequentist estimator for the strength of an isotropic stochastic GWB in PTA data 
-that searches for a common signal with a particular spatial correlation relation 
-using a matched filter approach. 
-Our method uses the output from a Bayesian analyses 
+The definitive signature of a GWB in PTA data is spatial correlations 
+described by the Hellings-Downs curve. 
+Searching for these using a full Bayesian approach is computationally expensive, 
+requiring \sv{estimate of CPU time}. 
+In contrast, the optimal statistic can be computed in minutes. 
+In this paper, we introduce an improved method 
+for computing the optimal statistic, 
+which uses the output from a Bayesian analysis 
 for individual and common red signals 
@@
 
-Searching for the spatial correlations characteristic of a GWB 
-using a full Bayesian approach is computationally inexpensive, 
-whereas the optimal statistic can be computed in a fraction of the time. 
-This makes the optimal statistic particularly useful for simulations 
-or for analyses where many spatial correlations are compared 
-such as sky scramble analyses.
+Although the noise-marginalized optimal statistic 
+requires computing the optimal statistic thousands of times, 
+it is still many orders of magnitude faster than a Bayesian search. 
+Furthermore, the results from a single Bayesian analysis, 
+which are needed to marginalize over the red noise parameters, 
+can be used to compute the optimal statistic for many different spatial correlations. 
+In Sec.~\ref{sec:spatial} we use the noise-marginalized optimal statistic 
+to compare the strength of monopole, dipole, and quadrupole correlations 
+in simulated PTA data with a GWB. 
+In Sec.~\ref{sec:skyscrambles} we use the noise-marginalized optimal statistic 
+to perform sky scramble analyses, where we compare the mean SNR 
+computed using the true ORF to the mean SNR computed using scrambled ORFs 
+and measure the significance of HD spatial correlations through the $p$-value.
+
+The primary strength of the optimal statistic is how quickly it can be computed. 
+This is useful for analyses where the significance of many spatial correlations are compared, 
+as with the sky scrambles. 
+It also makes the optimal statistic a valuable tool for analyzing simulations 
+where many realizations of the GWB are compared. 
+The noise marginalization technique described in this paper is key 
+to being able to accurately measure the GWB for PTA data 
+where significant red noise is present.
 
@@
 This work was supported by NSF Physics Frontier Center Grant 1430284. 
-JAE was supported by NASA through Einstein Fellowship Grant PF4-150120. 
-SRT was supported by appointment to the NASA Postdoctoral Program 
+SRT was partially supported by appointment to the NASA Postdoctoral Program 
 at the Jet Propulsion Laboratory, which is administered by Oak Ridge Associated Universities 
 and the Universities Space Research Association through a contract with NASA. 
+JAE was partially supported by NASA through Einstein Fellowship Grant PF4-150120.
 

### 2018-05-02 manuscript/manuscript.tex
updates to notation (Sec. II)

@@
 
-\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\Tr}{tr}
 
@@
 
-\title{Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic}
+\title{The noise-marginalized optimal statistic: A robust hybrid frequentist-Bayesian statistic for a stochastic gravitational-wave background in pulsar timing arrays}
+%\title{Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic}
 
@@
 In Sec.~\ref{sec:skyscrambles} we determine how well the noise-marginalized optimal statistic 
-can assess the significance of a GW detection by performing sky scrambles \citep{cs2016}. 
+can assess the significance of a GW detection by performing sky scrambles \citep{cs2016, tlb+2017}. 
 %We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
@@
 describe the contributions to the residuals from uncertainties in the timing model. 
+The term $U{\bf{j}}$ describes noise that is correlated for observations made at the same time 
+at different frequencies and uncorrelated over different observing epochs, 
+while ${\bf{n}}$ describes uncorrelated white noise.
 The term $F{\bf{a}}$ describes red noise, 
@@
 and a common red noise signal common to all the pulsar (such as a GW signal). 
-The term $U{\bf{j}}$ describes noise that is correlated for observations made at the same time 
-at different frequencies and uncorrelated over different observing epochs, 
-while ${\bf{n}}$ describes uncorrelated white noise.
+We model the red noise as a Fourier series,
+\begin{equation}
+	F {\bf{a}} = \sum_{j=1}^N \left[ a_j \sin\left( \frac{2\pi j t}{T} \right) + b_j \cos \left( \frac{2\pi j t}{T} \right) \right] \,,
+\end{equation}
+where $N$ is the number of Fourier modes used (typically $N=30$) 
+and $T$ is the span of the observations.
+
+%The optimal statistic is constructed from the autocovariance and cross-covariance matrices 
+%${\bf{P}}_a$ and ${\bf{S}}_{ab}$. 
+%%\begin{eqnarray}
+%%	{\bf{P}}_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
+%%	{\bf{S}}_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,.
+%%\end{eqnarray}
+%%where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
+%In the frequency domain, these can be written as
+%\begin{eqnarray}
+%	{\bf{P}}_a^{ij} &=& {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_a(f) \right] {\bf{R}}_a^T \,, \\
+%	{\bf{S}}_{ab}^{ij} &=& \Gamma_{ab} \; {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_b^T \,, \label{eq:Sab}
+%\end{eqnarray}
+%where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
+%two observations, 
+%$\mathbf{R}$ is the timing model fitting matrix, 
+%$P_\mathrm{gw}(f)$ is the GW power spectrum, 
+%and $P_a(f)$ is the power spectrum for the $a$th pulsar. 
+%$\Gamma_{ab}$ is the overlap reduction function (ORF) 
+%that describes the spatial correlation between two pulsars' residuals \citep{hd1983}, 
+%\begin{eqnarray}
+%	\Gamma_{ab} &=& \frac{1}{2} \left[ 1 - \frac{1}{2} \left( \frac{1-\cos\theta_{ab}}{2} \right) \right. \nonumber \\
+%		&& \left. + 3 \left( \frac{1-\cos\theta_{ab}}{2} \right) \ln \left( \frac{1-\cos\theta_{ab}}{2} \right) \right]  \,,
+%\end{eqnarray}
+%where $\theta_{ab}$ is the angle between the pulsars. 
+%The optimal statistic $\hat{A}^2$ is given by
+%\begin{equation}
+%	\hat{A}^2 = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) } \,,
+%\end{equation}
+%where $\tilde{{\bf{S}}}_{ab}$ is the amplitude-independent cross-correlation matrix,
+%\begin{equation}
+%	\Agw^2 \tilde{{\bf{S}}}_{ab} = {\bf{S}}_{ab} \,.
+%\end{equation}
+%This definition of the optimal statistic ensures that  
+%$\langle \hat{A}^2 \rangle = \Agw^2$. 
+%If $\Agw=0$, the variance of the optimal statistic is
+%\begin{equation}
+%	\sigma_0 = \left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{-1/2} \,.
+%\end{equation}
+%For a measured value of $\hat{A}^2$, 
+%the significance of $\hat{A}^2 \neq 0$ is given by the signal-to-noise ratio (SNR) 
+%\begin{equation}
+%	\rho = \frac{\hat{A}^2}{\sigma_0} = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2}} \,.
+%\end{equation}
+%%The expectation value of the SNR, averaged over all realizations of the stochastic background, is
+%%\begin{equation}
+%%	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{ab} \Tr\left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2} \,.
+%%\end{equation}
 
 The optimal statistic is constructed from the autocovariance and cross-covariance matrices 
-${\bf{P}}_a$ and ${\bf{S}}_{ab}$. 
-%\begin{eqnarray}
-%	{\bf{P}}_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
-%	{\bf{S}}_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,.
-%\end{eqnarray}
-%where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
-In the frequency domain, these can be written as
+$C_a$ and $S_{ab}$, 
 \begin{eqnarray}
-	{\bf{P}}_a^{ij} &=& {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_a(f) \right] {\bf{R}}_a^T \,, \\
-	{\bf{S}}_{ab}^{ij} &=& \Gamma_{ab} \; {\bf{R}}_a \left[ \int_{-\infty}^\infty df \; e^{2\pi i f \tau_{ij}} P_\mathrm{gw}(f) \right] {\bf{R}}_b^T \,, \label{eq:Sab}
+	C_a &=& \left\langle \res_a \res_a^T \right\rangle \,, \\
+	S_{ab} &=& \left. \left\langle \res_a \res_b^T \right\rangle \right|_{a \neq b} \,,
 \end{eqnarray}
-where $i,j$ index the matrix elements, $\tau_{ij} \equiv |t_i - t_j|$ is the time interval between 
-two observations, 
-$\mathbf{R}$ is the timing model fitting matrix, 
-$P_\mathrm{gw}(f)$ is the GW power spectrum, 
-and $P_a(f)$ is the power spectrum for the $a$th pulsar. 
-$\Gamma_{ab}$ is the overlap reduction function (ORF) 
-that describes the spatial correlation between two pulsars' residuals \citep{hd1983}, 
+where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
+Assuming a GWB with power spectrum $P_\mathrm{gw}(f)$ 
+and overlap reduction function (ORF) $\Gamma_{ab}$, 
+\begin{equation}
+	S_{ab} = F_a \, \phi_{ab}^\mathrm{gw} \, F_b^T \,,
+\end{equation}
+where
+\begin{equation}
+	\phi_{ab}^\mathrm{gw} = \Gamma_{ab} \, P_\mathrm{gw}(f) \,.
+	\label{eq:phi_gw}
+\end{equation}
+For a GWB, the ORF is the Hellings-Downs curve \citep{hd1983}, 
 \begin{eqnarray}
@@
 \begin{equation}
-	\hat{A}^2 = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) } \,,
+	\hat{A}^2 = \frac{\sum_{ab} \res_a^T {C_a^{-1} \tilde{S}_{ab} C_b^{-1} \res_b}}{\sum_{ab} \Tr \left( C_a^{-1} \tilde{S}_{ab} C_b^{-1} \tilde{S}_{ba} \right) } \,,
 \end{equation}
-where $\tilde{{\bf{S}}}_{ab}$ is the amplitude-independent cross-correlation matrix,
+where $\tilde{S}_{ab}$ is the amplitude-independent cross-correlation matrix,
 \begin{equation}
-	\Agw^2 \tilde{{\bf{S}}}_{ab} = {\bf{S}}_{ab} \,.
+	\Agw^2 \tilde{S}_{ab} = S_{ab} \,.
 \end{equation}
@@
 \begin{equation}
-	\sigma_0 = \left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{-1/2} \,.
+	\sigma_0 = \left[ \sum_{ab} \Tr \left( C_a^{-1} \tilde{S}_{ab} C_b^{-1} \tilde{S}_{ba} \right) \right]^{-1/2} \,.
 \end{equation}
@@
 \begin{equation}
-	\rho = \frac{\hat{A}^2}{\sigma_0} = \frac{\sum_{ab} \res_a^T {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \res_b}{\left[ \sum_{ab} \Tr \left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2}} \,.
+	\rho = \frac{\sum_{ab} \res_a^T {C_a^{-1} \tilde{S}_{ab} C_b^{-1} \res_b}}{ \left[ \sum_{ab} \Tr \left( C_a^{-1} \tilde{S}_{ab} C_b^{-1} \tilde{S}_{ba} \right) \right]^{1/2}} \,.
 \end{equation}
-%The expectation value of the SNR, averaged over all realizations of the stochastic background, is
-%\begin{equation}
-%	\left\langle \rho \right\rangle = \Agw^2 \left[ \sum_{ab} \Tr\left( {\bf{P}}_a^{-1} \tilde{{\bf{S}}}_{ab} {\bf{P}}_b^{-1} \tilde{{\bf{S}}}_{ba} \right) \right]^{1/2} \,.
-%\end{equation}
 
@@
 because using a different spatial correlation only requires changing the ORF 
-in Eq.~\eqref{eq:Sab}. 
+in Eq.~\eqref{eq:phi_gw}. 
 \citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 

### 2018-05-07 manuscript/manuscript.tex
added Kristina's comments

@@
 and we use the noise-marginalized optimal statistic to assess the 
-significance of different spatial correlations and 
-by performing sky scrambles.
+significance of the spatial correlations 
+by comparing monopole, dipole, and Hellings-Downs spatial correlations 
+and performing sky scrambles.
 \end{abstract}
@@
 which measure the difference between the expected and observed pulse times of arrival. 
-Current upper-limits on the stochastic background from PTAs are approaching 
+Current upper limits on the stochastic background from PTAs are approaching 
 theoretical predictions for the GWB \citep{PPTA2013,EPTA2015,abb+17b}
 
-The primary data analysis technique used by PTAs is Bayesian inference 
-to compare the probabilities of various models for the pulsar timing residuals, 
+PTAs primarily use Bayesian data analysis to compare the inferred probabilities 
+of various models for the pulsar timing residuals, 
 including one where the residuals contain evidence of the GWB 
@@
 However, running a full Bayesian analysis is computationally intensive, 
-particularly when searching for evidence of dipole spatial correlations:
-``the smoking gun'' of the GWB.
+particularly when searching for evidence of Hellings-Downs spatial correlations --
+the ``smoking gun'' of the GWB.
 
@@
 for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
-in the NANOGrav 11-year dataset \citep{abb+17b}.
+in the NANOGrav 11-year data set \citep{abb+17b}.
 
@@
 to the optimal statistic computed using fixed noise. 
-In Sec.~\ref{sec:spatial} determine how well 
+In Sec.~\ref{sec:spatial} we determine how well 
 the noise-marginalized optimal statistic can 
@@
 We use these methods to compute the optimal statistic 
-for simulated ``NANOGrav-like'' datasets 
+for simulated ``NANOGrav-like'' data sets 
 consisting of 18 MSPs with observation times up to 11 years 
@@
 	\caption{Optimal statistic 
-			for a simulated PTA dataset containing a GWB with $\Agw = 5\times10^{-15}$. 
+			for a simulated PTA data set containing a GWB with $\Agw = 5\times10^{-15}$. 
 			The fixed-noise analysis using the individual noise values (dashed blue lines) 
@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
-	\caption{Optimal statistic and SNR for 200 simulated datasets 
+	\caption{Optimal statistic and SNR for 200 simulated data sets 
 			containing a GWB with $\Agw = 5\times10^{-15}$. 
@@
 for $\Agw = 5\times10^{-15}$, 
-but there are differences between them when analyzing datasets 
+but there are differences between them when analyzing data sets 
 containing smaller injected values of $\Agw$. 
@@
 If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
-the curves should lie along a straight line with slope 1 (the dotted, diagonal lines in Fig.~\ref{fig:os_compare}).
+the curves should lie along a straight line with slope equal to unity (the dotted, diagonal lines in Fig.~\ref{fig:os_compare}).
 
@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
-	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated data sets 
 			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
@@
 We required the scrambled ORFs to have $\bar{M} < 0.2$ 
-with respect to the true ORF and each other 
-in order to minimize the overlap between the ORFs. 
+with respect to the true ORF and each other. 
+This threshold was chosen to be comparable to the match statistics 
+between the Hellings-Downs ORF with monopole and dipole ORFs 
+given in Sec.~\ref{sec:spatial}.
 For each simulation, 
@@
 Figure~\ref{fig:skyscrambles_dataset_sample} shows the results of a sky scramble analysis 
-for a sample dataset with $\Agw=5\times10^{-15}$. 
+for a sample data set with $\Agw=5\times10^{-15}$. 
 For this particular realization of the GWB, 
@@
 	\caption{Comparison between the noise-marginalized mean optimal statistic and mean SNR 
-			with and without sky scrambles for a simulated dataset 
+			with and without sky scrambles for a simulated data set 
 			containing a GWB with $\Agw = 5\times10^{-15}$. 

### 2018-05-07 manuscript/manuscript.tex
edits to abstract and introduction

@@
 which emit low-frequency gravitational waves (GWs). 
-The incoherent superposition of these GW sources produce a stochastic GW background (GWB) 
+The incoherent superposition of these sources produce a stochastic GW background (GWB) 
 that can be observed by pulsar timing arrays (PTAs). 
-The optimal statistic is a frequentist estimator for the amplitude of a GWB 
+The optimal statistic is a frequentist estimator of the amplitude of the GWB 
 that specifically looks for the spatial correlations induced by the GWB in PTA data. 
@@
 We use simulations to demonstrate that this method 
-more accurately determines the strength of a GWB, 
-and we use the noise-marginalized optimal statistic to assess the 
-significance of the spatial correlations 
-by comparing monopole, dipole, and Hellings-Downs spatial correlations 
-and performing sky scrambles.
+more accurately determines the strength of the GWB, 
+and we use the noise-marginalized optimal statistic 
+to compare the significance of monopole, dipole, and Hellings-Downs spatial correlations 
+and perform sky scrambles.
 \end{abstract}
@@
 PTAs primarily use Bayesian data analysis to compare the inferred probabilities 
-of various models for the pulsar timing residuals, 
-including one where the residuals contain evidence of the GWB 
+of various models for the residuals, 
+including one where they contain evidence of the GWB 
 \citep{vlm+2009,lah+2013}. 
@@
 %several weeks to run on a supercomputing cluster.
-Not only does it provide an independent detection procedure, complementing a more robust Bayesian analysis, but its calculation requires significantly less time to compute.
-In particular, the optimal statistic produces results for various spatial correlation functions, a bottleneck within the more traditional stochastic GWB analysis, within seconds; a full Bayesian analysis including correlations has to run for several weeks on a supercomputing cluster.
+Not only does it provide an independent detection procedure, complementing a more robust Bayesian analysis, but it requires significantly less time to compute.
+In particular, the optimal statistic produces results for various spatial correlation functions within seconds; a full Bayesian analysis including correlations has to run for many weeks on a supercomputing cluster.
 
@@
 Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
-can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), 
-allowing for quick testing of various spatial correlations. 
+can be used to compute the optimal statistic for any choice of spatial correlations 
+simply by changing the overlap-reduction function (ORF). 
 This technique is used to perform the frequentist searches 
@@
 We use simulations to compare the noise-marginalized optimal statistic 
-to the optimal statistic computed using fixed noise. 
+to the optimal statistic computed with fixed noise. 
 In Sec.~\ref{sec:spatial} we determine how well 
@@
 differentiate between monopole, dipole, and Hellings-Downs spatial correlations. 
-In Sec.~\ref{sec:skyscrambles} we determine how well the noise-marginalized optimal statistic 
-can assess the significance of a GW detection by performing sky scrambles \citep{cs2016, tlb+2017}. 
-%We also compare our sky scramble results to the Bayesian analysis of sky scrambles done by \citet{tlb+2017}. 
+In Sec.~\ref{sec:skyscrambles} we use the noise-marginalized optimal statistic 
+to perform sky scrambles, 
+which assess the significance of Hellings-Downs spatial correlations 
+by scrambling the pulsars' sky positions \citep{cs2016, tlb+2017}. 
 We summarize our results in Sec.~\ref{sec:conclusion} 

### 2018-05-07 manuscript/manuscript.tex
more edits, including discussion of red noise in pulsars

@@
 and the GWB amplitude. 
+Many individual pulsars show evidence for red noise \citep{lam+2017,abb+17}, 
+and uncertainty in the position of the Solar System barycenter leads to 
+a common red process in all pulsars \citep{abb+17b}. 
 %In this paper, we present a technique for improving the accuracy of the optimal statistic 
@@
 where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
-Assuming a GWB with power spectrum $P_\mathrm{gw}(f)$ 
+For the GWB with power spectrum $P_\mathrm{gw}(f)$ 
 and overlap reduction function (ORF) $\Gamma_{ab}$, 
+the cross-covariance matrices are
 \begin{equation}
@@
 \begin{equation}
-	\phi_{ab}^\mathrm{gw} = \Gamma_{ab} \, P_\mathrm{gw}(f) \,.
+	\phi_{ab}^\mathrm{gw} = \Gamma_{ab} \, P_\mathrm{gw}(f) \,,
 	\label{eq:phi_gw}
 \end{equation}
-For a GWB, the ORF is the Hellings-Downs curve \citep{hd1983}, 
+and the ORF is the Hellings-Downs curve \citep{hd1983}, 
 \begin{eqnarray}
@@
 where $\theta_{ab}$ is the angle between the pulsars. 
+We model the power-spectrum of the GWB as a power law:
+\begin{equation}
+	P_\mathrm{gw}(f) = \Agw \left( \frac{f}{f_\mathrm{yr}} \right)^{-\gamma} \,,
+\end{equation}
+where $\gamma = 13/3$ assuming SMBHBs evolve solely due to GW emission and $f_\mathrm{yr} \equiv 1/(1\,\mathrm{yr})$. 
 The optimal statistic $\hat{A}^2$ is given by
@@
 First, we fix the individual pulsars' red noise parameters to the 
-maximum-likelihood values from individual pulsar analyses. 
-Second, we fix the individual pulsars' red noise parameters to the 
-values that jointly maximize the likelihood for a common analysis 
+maximum-likelihood values from individual Bayesian pulsar noise analyses. 
+Second, we fix the pulsars' red noise parameters to the 
+values that jointly maximize the likelihood for a Bayesian analysis 
 of all of the pulsars in our PTA 
@@
 a common red process. 
-For the noise-marginalized method, we draw values of the individual pulsars' 
-red noise parameters from the posteriors generated by a Bayesian analysis 
-that includes both individual red noise and a common red noise process.
+For the noise-marginalized method, we draw values of the pulsars' 
+red noise parameters from the posteriors generated by the common Bayesian analysis.
 
@@
 The fixed-noise analysis using the individual noise results also gives a significantly lower SNR 
-than the analyses using the common noise results. 
+than the other two. 
 %The mean SNR from the noise-marginalized analysis is smaller 
@@
 The fixed-noise optimal statistic using the common noise results 
-estimates $\hat{A}^2$ well for large values of $\Agw$, but for small values it also underestimates $\hat{A}^2$ 
+recovers $\hat{A}^2$ well for large values of $\Agw$, but for small values it also underestimates $\hat{A}^2$ 
 (Fig.~\ref{fig:os_compare}, middle panel). 
@@
 Searching for these using a full Bayesian approach is computationally expensive, 
-requiring \sv{estimate of CPU time}. 
-In contrast, the optimal statistic can be computed in minutes. 
+requiring many weeks on a supercomputing cluster. 
+In contrast, the optimal statistic can be computed in seconds. 
 In this paper, we introduce an improved method 
@@
 The noise marginalization technique described in this paper is key 
-to being able to accurately measure the GWB for PTA data 
-where significant red noise is present.
+to being able to accurately measure the GWB with the optimal statistic 
+for real PTAs and realistic PTA simulations, 
+for which red noise is significant.
 

### 2018-05-14 manuscript/manuscript.tex
incorporated Steve's comments into manuscript

@@
 
-\title{The noise-marginalized optimal statistic: A robust hybrid frequentist-Bayesian statistic for a stochastic gravitational-wave background in pulsar timing arrays}
+\title{The noise-marginalized optimal statistic: A robust hybrid frequentist-Bayesian statistic for the stochastic gravitational-wave background in pulsar timing arrays}
 %\title{Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic}
@@
 
-\author{Kristina P.\ Islo}
+\author{Kristina Islo}
 \affiliation{Center for Gravitation, Cosmology and Astrophysics, University of Wisconsin--Milwaukee, PO Box 413, Milwaukee WI, 53201, USA}
@@
 \author{Stephen R.\ Taylor}
+\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, California 91125, USA}
 \affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, California 91106, USA}
-\affiliation{Theoretical AstroPhysics Including Relativity (TAPIR), MC 350-17, California Institute of Technology, Pasadena, California 91125, USA}
 
@@
 The optimal statistic is a frequentist estimator of the amplitude of the GWB 
-that specifically looks for the spatial correlations induced by the GWB in PTA data. 
+that specifically looks for the spatial correlations between pulsars induced by the GWB. 
 In this paper, we introduce an improved method for computing the optimal statistic 
@@
 \citep{1995ApJ...446..543R, 2003ApJ...583..616J, 2003ApJ...590..691W}. 
-By monitoring the ultra-regular periodic emission from these pulsars using radio telescopes, 
+By monitoring the periodic emission from these pulsars using radio telescopes, 
 we can probe the dynamics of the spacetime through which the pulses travel. 
@@
 Bayesian inference is a powerful tool because 
-it performs a non-linear fit and 
-accounts for degeneracies between parameters. 
+it properly accounts for degeneracies between parameters 
+and incorporate all sources of uncertainty into the analysis. 
 However, running a full Bayesian analysis is computationally intensive, 
-particularly when searching for evidence of Hellings-Downs spatial correlations --
+particularly when searching for evidence of Hellings-Downs (HD) spatial correlations --
 the ``smoking gun'' of the GWB.
@@
 
-However, the optimal statistic gives biased results 
-when the pulsars have significant red noise 
-due to the strong covariance between the individual red noise parameters 
-and the GWB amplitude. 
+However, when pulsars have significant red noise, 
+the optimal statistic gives biased results 
+due to the strong covariance 
+between the individual red noise parameters and the GWB amplitude. 
 Many individual pulsars show evidence for red noise \citep{lam+2017,abb+17}, 
-and uncertainty in the position of the Solar System barycenter leads to 
+and uncertainty in the position of the Solar System barycenter (SSB) leads to 
 a common red process in all pulsars \citep{abb+17b}. 
@@
 simply by changing the overlap-reduction function (ORF). 
+For example, clock errors lead to a common red signal with monopole spatial correlations \citep{hcm+2012}, 
+while uncertainty in the SSB produces dipole spatial correlations \citep{chm+2010}. 
 This technique is used to perform the frequentist searches 
-for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
+for common red signals with Hellings-Downs, 
+monopole, and dipole spatial correlations 
 in the NANOGrav 11-year data set \citep{abb+17b}.
@@
 We summarize our results in Sec.~\ref{sec:conclusion} 
-and discuss future applications of the noise-marginalized optimal statistic.
+as well as discuss future applications of the noise-marginalized optimal statistic.
 
@@
 and can be derived by analytically maximizing 
-the PTA likelihood function in the weak-signal regime (see~\citet{abc+2009}). 
+the PTA likelihood function in the weak-signal regime \citep{abc+2009,ccs+2015}. 
 It is constructed from the timing residuals $\res$, 
@@
 \end{equation}
-The design matrix $M$ and timing model parameter uncertainties $\epsilon$ 
-describe the contributions to the residuals from uncertainties in the timing model. 
+The term $M\epsilon$ describes the contributions to the residuals 
+from perturbations to the timing model. 
 The term $U{\bf{j}}$ describes noise that is correlated for observations made at the same time 
-at different frequencies and uncorrelated over different observing epochs, 
-while ${\bf{n}}$ describes uncorrelated white noise.
+at different frequencies and uncorrelated over different observing epochs (ECORR), 
+while ${\bf{n}}$ describes uncorrelated white noise from TOA measurement uncertainties.
 The term $F{\bf{a}}$ describes red noise, 
 including both red noise intrinsic to the pulsar 
-and a common red noise signal common to all the pulsar (such as a GW signal). 
+and a common red noise signal common to all pulsars (such as a GW signal). 
 We model the red noise as a Fourier series,
@@
 where $\res_a$ is a vector of the residuals of the $a$th pulsar in the PTA. 
-For the GWB with power spectrum $P_\mathrm{gw}(f)$ 
+For the GWB with power spectral density (PSD) $P_\mathrm{gw}(f)$ 
 and overlap reduction function (ORF) $\Gamma_{ab}$, 
@@
 where $\theta_{ab}$ is the angle between the pulsars. 
-We model the power-spectrum of the GWB as a power law:
+We model the PSD of the GWB as a power law:
 \begin{equation}
-	P_\mathrm{gw}(f) = \Agw \left( \frac{f}{f_\mathrm{yr}} \right)^{-\gamma} \,,
+	P_\mathrm{gw}(f) = \frac{\Agw^2}{12\pi^2} \left( \frac{f}{f_\mathrm{yr}} \right)^{-\gamma} \,,
 \end{equation}
@@
 we typically fix the red noise parameters to the values that 
-maximize the likelihood. 
+maximize the single-pulsar likelihood. 
 However, this leads to a bias in the optimal statistic 
 because the individual red noise and common red noise parameters 
-are highly covariant. 
+are highly covariant, 
+with the optimal statistic systematically lower than the true value of $\Agw^2$. 
 In this section, we compare three techniques for computing the optimal statistic. 
@@
 We include white noise for all pulsars, plus 
-red noise parametrized as a power law, $P_a(f) = A_\mathrm{red} f^{-\gamma}$, 
+red noise parametrized as a power law, 
+\begin{equation}
+	P_a(f) = \frac{A_\mathrm{red}^2}{12\pi^2} \left( \frac{f}{f_\mathrm{yr}} \right)^{-\gamma} \,,
+\end{equation}
 for those pulsars that show evidence of red noise 
@@
 
-The fixed-noise and noise-marginalized analyses give the same results 
+The fixed-noise and noise-marginalized analyses 
+using the common noise results give the same results 
 for $\Agw = 5\times10^{-15}$, 
@@
 of the cumulative fraction of simulations for which the injected $\Agw^2$ lies within 
-a given confidence interval of the measured $\hat{A}^2$, 
-assuming a Gaussian distribution with mean equal to the measured $\hat{A}^2$ 
-and variance $\sigma^2_{\hat{A}^2}$, 
-determined from the distribution of $\hat{A}^2$ for our simulations. 
+a given confidence interval of the measured $\hat{A}^2$. 
+The confidence interval of $\hat{A}^2$ is determined 
+assuming $\hat{A}^2$ follows a Gaussian distribution, 
+with mean and variance $\sigma^2_{\hat{A}^2}$ 
+taken from the distribution for $\hat{A}^2$ 
+found from our 200 realizations of the GWB (i.e.,~the top panel of Fig.~\ref{fig:os_datasetstats}).
 If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
@@
 3\% and 3.5\% of our simulations, respectively.
-\begin{figure}[tb]
+\begin{figure}[t]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
@@
 The significance of spatial correlations can also be tested with ``sky scrambles,'' 
-where the ORF is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
+where the ORF is altered in order to simulate changing the pulsars' positions \citep{cs2016,tlb+2017}. 
+The scrambled ORFs are required to have small values of $\bar{M}$ 
+with the true ORF and each other so that they form a nearly-orthogonal set. 
+This ensures that the distribution of $\hat{A}^2$ computed using the scrambled ORFs 
+forms the null hypothesis distribution. 
 \citet{tlb+2017} showed how sky scrambles affect the Bayes' factor for simulated data sets. 
@@
 given in Sec.~\ref{sec:spatial}. 
-For each simulation, 
-we computed the noise-marginalized mean optimal statistic and mean SNR 
-for each scrambled ORF, and compared to the values found using the true ORF.
+We did not chose a smaller threshold because significantly more time 
+would have been needed to generate 725 scrambled ORFs. 
+For each simulation, we computed the 
+noise-marginalized mean optimal statistic and mean SNR 
+for each scrambled ORF, 
+and compared them to the values found using the true ORF.
 
@@
 and 33\% have $p \leq 0.003$. 
-\begin{figure}[tb]
+This shows that for smaller values of \Agw, 
+there is a greater chance that noise fluctuations will appear 
+to have the spatial correlations of the GWB.
+\begin{figure}[t]
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_scrambled_dataset11.pdf}
@@
 as with the sky scrambles. 
+An upcoming paper will use the noise-marginalized optimal statistic 
+to determine how well the spatial correlations corresponding to alternate GW polarizations 
+can be measured. 
 It also makes the optimal statistic a valuable tool for analyzing simulations 
@@
 This work was supported by NSF Physics Frontier Center Grant 1430284. 
-SRT was partially supported by appointment to the NASA Postdoctoral Program 
-at the Jet Propulsion Laboratory, which is administered by Oak Ridge Associated Universities 
-and the Universities Space Research Association through a contract with NASA. 
+%SRT was partially supported by appointment to the NASA Postdoctoral Program 
+%at the Jet Propulsion Laboratory, which is administered by Oak Ridge Associated Universities 
+%and the Universities Space Research Association through a contract with NASA. 
 JAE was partially supported by NASA through Einstein Fellowship Grant PF4-150120.

### 2018-05-14 codes/optstat_v3.py
edits to main file

@@ -4,9 +4,24 @@ import h5py as h5
 import scipy.interpolate as interp
 import scipy.ndimage.filters as filter
 
+from astropy.io import fits
+from astropy.table import Table
+
 from PAL2 import PALmodels, PALutils, pputils as pp
 
 import makesims_nano
+
+
+def load_fits_file(filename):
+    
+    myfile = fits.open(filename)
+    
+    table = Table(myfile[1].data)
+    pars = table.colnames
+    data = np.array(table)
+    chain = np.array([data[ii] for ii in pars]).T
+    
+    return pars, chain
 
 
 def get_ml_vals(cp):
@@ -257,18 +272,24 @@ def compute_os(orf, rho, sig):
     return opt, sig
 
 
-# reads in sky scramble positions from files provided by Steve Taylor
 def read_in_skyscrambles(directory, nscrambles):
+
+    data = np.load('../data/scrambles_nano_thresh0.2.npz')
     
-    orfs = []
-    
-    for n in range(nscrambles):
-        data = np.loadtxt(directory + 'PositionSet_{0:d}.dat'.format(n),
-                          skiprows=1, usecols=(1,2))
-                          
-        orfs.append(compute_orf(data[:,1], data[:,0]))
-    
-    return orfs
+    return data['orfs'][:nscrambles]
+
+# reads in sky scramble positions from files provided by Steve Taylor
+#def read_in_skyscrambles(directory, nscrambles):
+#
+#    orfs = []
+#
+#    for n in range(nscrambles):
+#        data = np.loadtxt(directory + 'PositionSet_{0:d}.dat'.format(n),
+#                          skiprows=1, usecols=(1,2))
+#
+#        orfs.append(compute_orf(data[:,1], data[:,0]))
+#
+#    return orfs
 
 
 # initialize the optimal statistic using the noise values from file
@@ -637,7 +658,10 @@ if __name__ == '__main__':
         os.makedirs(outputdir)
 
     chaindir = '../data/chains/' + dataset + '/fix_spec_nf_30/'
-    chain = np.loadtxt(chaindir + 'chain_1.txt')
+    if os.path.isfile(chaindir + 'chain.fits'):
+        _, chain = load_fits_file(chaindir + 'chain.fits')
+    else:
+        chain = np.loadtxt(chaindir + 'chain_1.txt')
     burn = int(0.25*chain.shape[0])
 
[…]

### 2018-05-15 codes/makesims_nano_dipole.py
added script that makes simulations with an injected dipole red process

@@ -1,0 +1,224 @@
+# modification of simulation code that injects a common signal
+# with power-law PSD and dipole spatial correlations
+
+import numpy as np
+import os,sys,glob,math
+import json
+
+import ephem
+import libstempo.toasim as LT
+from libstempo import spharmORFbasis as anis
+
+
+def createdipole(psr, Amp, gam, noCorr=False, seed=None, turnover=False,
+                 clm=[np.sqrt(4.0*np.pi)], lmax=0, f0=1e-9, beta=1,
+                 power=1, userSpec=None, npts=600, howml=10):
+    """
+        Function to create GW-induced residuals from a dipole background as defined
+        in Chamberlin, Creighton, Demorest, et al. (2014)
+        with modification to the ORF by SJV
+        
+        :param psr: pulsar object for single pulsar
+        :param Amp: Amplitude of red noise in GW units
+        :param gam: Red noise power law spectral index
+        :param noCorr: Add red noise with no spatial correlations
+        :param seed: Random number seed
+        :param turnover: Produce spectrum with turnover at frequency f0
+        :param clm: coefficients of spherical harmonic decomposition of GW power
+        :param lmax: maximum multipole of GW power decomposition
+        :param f0: Frequency of spectrum turnover
+        :param beta: Spectral index of power spectram for f << f0
+        :param power: Fudge factor for flatness of spectrum turnover
+        :param userSpec: User-supplied characteristic strain spectrum
+        (first column is freqs, second is spectrum)
+        :param npts: Number of points used in interpolation
+        :param howml: Lowest frequency is 1/(howml * T)
+        
+        :returns: list of residuals for each pulsar
+        """
+    
+    if seed is not None:
+        np.random.seed(seed)
+    
+    # number of pulsars
+    Npulsars = len(psr)
+
+    # gw start and end times for entire data set
+    start = np.min([p.toas().min()*86400 for p in psr]) - 86400
+    stop = np.max([p.toas().max()*86400 for p in psr]) + 86400
+    
+    # duration of the signal
+    dur = stop - start
+    
+    # get maximum number of points
+    if npts is None:
[…]

### 2018-05-17 manuscript/manuscript.tex
updated sky scramble plot and stats

@@
 In Fig.~\ref{fig:skyscrambles_pvalues}, we plot the distribution of $p$-values of the 725 sky scrambles. 
-For a GWB with $\Agw = 5\times10^{-15}$, 96\% of the simulations have $p \leq 0.05$, 
-and 75\% of the simulations have $p \leq 0.003$. 
-For a GWB with $\Agw = 10^{-15}$, 60\% of the simulations have $p \leq 0.05$, 
-and 33\% have $p \leq 0.003$. 
+For a GWB with $\Agw = 5\times10^{-15}$, 95\% of the simulations have $p \leq 0.05$, 
+and 74\% of the simulations have $p \leq 0.003$. 
+For a GWB with $\Agw = 10^{-15}$, 76\% of the simulations have $p \leq 0.05$, 
+and 39\% have $p \leq 0.003$. 
 This shows that for smaller values of \Agw, 

### 2018-05-17 manuscript/manuscript.tex
updated Table 1 with pulsar parameters
added results from dipole injections

@@
 	\setlength{\tabcolsep}{5pt}
-	\caption{Pulsar parameters used in simulated PTAs. \sv{finish filling in this table}}
+	\caption{Pulsar parameters used in simulated PTA data sets.}
 	\begin{center}
@@
 		\hline
-		J0030+0451 & 11.0 & & -13.93 & 3.56 \\
-		J0613$-$0200 & 11.0 &  & -13.14 & 1.22 \\
-		J1012+5307 & 11.0 &  & -12.79 & 1.51 \\
-		J1024$-$0719 & 6.0 & & $-$ & $-$ \\
-		J1455$-$3330 & 11.0 & & $-$ & $-$ \\
-		J1600$-$3053 & 8.0 & & $-$ & $-$ \\
-		J1614$-$2230 & 7.0 & & $-$ & $-$ \\
-		J1640+2224 & 11.0 & & $-$ & $-$ \\
-		J1713+0747 & 11.0 & & -14.14 & 1.58 \\
-		J1741+1351 & 6.0 & & $-$ & $-$ \\
-		J1744$-$1134 & 11.0 & & $-$ & $-$ \\
-		B1855+09 & 11.0 & & -13.75 & 3.54 \\
-		J1853+1303 & 7.0 & & $-$ & $-$ \\
-		J1909$-$3744 & 11.0 &  & -13.84 & 1.74  \\
-		J1918$-$0642 & 11.0 & & $-$ & $-$ \\
-		J2010$-$1323 & 7.0 & & $-$ & $-$ \\
-		J2145$-$0750 & 11.0 &  & -12.69 & 1.30 \\
-		J2317+1439 & 11.0 & & $-$ & $-$ \\
+		J0030+0451 & 11.0 & 0.339 & -13.93 & 3.56 \\
+		J0613$-$0200 & 11.0 & 0.281 & -13.14 & 1.22 \\
+		J1012+5307 & 11.0 & 0.320 & -12.79 & 1.51 \\
+		J1024$-$0719 & 6.0 &0.421 & $-$ & $-$ \\
+		J1455$-$3330 & 11.0 & 0.773 & $-$ & $-$ \\
+		J1600$-$3053 & 8.0 & 0.146 & $-$ & $-$ \\
+		J1614$-$2230 & 7.0 & 0.261 & $-$ & $-$ \\
+		J1640+2224 & 11.0 & 0.202 & $-$ & $-$ \\
+		J1713+0747 & 11.0 & 0.093 & -14.14 & 1.58 \\
+		J1741+1351 & 6.0 & 0.106 & $-$ & $-$ \\
+		J1744$-$1134 & 11.0 & 0.096 & $-$ & $-$ \\
+		B1855+09 & 11.0 & 0.218 & -13.75 & 3.54 \\
+		J1853+1303 & 7.0 & 0.215 & $-$ & $-$ \\
+		J1909$-$3744 & 11.0 & 0.034  & -13.84 & 1.74  \\
+		J1918$-$0642 & 11.0 & 0.342 & $-$ & $-$ \\
+		J2010$-$1323 & 7.0 & 0.413 & $-$ & $-$ \\
+		J2145$-$0750 & 11.0 & 0.281 & -12.69 & 1.30 \\
+		J2317+1439 & 11.0 & 0.160 & $-$ & $-$ \\
     		\hline\hline
@@
 			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
-			We compare the values found using monopole (blue), dipole (orange), 
+			We compare the values of the mean optimal statistic and the SNR found 
+			using monopole (blue), dipole (orange), and quadrupole (green) spatial correlations. 
+			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
+	\label{fig:os_ORF}
+\end{figure}
+
+This overlap between the monopole, dipole, and quadrupole ORFs also means that 
+a common red process that does not have Hellings-Downs correlations  
+may be confused for a GWB. Figure~\ref{fig:os_dipole} shows the results of 25 simulations 
+containing a stochastic signal with dipole spatial correlations. 
+Although a dipole signal has been injected, 
+the Hellings-Downs ORF gives a mean SNR greater than 5 in 84\% of the simulations. 
+However, both the monopole and Hellings-Downs ORFs give smaller values of 
+the mean $\hat{A}^2$ and mean SNR compared to the dipole ORF. 
+This demonstrates the importance of comparing the SNR from different spatial correlations 
+when determining the type of spatial correlations present.
+\begin{figure}[t]
+	\includegraphics[width=0.9\columnwidth]{plots/optstat_dipole_A5e-15.pdf}
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 25 simulated data sets 
+			containing an injected stochastic signal with dipole spatial correlations and $A = 5\times10^{-15}$. 
+			We compare the values of the optimal statistic and mean SNR 
+			found using monopole (blue), dipole (orange), 
 			and quadrupole (green) spatial correlations. 
 			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
-	\label{fig:os_ORF}
+	\label{fig:os_dipole}
 \end{figure}
@@
 \acknowledgments
-We thank Joe Romano, Xavier Siemens, Jolien Creighton, and Jeffrey Hazboun for useful discussions. 
+We thank Joe Romano, Xavier Siemens, and Jolien Creighton for useful discussions. 
 This work was supported by NSF Physics Frontier Center Grant 1430284. 

### 2018-05-17 manuscript/manuscript.tex
fixed wording, etc

@@
 and we use the noise-marginalized optimal statistic 
-to compare the significance of monopole, dipole, and Hellings-Downs spatial correlations 
+to compare the significance of monopole, dipole, and Hellings-Downs (HD) spatial correlations 
 and perform sky scrambles.
@@
 This technique is used to perform the frequentist searches 
-for common red signals with Hellings-Downs, 
+for common red signals with HD, 
 monopole, and dipole spatial correlations 
@@
 the noise-marginalized optimal statistic can 
-differentiate between monopole, dipole, and Hellings-Downs spatial correlations. 
+differentiate between monopole, dipole, and HD spatial correlations. 
 In Sec.~\ref{sec:skyscrambles} we use the noise-marginalized optimal statistic 
 to perform sky scrambles, 
-which assess the significance of Hellings-Downs spatial correlations 
+which assess the significance of HD spatial correlations 
 by scrambling the pulsars' sky positions \citep{cs2016, tlb+2017}. 
@@
 \end{equation}
-and the ORF is the Hellings-Downs curve \citep{hd1983}, 
+and the ORF is the HD curve \citep{hd1983}, 
 \begin{eqnarray}
@@
 In Fig.~\ref{fig:os_datasetstats} we show the optimal statistic 
-for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
+for 300 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
 computed using the three techniques described above. 
@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
-	\caption{Optimal statistic and SNR for 200 simulated data sets 
+	\caption{Optimal statistic and SNR for 300 simulated data sets 
 			containing a GWB with $\Agw = 5\times10^{-15}$. 
@@
 taken from the distribution for $\hat{A}^2$ 
-found from our 200 realizations of the GWB (i.e.,~the top panel of Fig.~\ref{fig:os_datasetstats}).
+found from our 300 realizations of the GWB (i.e.,~the top panel of Fig.~\ref{fig:os_datasetstats}).
 If $\hat{A}^2$ has a Gaussian distribution centered around $\Agw^2$, 
@@
 For the 18 pulsars used in these simulations, the 
-match statistic for monopole and Hellings-Downs correlations is $\bar{M} = 0.264$, 
-and the match statistic for dipole and Hellings-Downs correlations is $\bar{M} = 0.337$. 
+match statistic for monopole and HD correlations is $\bar{M} = 0.264$, 
+and the match statistic for dipole and HD correlations is $\bar{M} = 0.337$. 
 These match statistics describe a fundamental limit on our ability 
-to identify the spatial correlations of a common red signal as Hellings-Downs 
+to identify the spatial correlations of a common red signal as HD 
 rather than monopole or dipole 
@@
 mean value of $\hat{A}^2$ and the mean SNR 
-computed assuming monopole, dipole, and quadrupole spatial correlations 
-for 200 simulated data sets. Using a monopole or dipole ORF 
+computed assuming monopole, dipole, and HD spatial correlations 
+for 300 simulated data sets. Using a monopole or dipole ORF 
 gives a lower value for the mean optimal statistic and mean SNR compared to the 
-Hellings-Downs ORF. 
-Using quadrupole spatial correlations gives $\hat{A}^2 = (2.4 \pm 1.2) \times 10^{-29}$, 
+HD ORF. 
+Using HD spatial correlations gives $\hat{A}^2 = (2.4 \pm 1.2) \times 10^{-29}$, 
 while using monopole spatial correlations gives $\hat{A}^2 = (2.5 \pm 3.1) \times 10^{-30}$, 
@@
 We find a noise-marginalized mean SNR above 1.0 in 97\% of our simulated data sets 
-using the quadrupole ORF, and in 50\% and 68\% of our simulated data sets 
+using the HD ORF, and in 50\% and 68\% of our simulated data sets 
 using the monopole and dipole ORFs, respectively. 
-The mean SNR using the quadrupole ORF, averaged over realizations of the GWB, is 4.1, 
+The mean SNR using the HD ORF, averaged over realizations of the GWB, is 4.1, 
 and we find an SNR greater than this using the monopole and dipole ORFs in just 
@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
-	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated data sets 
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 300 simulated data sets 
 			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
 			We compare the values of the mean optimal statistic and the SNR found 
-			using monopole (blue), dipole (orange), and quadrupole (green) spatial correlations. 
+			using monopole (blue), dipole (orange), and HD (green) spatial correlations. 
 			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
@@
 
-This overlap between the monopole, dipole, and quadrupole ORFs also means that 
-a common red process that does not have Hellings-Downs correlations  
+This overlap between the monopole, dipole, and HD ORFs also means that 
+a common red process that does not have HD correlations  
 may be confused for a GWB. Figure~\ref{fig:os_dipole} shows the results of 25 simulations 
@@
 Although a dipole signal has been injected, 
-the Hellings-Downs ORF gives a mean SNR greater than 5 in 84\% of the simulations. 
-However, both the monopole and Hellings-Downs ORFs give smaller values of 
+the HD ORF gives a mean SNR greater than 5 in 84\% of the simulations. 
+However, both the monopole and HD ORFs give smaller values of 
 the mean $\hat{A}^2$ and mean SNR compared to the dipole ORF. 
@@
 			found using monopole (blue), dipole (orange), 
-			and quadrupole (green) spatial correlations. 
+			and HD (green) spatial correlations. 
 			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
@@
 This threshold was chosen to be comparable to the match statistics 
-between the Hellings-Downs ORF with monopole and dipole ORFs 
+between the HD ORF with monopole and dipole ORFs 
 given in Sec.~\ref{sec:spatial}. 
@@
 the mean SNR using the true ORF ($p < 0.0014$).
-In Fig.~\ref{fig:skyscrambles_pvalues}, we plot the distribution of $p$-values of the 725 sky scrambles. 
+In Fig.~\ref{fig:skyscrambles_pvalues}, we plot the distribution of $p$-values of the 725 sky scrambles 
+for 300 realizations of the GWB. 
 For a GWB with $\Agw = 5\times10^{-15}$, 95\% of the simulations have $p \leq 0.05$, 
@@
 	\caption{Distribution of $p$-values for the noise-marginalized optimal statistic mean SNR 
-			using the true ORF compared to 725 sky scrambles. 
+			using the true ORF compared to 725 sky scrambles from 300 realizations of the GWB. 
 			We show results for simulations with $\Agw=5\times10^{-15}$ (blue) and $\Agw=10^{-15}$ (orange).}
@@
 The definitive signature of a GWB in PTA data is spatial correlations 
-described by the Hellings-Downs curve. 
+described by the HD curve. 
 Searching for these using a full Bayesian approach is computationally expensive, 
@@
 In Sec.~\ref{sec:spatial} we use the noise-marginalized optimal statistic 
-to compare the strength of monopole, dipole, and quadrupole correlations 
+to compare the strength of monopole, dipole, and HD correlations 
 in simulated PTA data with a GWB. 

### 2018-05-19 codes/makesims_nano_dipole.py
fixed bugs

@@ -1,9 +1,13 @@
 # modification of simulation code that injects a common signal
 # with power-law PSD and dipole spatial correlations
 
+from __future__ import division
+
 import numpy as np
 import os,sys,glob,math
 import json
+
+import scipy.interpolate as interp
 
 import ephem
 import libstempo.toasim as LT
@@ -84,17 +88,18 @@ def createdipole(psr, Amp, gam, noCorr=False, seed=None, turnover=False,
     ORF = np.zeros((Npulsars,Npulsars))
     for i in range(Npulsars):
         for j in range(Npulsars):
-            zeta = anis.calczeta(psrlocs[:,0][i],psrlocs[:,0][j],
-                                 psrlocs[:,1][i],psrlocs[:,1][j])
-            ORF[i,j] = np.cos(zeta)
-        
+#            zeta = anis.calczeta(psrlocs[:,0][i],psrlocs[:,0][j],
+#                                 psrlocs[:,1][i],psrlocs[:,1][j])
+#            ORF[i,j] = np.cos(zeta)
+            if i == j:
+                # linear algebra wizardry courtesy of Steve
+                # this ensures the matrix is positive definite so the Cholesky transform can be used
+                ORF[i,j] = 1. + 1e-6
+            else:
+                ORF[i,j] = np.sin(psrlocs[i,1])*np.sin(psrlocs[j,1])*np.cos(psrlocs[i,0]-psrlocs[j,0]) + \
+                            np.cos(psrlocs[i,1])*np.cos(psrlocs[j,1])
     ORF *= 2.0
-        
-    # linear algebra wizardry courtesy of Steve
-    # this ensures the matrix is positive definite so the Cholesky transform can be used
-    for i in range(Npulsars):
-        ORF[i,i] += 1e-6
-
+   
     # Define frequencies spanning from DC to Nyquist.
     # This is a vector spanning these frequencies in increments of 1/(dur*howml).
     f = np.arange(0, 1/(2*dt), 1/(dur*howml))
@@ -203,7 +208,7 @@ def create_dataset(dataset, Agwb):
                             components=30)
 
     # dipole signal
-    createdipole(psrs, 5e-15, 13./3., seed=None)
+    createdipole(psrs, Agwb, 13./3., seed=None)
     
     for psr in psrs:
         psr.fit(iters=2)

### 2018-05-19 codes/get_skyscramble_stats.py
modified script for getting stats from sky scrambles

@@ -1,33 +1,17 @@
 import numpy as np
-import os,sys,glob,math
 
-datadir = '../data/optstat/'
-datasetname = 'dataset'
-ndatasets = 100
 
-outputfilename = datadir + datasetname + 'stats_skyscrambles.dat'
-
-f = open(outputfilename, 'w')
-
-for n in range(ndatasets):
-    name = datadir + datasetname + str(n)
-    if os.path.exists(name):
-        data = np.loadtxt(name + '/marg_os.dat')
-        snr = np.mean(data[:,2])
-        
-        nscrambles = 210
-        scrambled_snr = []
-        for N in range(nscrambles):
-            ss = np.loadtxt(name + '/marg_os_skyscramble{0}.dat'.format(N))
-            scrambled_snr.append(np.mean(ss[:,2]))
-        
-        count = 0
-        for i in range(len(scrambled_snr)):
-            if scrambled_snr[i] >= snr:
-                count += 1
+def get_skyscramble_stats(datasetname, nscrambles):
     
-        pvalue = float(count)/float(nscrambles)
+    output = open('../data/optstat_common/{0}/skyscramble_summary.dat'.format(datasetname), 'w')
 
-        f.write('{0:<10}  {1:>3.0f}  {2:.4f}\n'.format(datasetname+str(n), count, pvalue))
+    for n in range(nscrambles):
+        ss = np.loadtxt('../data/optstat_common/{0}/marg_os_skyscramble{1}.dat'.format(datasetname,n))
+        output.write('skyscramble{0:<3}  {1:>18}  {2}\n'.format(n, np.mean(ss[:,0]), np.mean(ss[:,2])))
 
-f.close()
+    output.close()
+
+
+for i in range(299,300):
+    print 'Making skyscramble summary file for dataset_nano_A1e-15_{0}'.format(i)
+    get_skyscramble_stats('dataset_nano_A1e-15_{0}'.format(i), 725)

### 2018-05-19 codes/runscript.sh
modifications to runscript.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
 # script for generating and analyzing many datasets
-# to modify, change the numbers 30 and 33 to correspond to the numbers 
+# to modify, change the numbers 1 and 9 to correspond to the numbers 
 # of the first and last set you want to generate
 
-for i in $(seq 0 1 9)
+for i in $(seq 1 1 9)
 do
-	python optstat_v2.py --datasetname dataset_nano_$i --computeMonopole --computeDipole --computeSkyScrambles
+    python optstat_v3.py --Agw 5e-15 --datasetname dataset_nano_$i --nreal 10000 --computeMonopole --computeDipole
 done

### 2018-05-19 codes/makesims_nano_monopole.py
script for making simulated data sets with an injected monopole

@@ -1,0 +1,205 @@
+# modification of simulation code that injects a common signal
+# with power-law PSD and dipole spatial correlations
+
+from __future__ import division
+
+import numpy as np
+import os,sys,glob,math
+import json
+
+import scipy.interpolate as interp
+
+import ephem
+import libstempo.toasim as LT
+from libstempo import spharmORFbasis as anis
+
+
+def createmonopole(psr, Amp, gam, noCorr=False, seed=None, turnover=False,
+                 clm=[np.sqrt(4.0*np.pi)], lmax=0, f0=1e-9, beta=1,
+                 power=1, userSpec=None, npts=600, howml=10):
+    """
+        Function to create GW-induced residuals from a dipole background as defined
+        in Chamberlin, Creighton, Demorest, et al. (2014)
+        with modification to the ORF by SJV
+        
+        :param psr: pulsar object for single pulsar
+        :param Amp: Amplitude of red noise in GW units
+        :param gam: Red noise power law spectral index
+        :param noCorr: Add red noise with no spatial correlations
+        :param seed: Random number seed
+        :param turnover: Produce spectrum with turnover at frequency f0
+        :param clm: coefficients of spherical harmonic decomposition of GW power
+        :param lmax: maximum multipole of GW power decomposition
+        :param f0: Frequency of spectrum turnover
+        :param beta: Spectral index of power spectram for f << f0
+        :param power: Fudge factor for flatness of spectrum turnover
+        :param userSpec: User-supplied characteristic strain spectrum
+        (first column is freqs, second is spectrum)
+        :param npts: Number of points used in interpolation
+        :param howml: Lowest frequency is 1/(howml * T)
+        
+        :returns: list of residuals for each pulsar
+        """
+    
+    if seed is not None:
+        np.random.seed(seed)
+    
+    # number of pulsars
+    Npulsars = len(psr)
+
+    # gw start and end times for entire data set
+    start = np.min([p.toas().min()*86400 for p in psr]) - 86400
+    stop = np.max([p.toas().max()*86400 for p in psr]) + 86400
+    
+    # duration of the signal
+    dur = stop - start
+    
[…]

### 2018-05-23 manuscript/manuscript.tex
added Joe Lazio and Andrea Lommen to acknowledgements
added text to acknowledgements for nemo

@@
 \acknowledgments
-We thank Joe Romano, Xavier Siemens, and Jolien Creighton for useful discussions. 
+We thank Joe Lazio, Andrea Lommen, Joe Romano, Xavier Siemens, and Jolien Creighton for useful discussions. 
 This work was supported by NSF Physics Frontier Center Grant 1430284. 
@@
 JAE was partially supported by NASA through Einstein Fellowship Grant PF4-150120. 
+We are grateful for computational resources provided by the Leonard E Parker
+Center for Gravitation, Cosmology and Astrophysics at the University of
+Wisconsin-Milwaukee.
 

### 2018-05-24 manuscript/manuscript.tex
updated results for dipole injection simulations

@@
 Although a dipole signal has been injected, 
-the HD ORF gives a mean SNR greater than 5 in 84\% of the simulations. 
+the HD ORF gives a mean SNR greater than 5 in 82\% of the simulations. 
 However, both the monopole and HD ORFs give smaller values of 
 the mean $\hat{A}^2$ and mean SNR compared to the dipole ORF. 
+Furthermore, there are no simulations for which the mean SNR with HD ORF 
+is greater than the mean SNR with dipole ORF. 
 This demonstrates the importance of comparing the SNR from different spatial correlations 

### 2018-05-30 manuscript/manuscript.tex
final round of edits

@@
 contain supermassive black holes (SMBHs) at their centers. 
-When galaxies merge, these SMBHs form SMBH binaries (SMBHBs), 
-which emit low-frequency gravitational waves (GWs). 
+When galaxies merge, these SMBHs form SMBH binaries (SMBHBs) 
+that emit low-frequency gravitational waves (GWs). 
 The incoherent superposition of these sources produce a stochastic GW background (GWB) 
@@
 This is done by searching for correlations in the pulsar timing residuals, 
-which measure the difference between the expected and observed pulse times of arrival. 
+which measure the differences between the expected and observed pulse times of arrival. 
 Current upper limits on the stochastic background from PTAs are approaching 
@@
 of various models for the residuals, 
-including one where they contain evidence of the GWB 
+including one where they contain the GWB 
 \citep{vlm+2009,lah+2013}. 
@@
 it properly accounts for degeneracies between parameters 
-and incorporate all sources of uncertainty into the analysis. 
+and incorporates all sources of uncertainty into the analysis. 
 However, running a full Bayesian analysis is computationally intensive, 
@@
 Not only does it provide an independent detection procedure, complementing a more robust Bayesian analysis, but it requires significantly less time to compute.
-In particular, the optimal statistic produces results for various spatial correlation functions within seconds; a full Bayesian analysis including correlations has to run for many weeks on a supercomputing cluster.
+In particular, the optimal statistic produces results for a given spatial correlation function within seconds; a full Bayesian analysis including correlations has to run for many weeks on a supercomputing cluster.
 
-However, when pulsars have significant red noise, 
+However, when pulsars have significant red noise 
 the optimal statistic gives biased results 
@@
 Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: 
-marginalizing over the individual pulsar red noise parameters 
+marginalizing over the individual pulsars' red noise parameters 
 using the posterior distributions from a full Bayesian analysis of all the pulsars. 
@@
 \begin{equation}
-	\phi_{ab}^\mathrm{gw} = \Gamma_{ab} \, P_\mathrm{gw}(f) \,,
+	\phi_{ab}^\mathrm{gw} = \Gamma_{ab} \, P_\mathrm{gw}(f) \,.
 	\label{eq:phi_gw}
 \end{equation}
-and the ORF is the HD curve \citep{hd1983}, 
+The ORF is the HD curve \citep{hd1983}, 
 \begin{eqnarray}
@@
 are highly covariant, 
-with the optimal statistic systematically lower than the true value of $\Agw^2$. 
+with the optimal statistic computed using fixed red noise parameters 
+systematically lower than the true value of $\Agw^2$. 
 In this section, we compare three techniques for computing the optimal statistic. 
@@
 for simulated ``NANOGrav-like'' data sets 
-consisting of 18 MSPs with observation times up to 11 years 
-and sky positions matching the 18 longest-observed pulsars in the 
+consisting of 18 MSPs with observation times, 
+sky positions, and noise properties 
+matching the 18 longest-observed pulsars in the 
 NANOGrav 11-year data set \citep{abb+17}. 
@@
 a common red process that does not have HD correlations  
-may be confused for a GWB. Figure~\ref{fig:os_dipole} shows the results of 25 simulations 
+may be confused for a GWB. Figure~\ref{fig:os_dipole} shows the results of 300 simulations 
 containing a stochastic signal with dipole spatial correlations. 
@@
 	\includegraphics[width=0.9\columnwidth]{plots/optstat_dipole_A5e-15.pdf}
-	\caption{Noise-marginalized mean optimal statistic and mean SNR for 25 simulated data sets 
+	\caption{Noise-marginalized mean optimal statistic and mean SNR for 300 simulated data sets 
 			containing an injected stochastic signal with dipole spatial correlations and $A = 5\times10^{-15}$. 
@@
 \citet{tlb+2017} showed how sky scrambles affect the Bayes' factor for simulated data sets. 
-Here we perform a similar analysis using frequentist methods.
+We performed a similar analysis using frequentist methods.
 
@@
 \acknowledgments
-We thank Joe Lazio, Andrea Lommen, Joe Romano, Xavier Siemens, and Jolien Creighton for useful discussions. 
+We thank Joe Lazio, Andrea Lommen, Joe Romano, Xavier Siemens, 
+and Jolien Creighton for useful discussions. 
 This work was supported by NSF Physics Frontier Center Grant 1430284. 
@@
 Center for Gravitation, Cosmology and Astrophysics at the University of
-Wisconsin-Milwaukee.
+Wisconsin-Milwaukee supported by NSF Grant 0923409.
 

### 2018-05-30 manuscript/manuscript.tex
added grant numbers

@@
 Center for Gravitation, Cosmology and Astrophysics at the University of
-Wisconsin-Milwaukee supported by NSF Grant 0923409.
+Wisconsin-Milwaukee, which is supported by NSF Grants 0923409 and 1626190.
 
